---
title: "FLoRA Preparation Pipeline"
subtitle: "Download, deduplicate, and prepare paper-level dataset (Replications + Reproductions)"
author: "FORRT Team"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 2
    theme: cosmo
    code-fold: true
    code-summary: "Show code"
params:
  include_score: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = TRUE,
  warning = TRUE,
  error = FALSE,
  cache = FALSE
)

library(tidyverse)
library(readxl)
library(jsonlite)
library(glue)
library(httr)

# Use project-root paths via here
if (!file.exists(here::here("R", "cache_config.R"))) {
  stop("Cannot find cache config. Ensure you're in the project or run from repo root.")
}

# Source Helper Functions
source(here::here("R", "cache_config.R"))
source(here::here("R", "augmentation.R"))
source(here::here("R", "crossref_cache.R"))
source(here::here("R", "openalex_cache.R"))

```

## Configuration

```{r data-sources}
FReD_FOUNDATION_SUCCESS_CODING <- "https://docs.google.com/spreadsheets/d/e/2PACX-1vSlPVImL7kjzMfpfX08JtAikvEiBg8uQ3HyEJs-YvuQMUa2RMvip-4v8-DilWl2PdwOD_UTD8gJ5wKA/pub?gid=984458430&single=true&output=csv"
REPLICATION_GSHEET_URL <- "https://docs.google.com/spreadsheets/d/e/2PACX-1vT0VnLyrf9GCYXtN6l1DgaoLlg6H5r-08Op9eJzSripS1QBSHL031Arc27yUDe0YY7cB4TOnYMm2Vh1/pub?gid=863031634&single=true&output=csv"
REPRODUCTION_GSHEET_URL <- "https://docs.google.com/spreadsheets/d/e/2PACX-1vT0VnLyrf9GCYXtN6l1DgaoLlg6H5r-08Op9eJzSripS1QBSHL031Arc27yUDe0YY7cB4TOnYMm2Vh1/pub?gid=984458430&single=true&output=csv"
EXCLUSIONS_GSHEET_URL <- "https://docs.google.com/spreadsheets/d/e/2PACX-1vT0VnLyrf9GCYXtN6l1DgaoLlg6H5r-08Op9eJzSripS1QBSHL031Arc27yUDe0YY7cB4TOnYMm2Vh1/pub?gid=1098183075&single=true&output=csv"
FLORA_OUTPUT <- here::here("output", "flora.csv")

# SCORE Integration (optional)
SCORE_INTEGRATION_ENABLED <- params$include_score
SCORE_DATA_PATH <- here::here("cos_score_integration", "cos_test_set_phase1_prepared.xlsx")

cat("\n=== FLoRA Preparation Pipeline ===\n")
cat(sprintf("  SCORE Integration: %s\n", if (SCORE_INTEGRATION_ENABLED) "ENABLED" else "DISABLED"))
```

## Download FLoRA Data (Replications + Reproductions)

```{r download-data}
# Download replications
replications <- tryCatch({
  read_csv(REPLICATION_GSHEET_URL, show_col_types = FALSE)
}, error = function(e) {
  stop("Failed to download replications data. Check URL and network connection.\n",
       "Error: ", e$message)
})

cat(sprintf("    ✓ Downloaded %d replication rows × %d columns\n", nrow(replications), ncol(replications)))

# Filter for validated entries only
replications <- replications %>%
  filter(validation_status == "validated - chosen")
cat(sprintf("    ✓ Filtered to %d validated replications\n", nrow(replications)))

# Download reproductions
reproductions <- tryCatch({
  read_csv(REPRODUCTION_GSHEET_URL, show_col_types = FALSE)
}, error = function(e) {
  stop("Failed to download reproductions data. Check URL and network connection.\n",
       "Error: ", e$message)
})

cat(sprintf("    ✓ Downloaded %d reproduction rows × %d columns\n", nrow(reproductions), ncol(reproductions)))

# Filter for validated entries only
reproductions <- reproductions %>%
  filter(validation == "validated - chosen")
cat(sprintf("    ✓ Filtered to %d validated reproductions\n", nrow(reproductions)))

# Combine both datasets (handle case where one might be empty)
# Add source variable to track data origin
if (nrow(reproductions) == 0) {
  flora_raw <- replications %>% mutate(type = "replication", source = "replications")
} else if (nrow(replications) == 0) {
  flora_raw <- reproductions %>% mutate(type = "reproduction", source = "reproductions")
} else {
  flora_raw <- bind_rows(
    replications %>% mutate(type = "replication", source = "replications"),
    reproductions %>% mutate(type = "reproduction", source = "reproductions")
  )
}
```

### Add COS source data

```{r}
replications_cos <- read_csv(FReD_FOUNDATION_SUCCESS_CODING, show_col_types = FALSE, skip = 1)

cat(sprintf("    ✓ Downloaded %d COS rows\n", nrow(replications_cos)))

# Filter for validated entries only
replications_cos <- replications_cos %>%
  filter(validation == "validated - chosen")
cat(sprintf("    ✓ Filtered to %d validated COS entries\n", nrow(replications_cos)))

replications_cos <- replications_cos %>%
  mutate(repl_id = coalesce(url_r, doi_r)) %>%
  group_by(doi_o, repl_id) %>%
  summarise(
    ref_o = first(ref_o),
    ref_r = first(ref_r),
    url_r = first(url_r),
    doi_r = first(doi_r),
    study_o = paste(unique(na.omit(study_o)), collapse = "; "),
    abstract_r = first(abstract_r),
    outcome = if (length(unique(outcome)) == 1) unique(outcome) else "mixed",
    outcome_quote = paste(unique(na.omit(outcome_quote)), collapse = " || "),
    outcome_quote_source = paste(unique(na.omit(out_quote_source)), collapse = " || "),
    type = "replication",
    source = "COS",
    .groups = "drop"
  )

cat(sprintf("    ✓ Processed %d COS replication rows × %d columns\n", nrow(replications_cos), ncol(replications_cos)))

flora_raw <- bind_rows(
  flora_raw,
  replications_cos
)

```

### Add SCORE data (Optional)

```{r score-integration}
cat("\n### SCORE Integration\n")

if (SCORE_INTEGRATION_ENABLED) {
  cat("Status: ENABLED\n")

  if (file.exists(SCORE_DATA_PATH)) {
    score_data <- readxl::read_excel(SCORE_DATA_PATH)
    cat(sprintf("✓ Loaded SCORE data: %d rows\n", nrow(score_data)))

    # Prepare SCORE data to match FLoRA format
    # Select and rename columns to match flora_raw structure
    score_prepared <- score_data %>%
      mutate(
        type = "replication",
        source = "SCORE"
      )

    # Merge using common columns
    common_cols <- intersect(names(flora_raw), names(score_prepared))
    cat(sprintf("  Common columns: %d\n", length(common_cols)))

    flora_raw <- bind_rows(
      flora_raw,
      score_prepared %>% select(all_of(common_cols))
    )

    cat(sprintf("✓ Merged with FLoRA data\n"))
    cat(sprintf("  Total rows after SCORE merge: %d\n", nrow(flora_raw)))
  } else {
    warning(sprintf("SCORE Integration enabled but file not found: %s\n", SCORE_DATA_PATH))
  }
} else {
  cat("Status: DISABLED (use params: include_score: true to enable)\n")
}
```

### Apply Exclusions

```{r exclusions}
# Download exclusions list
exclusions <- tryCatch({
  read_csv(EXCLUSIONS_GSHEET_URL, show_col_types = FALSE)
}, error = function(e) {
  warning("Failed to download exclusions. Proceeding without exclusions.\n", e$message)
  tibble(doi_r = character(), url_r = character(), reason = character())
})

cat(sprintf("    ✓ Downloaded %d exclusions\n", nrow(exclusions)))

# Apply exclusions based on doi_r or url_r
pre_exclusion <- nrow(flora_raw)

flora_raw <- flora_raw %>%
  filter(
    !(doi_r %in% exclusions$doi_r[!is.na(exclusions$doi_r)]) &
    !(url_r %in% exclusions$url_r[!is.na(exclusions$url_r)])
  )

post_exclusion <- nrow(flora_raw)
cat(sprintf("    ✓ Excluded %d rows based on exclusion list\n", pre_exclusion - post_exclusion))
```

## Data Preparation

```{r prepare}
# Select relevant fields for FLoRA
flora_cols <- c("doi_o", "ref_o", "study_o", "doi_r", "ref_r",  "url_r",
                "abstract_r", "outcome", "outcome_quote", "outcome_quote_source", "type", "source")

flora_data <- flora_raw %>%
  select(all_of(flora_cols)) %>%
  # Normalize DOIs: lowercase and strip common prefixes

  mutate(
    doi_o = tolower(doi_o),
    doi_r = tolower(doi_r),
    # Strip "doi:" prefix and URL prefixes
    doi_o = gsub("^(https?://)?(dx\\.)?doi\\.org/", "", doi_o),
    doi_r = gsub("^(https?://)?(dx\\.)?doi\\.org/", "", doi_r),
    doi_o = gsub("^doi:", "", doi_o),
    doi_r = gsub("^doi:", "", doi_r)
  )
```

## Deduplication

Currently minimal - might want to also check blank vs filled in study_o

```{r deduplicate}

pre_dedup <- nrow(flora_data)

flora_data <- flora_data %>%
  distinct(doi_o, study_o, coalesce(url_r, doi_r), .keep_all = TRUE)

post_dedup <- nrow(flora_data)

cat(sprintf("✓ Removed %d duplicate rows\n", post_dedup - pre_dedup))
```

## Step 5: DOI Validation

Validate whether DOIs are in correct format and can be resolved

```{r validate-dois}
# Check DOI format (should start with "10.")
flora_data <- flora_data %>%
  mutate(
    doi_o_valid = is.na(doi_o) | str_starts(tolower(doi_o), "10."),
    doi_r_valid = is.na(doi_r) | str_starts(tolower(doi_r), "10.")
  )


invalid_indices <- !flora_data$doi_o_valid
invalid_o <- sum(invalid_indices)

if (invalid_o > 0) {
  # Logic: Show all if < 10, otherwise show 5
  show_all <- invalid_o < 10
  n_to_show <- if (show_all) invalid_o else 5
  
  # Select values
  vals <- head(flora_data$doi_o[invalid_indices], n_to_show)
  
  # Determine label based on whether we are showing the full set or a subset
  label <- if (show_all) "Invalid doi_o values:" else "Example invalid doi_o values:"
  
  cat(sprintf("⚠ Found %d invalid doi_o values\n", invalid_o),
      paste(label, paste(vals, collapse = ", ")), "\n")
}

invalid_indices <- !flora_data$doi_r_valid
invalid_r <- sum(invalid_indices)

if (invalid_r > 0) {
  # Logic: Show all if < 10, otherwise show 5
  show_all <- invalid_o < 10
  n_to_show <- if (show_all) invalid_r else 5
  
  # Select values
  vals <- head(flora_data$doi_r[invalid_indices], n_to_show)
  
  # Determine label based on whether we are showing the full set or a subset
  label <- if (show_all) "Invalid doi_r values:" else "Example invalid doi_r values:"
  
  cat(sprintf("⚠ Found %d invalid doi_r values\n", invalid_o),
      paste(label, paste(vals, collapse = ", ")), "\n")
}

flora_data <- flora_data %>% select(-doi_o_valid, -doi_r_valid)

#TK add DOI resolution check when available in metacheck
    
```

## Step 6a: Fetch DOI References (APA + BibTeX)

```{r fetch-references}
# Get unique DOIs
unique_dois <- unique(tolower(c(flora_data$doi_o, flora_data$doi_r)))
unique_dois <- unique_dois[!is.na(unique_dois)]

# Fetch both APA and BibTeX references with CrossRef + DataCite fallback
refs <- get_references(unique_dois)
```

<!-- Step 6b: Fetch url_r References - TODO: implement if needed
```{r fetch-url-refs, eval=FALSE}
# Placeholder for URL-keyed reference fetching
```
-->

## Step 7: Add clean References and Individual Fields

```{r clean-refs}

flora_augmented <- flora_data %>%
  # Add original references and fields
  left_join(
    refs %>% rename(
      doi_o = doi,
      ref_o_clean = reference_apa,
      bibtex_o = reference_bibtex,
      title_o = title,
      author_o = author,
      journal_o = journal,
      year_o = year,
      volume_o = volume,
      issue_o = issue,
      pages_o = pages
    ),
    by = "doi_o"
  ) %>%
  # Add replication references and fields
  left_join(
    refs %>% rename(
      doi_r = doi,
      ref_r_clean = reference_apa,
      bibtex_r = reference_bibtex,
      title_r = title,
      author_r = author,
      journal_r = journal,
      year_r = year,
      volume_r = volume,
      issue_r = issue,
      pages_r = pages
    ),
    by = "doi_r"
  ) %>%
  mutate(
    ref_o = coalesce(ref_o_clean, ref_o),
    ref_r = coalesce(ref_r_clean, ref_r)
  )

# ---- Step 7 stats helpers ----
has_text <- function(x) !is.na(x) & nzchar(trimws(as.character(x)))

count_missing <- function(x) sum(!has_text(x))
count_present <- function(x) sum(has_text(x))

cat(sprintf("✓ Added clean references (APA + BibTeX) and individual fields\n"))

# ---- BEFORE patch stats ----
cat("\n--- Step 7 stats (BEFORE manual/BibTeX recovery) ---\n")
cat(sprintf("Rows: %d\n", nrow(flora_augmented)))

cat(sprintf("APA present:    %d original, %d replication\n",
            count_present(flora_augmented$ref_o_clean),
            count_present(flora_augmented$ref_r_clean)))
cat(sprintf("APA missing:    %d original, %d replication\n",
            count_missing(flora_augmented$ref_o_clean),
            count_missing(flora_augmented$ref_r_clean)))

cat(sprintf("BibTeX present: %d original, %d replication\n",
            count_present(flora_augmented$bibtex_o),
            count_present(flora_augmented$bibtex_r)))
cat(sprintf("BibTeX missing: %d original, %d replication\n",
            count_missing(flora_augmented$bibtex_o),
            count_missing(flora_augmented$bibtex_r)))

cat(sprintf("Title present:  %d original, %d replication\n",
            count_present(flora_augmented$title_o),
            count_present(flora_augmented$title_r)))
cat(sprintf("Title missing:  %d original, %d replication\n",
            count_missing(flora_augmented$title_o),
            count_missing(flora_augmented$title_r)))

cat("  Consider adding missing refs to manual_references.xlsx\n")

# ---- Run patch ----
cat("\n### Step 7b: Manual URL overrides + BibTeX title recovery (R)\n")
before_title_r_missing <- count_missing(flora_augmented$title_r)
before_ref_r_missing   <- count_missing(flora_augmented$ref_r_clean)
before_bib_r_missing   <- count_missing(flora_augmented$bibtex_r)

flora_augmented <- augment_with_manual_url_refs_r(flora_augmented, verbose = TRUE)

# ---- AFTER patch stats ----
cat("\n--- Step 7 stats (AFTER manual/BibTeX recovery) ---\n")
cat(sprintf("Rows: %d\n", nrow(flora_augmented)))

cat(sprintf("APA present:    %d original, %d replication\n",
            count_present(flora_augmented$ref_o_clean),
            count_present(flora_augmented$ref_r_clean)))
cat(sprintf("APA missing:    %d original, %d replication\n",
            count_missing(flora_augmented$ref_o_clean),
            count_missing(flora_augmented$ref_r_clean)))

cat(sprintf("BibTeX present: %d original, %d replication\n",
            count_present(flora_augmented$bibtex_o),
            count_present(flora_augmented$bibtex_r)))
cat(sprintf("BibTeX missing: %d original, %d replication\n",
            count_missing(flora_augmented$bibtex_o),
            count_missing(flora_augmented$bibtex_r)))

cat(sprintf("Title present:  %d original, %d replication\n",
            count_present(flora_augmented$title_o),
            count_present(flora_augmented$title_r)))
cat(sprintf("Title missing:  %d original, %d replication\n",
            count_missing(flora_augmented$title_o),
            count_missing(flora_augmented$title_r)))

cat("\nRecovered on R side (replication/reproduction):\n")
cat(sprintf("  title_r recovered:  %d\n", before_title_r_missing - count_missing(flora_augmented$title_r)))
cat(sprintf("  apa ref recovered:  %d\n", before_ref_r_missing   - count_missing(flora_augmented$ref_r_clean)))
cat(sprintf("  bibtex recovered:   %d\n", before_bib_r_missing   - count_missing(flora_augmented$bibtex_r)))

```

## Step 8: Add Privacy-Preserving IDs

```{r add-ids}
# Create 3-character hash hash for privacy-preserving API lookup
flora_augmented <- flora_augmented %>%
  mutate(
    doi_o_hash = as.character(substr(openssl::md5(doi_o), 1, 3)),
    doi_r_hash = as.character(substr(openssl::md5(doi_r), 1, 3))
  )
cat(sprintf(
  "  Unique original hashes: %d\n",
  n_distinct(flora_augmented$doi_o_hash, na.rm = TRUE)
))
cat(sprintf(
  "  Unique original hashes: %d\n",
  n_distinct(flora_augmented$doi_o_hash, na.rm = TRUE)
))
cat(sprintf(
  "  Unique replication hashes: %d\n",
  n_distinct(flora_augmented$doi_r_hash, na.rm = TRUE)
))

```

## Step 9: Add Language from OpenAlex

```{r add-metadata}
cat("\n### Step 9: Adding language from OpenAlex\n")

# OpenAlex metadata functions are provided by R/openalex_cache.R (get_openalex_language)

# Get OpenAlex mailto from environment or use default
openalex_mailto <- Sys.getenv("OPENALEX_MAILTO", NA_character_)

# add language columns if not already present
if (!"language_o" %in% names(flora_augmented)) {
  flora_augmented <- flora_augmented %>%
    mutate(language_o = NA_character_)
}
if (!"language_r" %in% names(flora_augmented)) {
  flora_augmented <- flora_augmented %>%
    mutate(language_r = NA_character_)
}

# Add metadata only where fields are currently empty
dois_to_fetch_o <- flora_augmented %>%
  filter(!is.na(doi_o), is.na(language_o)) %>%
  pull(doi_o) %>%
  unique()

dois_to_fetch_r <- flora_augmented %>%
  filter(!is.na(doi_r), is.na(language_r)) %>%
  pull(doi_r) %>%
  unique()

cat(sprintf(
  "Fetching language for %d unique original DOIs and %d unique replication DOIs...\n",
  length(dois_to_fetch_o), length(dois_to_fetch_r)
))

# Fetch language for original DOIs (vectorised, cache-backed)
if (length(dois_to_fetch_o) > 0) {
  meta_o <- get_openalex_language(dois_to_fetch_o, mailto = openalex_mailto)
  meta_o <- meta_o[!is.na(meta_o$doi), ]

  flora_augmented <- flora_augmented %>%
    left_join(
      meta_o %>% rename(doi_o = doi, language_o_new = language),
      by = "doi_o"
    ) %>%
    mutate(
      language_o = coalesce(language_o, language_o_new)
    ) %>%
    select(-language_o_new)
}

# Fetch language for replication DOIs (vectorised, cache-backed)
if (length(dois_to_fetch_r) > 0) {
  meta_r <- get_openalex_language(dois_to_fetch_r, mailto = openalex_mailto)
  meta_r <- meta_r[!is.na(meta_r$doi), ]

  flora_augmented <- flora_augmented %>%
    left_join(
      meta_r %>% rename(doi_r = doi, language_r_new = language),
      by = "doi_r"
    ) %>%
    mutate(
      language_r = coalesce(language_r, language_r_new)
    ) %>%
    select(-language_r_new)
}


cat(sprintf(
  "  Original papers with language: %d\n",
  sum(!is.na(flora_augmented$language_o), na.rm = TRUE)
))
cat(sprintf(
  "  Replication papers with language: %d\n",
  sum(!is.na(flora_augmented$language_r), na.rm = TRUE)
))
```

## Step 10: Format for Output and Remove NAs

```{r format-output}
cat("\n### Step 11: Formatting for output\n")

# Select and reorder columns for output
output_cols <- c(
"doi_o", "doi_o_hash", "title_o", "author_o", "journal_o", "year_o", "volume_o", "issue_o", "pages_o", "apa_ref_o", "bibtex_ref_o", "language_o", "study_o",
"doi_r", "doi_r_hash", "title_r", "author_r", "journal_r", "year_r", "volume_r", "issue_r", "pages_r", "apa_ref_r", "bibtex_ref_r", "url_r", "language_r",
"outcome", "outcome_quote", "outcome_quote_source", "type", "source"
)

flora_renames <- c(
  c(apa_ref_o = "ref_o_clean", apa_ref_r = "ref_r_clean", bibtex_ref_o = "bibtex_o",
    bibtex_ref_r = "bibtex_r")
)

flora_augmented <- flora_augmented %>%
  rename(!!!flora_renames)

available_output <- intersect(output_cols, names(flora_augmented))
setdiff(available_output, output_cols)

flora_all <- flora_augmented %>% select(all_of(available_output))

# Log rows that will be excluded due to missing titles
excluded_rows <- flora_all %>%
  filter(is.na(title_o) | title_o == "" | is.na(title_r) | title_r == "") %>%
  mutate(
    reason = case_when(
      (is.na(title_o) | title_o == "") & (is.na(title_r) | title_r == "") ~ "missing both title_o and title_r",
      is.na(title_o) | title_o == "" ~ "missing title_o",
      is.na(title_r) | title_r == "" ~ "missing title_r"
    )
  ) %>%
  select(reason, source, doi_o, apa_ref_o, doi_r, url_r, apa_ref_r)

if (nrow(excluded_rows) > 0) {
  export_log_path <- here::here("output", "flora_export_log.csv")
  write_excel_csv(excluded_rows, export_log_path)
  cat(sprintf("⚠ %d rows excluded due to missing reference details\n", nrow(excluded_rows)))
  cat(sprintf("  Export log saved to: %s\n", export_log_path))
} else {
  cat("✓ No rows excluded due to missing reference details\n")
}

# Apply title filter
flora_final <- flora_all %>%
  filter(!is.na(title_o), title_o != "", !is.na(title_r), title_r != "")

cat(sprintf("✓ Formatted for output\n"))
cat(sprintf("  Final columns: %d\n", ncol(flora_final)))
cat(sprintf("  Final rows: %d\n", nrow(flora_final)))
```

## Step 11: Save Output

```{r save}
cat("\n### Step 12: Saving output\n")

# Create output directory if needed
if (!dir.exists(here::here("output"))) {
  dir.create(here::here("output"))
}

# Save as CSV
write_excel_csv(flora_final, FLORA_OUTPUT)

cat(sprintf("✓ Saved to: %s\n", FLORA_OUTPUT))
cat(sprintf("  Format: CSV (UTF-8)\n"))
cat(sprintf("  Rows: %d (Replications + Reproductions)\n", nrow(flora_final)))
cat(sprintf("  Columns: %d\n", ncol(flora_final)))

# Summary statistics (flora_final has renamed columns!)
has_text <- function(x) !is.na(x) & nzchar(trimws(as.character(x)))

cat("\n### Summary Statistics\n")
cat(sprintf("Total paper pairs: %d\n", nrow(flora_final)))

# by type
if ("type" %in% names(flora_final)) {
  cat("\nRows by type:\n")
  print(table(flora_final$type, useNA = "ifany"))
}

# outcomes
if ("outcome" %in% names(flora_final)) {
  cat(sprintf("Rows with outcome codes: %d\n",
              sum(has_text(flora_final$outcome))))
}

# APA refs (final names)
if ("apa_ref_o" %in% names(flora_final)) {
  cat(sprintf("Original papers with APA reference: %d\n",
              sum(has_text(flora_final$apa_ref_o))))
}
if ("apa_ref_r" %in% names(flora_final)) {
  cat(sprintf("Replication/reproduction papers with APA reference: %d\n",
              sum(has_text(flora_final$apa_ref_r))))
}

# BibTeX refs (final names)
if ("bibtex_ref_o" %in% names(flora_final)) {
  cat(sprintf("Original papers with BibTeX: %d\n",
              sum(has_text(flora_final$bibtex_ref_o))))
}
if ("bibtex_ref_r" %in% names(flora_final)) {
  cat(sprintf("Replication/reproduction papers with BibTeX: %d\n",
              sum(has_text(flora_final$bibtex_ref_r))))
}

# Language
if ("language_o" %in% names(flora_final)) {
  cat(sprintf("Original papers with language metadata: %d\n",
              sum(has_text(flora_final$language_o))))
}
if ("language_r" %in% names(flora_final)) {
  cat(sprintf("Replication/reproduction papers with language metadata: %d\n",
              sum(has_text(flora_final$language_r))))
}

# DOIs
if ("doi_o" %in% names(flora_final)) {
  cat(sprintf("Rows with original DOI: %d\n",
              sum(has_text(flora_final$doi_o))))
}
if ("doi_r" %in% names(flora_final)) {
  cat(sprintf("Rows with replication/reproduction DOI: %d\n",
              sum(has_text(flora_final$doi_r))))
}
```

## Pipeline Complete

```{r complete}
cat("\n=== FLoRA Preparation Pipeline Complete ===\n")
cat("(Includes both Replications and Reproductions)\n")
cat(sprintf("Completion time: %s\n", format(Sys.time(), "%Y-%m-%d %H:%M:%S")))
cat(sprintf("Output file: %s\n", FLORA_OUTPUT))
cat("\nNext steps:\n")
cat("1. Review the generated flora.csv file\n")
cat("2. Optionally: pipelines/flora/release_flora.qmd to release to OSF\n")
```
