---
title: "Author retrieval and overlap detection"
format: html
---

## Retrieving author data from CrossRef

This code retrieves author information from CrossRef for original and replication studies, then checks for author overlap between them.

**Note:** This script should be run periodically (similar to `crossref_ref_retrieval.qmd`) to update the author data for new entries. The results are stored in `crossref_retrieved_authors.xlsx` and `crossref_author_overlap.xlsx`, which are then loaded by the COS report to add the `author_overlap` variable to the dataset.

```{r}
library(tidyverse)
library(rcrossref)

fred_validated_raw <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vR_62H8liC0P3-WQ1xx4UalDnvqzPfObWBFwQAgVH6JhF3dPBFXd7AgFK_OljxFWQkNtUYdnUbuFTpe/pub?output=csv", skip = 1)

source("https://raw.githubusercontent.com/forrtproject/FReD-data/refs/heads/main/clean_cos_fred.R")

# No filter for validated so that we get as many authors as possible
clean_res <- clean_fred_data(fred_validated_raw)$cleaned_data

# Get unique combinations of original and replication DOIs
study_pairs <- clean_res %>%
  filter(!is.na(doi_o) & !is.na(doi_r)) %>%
  distinct(doi_o, doi_r) %>%
  mutate(doi_o = tolower(doi_o),
         doi_r = tolower(doi_r))

# Collect all unique DOIs
all_dois <- c(study_pairs$doi_o, study_pairs$doi_r) %>%
  unique() %>%
  str_subset("^10\\.")

```

## Retrieve author data from CrossRef

```{r}
# Check if we have an existing author file
authors_df <- if (file.exists("crossref_retrieved_authors.xlsx")) {
  df <- readxl::read_excel("crossref_retrieved_authors.xlsx")
  # Ensure columns are character type (even if file is empty)
  df %>%
    mutate(across(everything(), as.character)) %>%
    filter(!is.na(doi) | !is.na(authors))
} else {
  tibble(doi = character(), authors = character())
}

# Find missing DOIs
missing_dois <- setdiff(tolower(all_dois), tolower(authors_df$doi))

if (length(missing_dois) > 0) {
  message(sprintf("Retrieving author data for %d DOIs", length(missing_dois)))
  
  # Retrieve data from CrossRef in batches
  authors_new <- tibble()
  
  for (i in seq(1, length(missing_dois), by = 100)) {
    dois_batch <- missing_dois[i:min(i+99, length(missing_dois))]
    names(dois_batch) <- rep("doi", length(dois_batch))
    
    message(sprintf("  Batch %d: %d DOIs", ceiling(i/100), length(dois_batch)))
    
    res <- cr_works(filter = dois_batch, limit = 100)
    
    if (!is.null(res$data) && nrow(res$data) > 0) {
      batch_authors <- res$data %>%
        select(doi, author) %>%
        mutate(doi = tolower(doi)) %>%
        filter(!is.na(author))
      
      authors_new <- bind_rows(authors_new, batch_authors)
    }
  }
  
  if(nrow(authors_new) == 0) {
    message("No author data retrieved for missing DOIs.")
  } else {
  
  # Convert author lists to JSON strings for storage
  authors_new <- authors_new %>%
    mutate(
      authors = sapply(author, function(x) {
        tryCatch(
          jsonlite::toJSON(x, auto_unbox = TRUE),
          error = function(e) NA_character_
        )
      }, USE.NAMES = FALSE)
    ) %>%
    select(doi, authors) %>%
    mutate(across(everything(), as.character))

  # Combine with existing data
  authors_df <- authors_df %>%
    mutate(across(everything(), as.character)) %>%
    bind_rows(authors_new) %>%
    distinct(doi, .keep_all = TRUE)
  
  message(sprintf("Retrieved author data for %d new DOIs", nrow(authors_new)))
  }
} else {
  message("No new DOIs to retrieve")
}

```

## Compute author overlap

```{r}
# Function to normalize names to ASCII (remove diacritics)
normalize_to_ascii <- function(text) {
  # Convert to NFD (decomposed form), remove combining marks, convert back to ASCII
  normalized <- iconv(text, from = "UTF-8", to = "ASCII//TRANSLIT")
  # Handle cases where transliteration results in special characters
  normalized <- gsub("[^a-zA-Z0-9\\s]", "", normalized)
  return(normalized)
}

# Function to extract author identifiers from JSON
# Returns a dataframe with family name, first initial (and full given name for debugging)
# Handles character encoding inconsistencies and requires both family name and first initial for matching
extract_author_identifiers <- function(author_json) {
  # Extract author identifiers from CrossRef author JSON data
  # CrossRef consistently returns dataframes with "family", "given", "sequence" columns
  if (is.na(author_json) || author_json == "" || is.null(author_json)) {
    return(data.frame(family = character(), first_initial = character(), stringsAsFactors = FALSE))
  }

  tryCatch({
    authors <- jsonlite::fromJSON(author_json)

    # Handle empty author lists (e.g., from OSF preprints)
    if (is.null(authors) || nrow(authors) == 0) {
      return(data.frame(family = character(), first_initial = character(), stringsAsFactors = FALSE))
    }

    # Extract family names and given names
    if ("family" %in% names(authors)) {
      author_identifiers <- data.frame(
        family = authors$family %>% as.character() %>% trimws(),
        given = if ("given" %in% names(authors)) {
          authors$given %>% as.character() %>% trimws()
        } else {
          rep("", length(authors$family))
        },
        stringsAsFactors = FALSE
      )

      # Normalize family names to ASCII and lowercase
      author_identifiers <- author_identifiers %>%
        mutate(
          family_normalized = tolower(normalize_to_ascii(family)),
          # Extract first initial from given name (before space or first char)
          first_initial = ifelse(
            nzchar(given),
            tolower(substr(given, 1, 1)),
            ""
          ),
          # Create combined identifier: family_given_initial
          author_id = paste0(family_normalized, "_", first_initial)
        ) %>%
        filter(nzchar(family_normalized) & nzchar(first_initial)) %>%
        distinct(author_id, .keep_all = TRUE) %>%
        select(family_normalized, first_initial, author_id)

      return(author_identifiers)
    }

    return(data.frame(family = character(), first_initial = character(), stringsAsFactors = FALSE))

  }, error = function(e) {
    return(data.frame(family = character(), first_initial = character(), stringsAsFactors = FALSE))
  })
}

# Compute author overlap for each study pair
# Now requires matching on both family name AND first initial
author_overlap <- study_pairs %>%
  left_join(authors_df %>% rename(doi_o = doi, authors_o = authors), by = "doi_o") %>%
  left_join(authors_df %>% rename(doi_r = doi, authors_r = authors), by = "doi_r") %>%
  rowwise() %>%
  mutate(
    authors_o_ids = list(extract_author_identifiers(authors_o)),
    authors_r_ids = list(extract_author_identifiers(authors_r)),
    # Flag extraction status for quality tracking
    o_extracted = nrow(authors_o_ids) > 0,
    r_extracted = nrow(authors_r_ids) > 0,
    both_extracted = o_extracted & r_extracted,
    # Check for overlap: must match on both family name AND first initial
    author_overlap = if (both_extracted) {
      length(intersect(authors_o_ids$author_id, authors_r_ids$author_id)) > 0
    } else {
      NA
    }
  ) %>%
  ungroup()

# Generate quality report
extraction_quality <- author_overlap %>%
  summarise(
    total_pairs = n(),
    both_have_data = sum(!is.na(authors_o) & !is.na(authors_r)),
    both_extracted = sum(both_extracted, na.rm = TRUE),
    extraction_rate = if (sum(!is.na(authors_o) & !is.na(authors_r)) > 0) {
      100 * sum(both_extracted, na.rm = TRUE) / sum(!is.na(authors_o) & !is.na(authors_r))
    } else {
      0
    }
  )

message(sprintf("\n=== AUTHOR EXTRACTION QUALITY ==="))
message(sprintf("Study pairs with both studies having data: %d",
                extraction_quality$both_have_data[1]))
message(sprintf("Pairs where names could be extracted from both: %d (%.1f%% of pairs with data)",
                extraction_quality$both_extracted[1],
                extraction_quality$extraction_rate[1]))

# Keep only the essential columns for final output
author_overlap <- author_overlap %>%
  select(doi_o, doi_r, author_overlap)

# Handle cases where author data is missing
author_overlap <- author_overlap %>%
  mutate(author_overlap = if_else(is.na(author_overlap), NA, author_overlap))

```

## Save results

```{r}
# Save author data
writexl::write_xlsx(authors_df, "crossref_retrieved_authors.xlsx")

# Save author overlap results
writexl::write_xlsx(author_overlap, "crossref_author_overlap.xlsx")

message(sprintf("Saved author data for %d DOIs", nrow(authors_df)))
message(sprintf("Computed author overlap for %d study pairs", nrow(author_overlap)))
message(sprintf("Study pairs with author overlap: %d (%.1f%%)", 
                sum(author_overlap$author_overlap, na.rm = TRUE),
                100 * mean(author_overlap$author_overlap, na.rm = TRUE)))
```

## Summary statistics

```{r}
# Show summary
overlap_summary <- author_overlap %>%
  summarise(
    total_pairs = n(),
    overlap_true = sum(author_overlap, na.rm = TRUE),
    overlap_false = sum(!author_overlap, na.rm = TRUE),
    overlap_na = sum(is.na(author_overlap)),
    overlap_pct = 100 * mean(author_overlap, na.rm = TRUE)
  )

print(overlap_summary)

# Show some examples of overlapping authors
examples <- study_pairs %>%
  left_join(authors_df %>% rename(doi_o = doi, authors_o = authors), by = "doi_o") %>%
  left_join(authors_df %>% rename(doi_r = doi, authors_r = authors), by = "doi_r") %>%
  left_join(author_overlap, by = c("doi_o", "doi_r")) %>%
  filter(author_overlap == TRUE) %>%
  select(doi_o, doi_r) %>%
  head(10)

if (nrow(examples) > 0) {
  message("\nExamples of study pairs with author overlap:")
  print(examples)
}
```
