---
title: "FLoRA"
format: html
---

## FLoRA for hackathon/tools

For the Zotero plugin and preprint bot, we need a list of originals and replications, with outcomes and full bibliographic information.

# --- Load replication data and keep only rows with valid DOIs ----------------
```{r setup}
# --- Load required packages ---------------------------------------------------
# dplyr/readr: data wrangling + csv IO
# rcrossref: Crossref API client
# purrr: mapping helpers (map_chr, compact)
# jsonlite: JSON encode/decode for authors
# glue/stringr: string handling and readable messages
# rlang: provides %||% (null-coalescing) via rlang
# httr: DataCite API calls (fallback)
# digest: short DOI hashes for stable IDs
library(dplyr)
library(readr)
library(rcrossref)
library(purrr)
library(jsonlite)
library(glue)
library(stringr)
library(rlang)
library(httr)
library(digest)

# --- Load replication data ---------------------------------------------------
# Pull the public Google Sheet as CSV.
# skip = 1 removes the first "header row" in the sheet.
reported_success <- read_csv(
  "https://docs.google.com/spreadsheets/d/e/2PACX-1vSlPVImL7kjzMfpfX08JtAikvEiBg8uQ3HyEJs-YvuQMUa2RMvip-4v8-DilWl2PdwOD_UTD8gJ5wKA/pub?gid=984458430&single=true&output=csv",
  skip = 1, show_col_types = FALSE
) %>%
  # Keep only rows where original DOI and replication DOI look like a DOI prefix.
  filter(str_detect(doi_o, "^10\\.[0-9]{4,9}/")) %>%
  filter(str_detect(doi_r, "^10\\.[0-9]{4,9}/"))
```

# --- Select relevant fields and deduplicate O–R pairs ------------------------
```{r}

flora <- reported_success %>%
  # keep only the columns we care about for the FLoRA table
  select(doi_o, doi_r, ref_o, ref_r, url_r, outcome) %>%
  # group rows that refer to the same original–replication combination
  group_by(doi_o, doi_r, ref_o, ref_r, url_r) %>%
  summarise(
    # if there’s only one outcome in the group, keep it;
    # if there are multiple different outcomes, label as "mixed"
    outcome = if (length(unique(outcome)) == 1) unique(outcome) else "mixed",
    .groups = "drop"
  ) %>%
  # ensure only one row per (doi_o, doi_r) pair in the final data
  distinct(doi_o, doi_r, .keep_all = TRUE)

```

```{r}
# --- Convert CrossRef / DataCite author lists into a JSON-encoded tibble ------
jsonify_authors <- function(author_list) {
  # Handle completely missing or empty author data
  if (is.null(author_list) || length(author_list) == 0) {
    return(jsonlite::toJSON(tibble(
      given = NA_character_,
      family = NA_character_,
      sequence = NA_character_,
      ORCID = NA_character_
    ), auto_unbox = TRUE, pretty = FALSE, na = "null"))
  }

  # Crossref often returns a data.frame; DataCite may return a list
  if (is.data.frame(author_list)) {
    out <- tibble(
      given    = author_list[["given"]]    %||% NA_character_,
      family   = author_list[["family"]]   %||% author_list[["name"]] %||% NA_character_,
      sequence = author_list[["sequence"]] %||% NA_character_,
      ORCID    = author_list[["ORCID"]]    %||% NA_character_
    )
    return(jsonlite::toJSON(out, auto_unbox = TRUE, pretty = FALSE, na = "null"))
  }

  # Single-author list (like list(name="World Bank"))
  if (is.list(author_list) && (("given" %in% names(author_list)) || ("name" %in% names(author_list)))) {
    out <- tibble(
      given    = author_list[["given"]]    %||% NA_character_,
      family   = author_list[["family"]]   %||% author_list[["name"]] %||% NA_character_,
      sequence = author_list[["sequence"]] %||% NA_character_,
      ORCID    = author_list[["ORCID"]]    %||% NA_character_
    )
    return(jsonlite::toJSON(out, auto_unbox = TRUE, pretty = FALSE, na = "null"))
  }

  # If we get something completely unexpected, return a neutral row instead of crashing
  message("Unknown author format - field names: ",
          paste(names(author_list), collapse = ", "),
          ". Returning empty entry.")
  out <- tibble(
    given = NA_character_,
    family = NA_character_,
    sequence = NA_character_,
    ORCID = NA_character_
  )
  jsonlite::toJSON(out, auto_unbox = TRUE, pretty = FALSE, na = "null")
}
```

```

# --- Helpers for RDS-based DOI cache -----------------------------------------
```{r}
load_doi_cache <- function(cache_file) {
  if (file.exists(cache_file)) {
    # Load the cached tibble/data.frame from RDS
    readRDS(cache_file)
  } else {
    # Empty tibble with the expected columns
    tibble::tibble(
      doi = character(), title = character(), author = character(),
      container.title = character(), published = character(),
      volume = character(), issue = character(), page = character(),
      type = character(), source = character()
    )
  }
}

save_doi_cache <- function(cache_df, cache_file) {
  saveRDS(cache_df, cache_file)
}




# --- DOI metadata fetcher (Crossref-first + DataCite fallback + cache) -------
```{r}
get_doi_data_unified_cached <- function(
  dois,
  cache_file = "flora_doi_cache.rds",
  save_every = 100,
  force_refresh = FALSE
) {
  # --- 1) Clean DOIs ---------------------------------------------------------
  clean_doi <- function(x) {
    x <- gsub("https?://(dx\\.)?doi\\.org/", "", x, ignore.case = TRUE)
    x <- URLdecode(x)
    x <- gsub("/%?2[fF]", "/", x, perl = TRUE)  # fix 2F artifacts
    x <- gsub("//+", "/", x)                    # collapse double slashes
    trimws(tolower(x))
  }

  dois <- unique(clean_doi(dois))
  dois <- stringr::str_subset(dois, "^10\\.\\d+/\\S+$")
  if (length(dois) == 0) return(tibble::tibble())

  # --- 2) Load cache from RDS ------------------------------------------------
  cache <- load_doi_cache(cache_file)
  cache[] <- lapply(cache, as.character)  # keep everything as character
  cached_dois <- unique(cache$doi)

  # Refetch if cached 'published' is missing/blank
  needs_refresh <- cache %>%
    dplyr::filter(
      doi %in% dois,
      is.na(published) | published %in% c("", "NA", "null")
    ) %>%
    dplyr::pull(doi) %>%
    unique()

  new_dois <- if (force_refresh) {
    dois
  } else {
    unique(c(setdiff(dois, cached_dois), needs_refresh))
  }

  message(glue::glue(
    "Total DOIs: {length(dois)} | Cached: {length(cached_dois)} | To fetch: {length(new_dois)}"
  ))

  # --- 3) Year extractor -----------------------------------------------------
  get_year_safe <- function(d) {
    extract_year <- function(x) {
      if (is.null(x)) return(NA_integer_)
      if (is.list(x) && length(x) == 1 && is.list(x[[1]]) && !is.null(names(x[[1]]))) {
        x <- x[[1]]
      }
      if (is.list(x) && "date-parts" %in% names(x)) {
        dp <- x[["date-parts"]]
        if (is.list(dp) && length(dp) >= 1 && length(dp[[1]]) >= 1) {
          return(as.integer(dp[[1]][1]))
        }
      }
      suppressWarnings(as.integer(stringr::str_sub(as.character(x), 1, 4)))
    }

    candidates <- c(
      "issued", "published.print", "published.online", "published",
      "created", "deposited", "indexed"
    )

    for (nm in candidates) {
      if (nm %in% names(d)) {
        val <- tryCatch(d[[nm]][[1]], error = function(e) d[[nm]][1])
        yr  <- tryCatch(extract_year(val), error = function(e) NA_integer_)
        if (!is.na(yr)) return(yr)
      }
    }
    NA_integer_
  }

  # --- 4) Crossref fetch -----------------------------------------------------
  try_crossref <- function(doi) {
    res <- tryCatch(rcrossref::cr_works(dois = doi), error = function(e) NULL)
    if (is.null(res) || nrow(res$data) == 0) return(NULL)

    d  <- res$data
    yr <- get_year_safe(d)

    out <- tibble::tibble(
      doi = tolower(doi),
      title = as.character(d$title[1]),
      author = jsonlite::toJSON(d$author[[1]] %||% list(), auto_unbox = TRUE),
      container.title = as.character(d$container.title[1] %||% NA_character_),
      published = as.character(yr),
      volume = as.character(d$volume[1] %||% NA_character_),
      issue = as.character(d$issue[1] %||% NA_character_),
      page = as.character(d$page[1] %||% NA_character_),
      type = as.character(d$type[1] %||% NA_character_),
      source = "crossref"
    )
    out[] <- lapply(out, as.character)
    out
  }

  # --- 5) DataCite fetch (fallback) ------------------------------------------
  try_datacite <- function(doi) {
    tryCatch({
      url <- glue::glue("https://api.datacite.org/dois/{URLencode(doi)}")
      res <- httr::GET(url)
      if (res$status_code != 200) return(NULL)

      j <- jsonlite::fromJSON(httr::content(res, "text", encoding = "UTF-8"))
      attr <- j$data$attributes

      out <- tibble::tibble(
        doi = tolower(doi),
        title = as.character(attr$title[1] %||% NA_character_),
        author = jsonlite::toJSON(attr$creators %||% list(), auto_unbox = TRUE),
        container.title = as.character(attr$publisher %||% NA_character_),
        published = as.character(attr$publicationYear %||% NA_character_),
        volume = NA_character_,
        issue = NA_character_,
        page = NA_character_,
        type = as.character(attr$type[1] %||% NA_character_),
        source = "datacite"
      )
      out[] <- lapply(out, as.character)
      out
    }, error = function(e) NULL)
  }

  # --- 6) Fetch single DOI ---------------------------------------------------
  fetch_one_doi <- function(doi) {
    if (!force_refresh && doi %in% cached_dois && !doi %in% needs_refresh) {
      return(NULL)
    }
    out <- try_crossref(doi)
    if (is.null(out) || nrow(out) == 0) out <- try_datacite(doi)
    if (is.null(out) || nrow(out) == 0) {
      message(glue::glue("⚠️ No metadata found for {doi}"))
      return(NULL)
    }
    out[] <- lapply(out, as.character)
    out
  }

  # --- 7) Fetch loop + periodic saves ----------------------------------------
  new_records <- vector("list", length(new_dois))

  if (length(new_dois) > 0) {
    for (i in seq_along(new_dois)) {
      d <- new_dois[i]
      new_records[[i]] <- fetch_one_doi(d)

      if (i %% save_every == 0) {
        temp <- dplyr::bind_rows(cache, dplyr::bind_rows(purrr::compact(new_records)))
        save_doi_cache(temp, cache_file)
        message(glue::glue("💾 Saved intermediate cache after {i} fetched DOIs"))
      }
    }
  }

  valid_new <- purrr::compact(new_records)
  new_tbl <- if (length(valid_new) > 0) dplyr::bind_rows(valid_new) else tibble::tibble()

  # --- 8) Combine + dedupe + save cache --------------------------------------
  all_data <- dplyr::bind_rows(cache, new_tbl) %>%
    dplyr::mutate(dplyr::across(dplyr::everything(), as.character)) %>%
    dplyr::distinct(doi, .keep_all = TRUE)

  save_doi_cache(all_data, cache_file)
  message(glue::glue("✅ Total {nrow(all_data)} records (cached + new)"))

  all_data
}



```







































# --- Fetch + parse DOI metadata for originals and replications (from cache) --
```{r}
# --- Fetch metadata for originals and replications (CrossRef + DataCite + cache) ----
doi_data_o <- get_doi_data_unified_cached(flora$doi_o, cache_file = "flora_doi_cache.rds")
doi_data_r <- get_doi_data_unified_cached(flora$doi_r, cache_file = "flora_doi_cache.rds")


# --- Parse and rename the fields for consistency ------------------------------------
doi_data_o_parsed <- doi_data_o %>%
  mutate(
    # If author looks like JSON text, keep as-is; otherwise jsonify
    author = ifelse(
      str_detect(author, "^\\s*\\[\\s*\\{"),
      author,
      map_chr(author, jsonify_authors)
    )
  ) %>%
  transmute(
    doi       = tolower(doi),
    title_o   = title,
    author_o  = author,     # already JSON-safe now
    journal_o = container.title,
    year_o = readr::parse_integer(published, na = c("", "NA", "null")),
    volume_o  = volume,
    issue_o   = issue,
    pages_o   = page,
    type_o    = type,
    source_o  = source
  )

doi_data_r_parsed <- doi_data_r %>%
  mutate(
    author = ifelse(
      str_detect(author, "^\\s*\\[\\s*\\{"),
      author,
      map_chr(author, jsonify_authors)
    )
  ) %>%
  transmute(
    doi       = tolower(doi),
    title_r   = title,
    author_r  = author,
    journal_r = container.title,
    year_r = readr::parse_integer(published, na = c("", "NA", "null")),
    volume_r  = volume,
    issue_r   = issue,
    pages_r   = page,
    type_r    = type,
    source_r  = source
  )


```

# --- Normalise metadata for originals and replications -----------------------
```{r}

# Build a unified metadata table for ORIGINAL articles
# (common column names so later functions can treat O and R the same)
meta_o <- doi_data_o_parsed %>%
  transmute(
    doi,
    title       = title_o,
    author_json = author_o,
    journal     = journal_o,
    year        =  year_o,
    volume      = volume_o,
    issue       = issue_o,
    pages       = pages_o,
    type        = type_o
  )

# Build a unified metadata table for REPLICATION articles
meta_r <- doi_data_r_parsed %>%
  transmute(
    doi,
    title       = title_r,
    author_json = author_r,
    journal     = journal_r,
    year        =year_r,
    volume      = volume_r,
    issue       = issue_r,
    pages       = pages_r,
    type        = type_r
  )

```

# --- CrossRef citation cache + helper for APA / BibTeX lookups ----------------
```{r}
# File on disk where we cache citation strings from CrossRef
CIT_CACHE_FILE <- "crossref_citation_cache.rds"

# Load cached APA/BibTeX results if the cache file exists;
# otherwise start with an empty list structure
load_citation_cache <- function() {
  if (file.exists(CIT_CACHE_FILE)) {
    readRDS(CIT_CACHE_FILE)
  } else {
    list(apa = list(), bibtex = list())
  }
}

# Save the updated cache back to disk as an RDS file
save_citation_cache <- function(cache) {
  saveRDS(cache, CIT_CACHE_FILE)
}

# Fetch a citation for a DOI from CrossRef, either in BibTeX or APA style
get_crossref_citation <- function(doi, type = c("bibtex", "apa")) {
  type <- match.arg(type)

  if (type == "bibtex") {
    fmt   <- "bibtex"  # CrossRef content-negotiation format for BibTeX
    style <- NULL      # no citation style needed for BibTeX
  } else if (type == "apa") {
    fmt   <- "text"    # plain text output
    style <- "apa"     # APA citation style
  }

  # Try the CrossRef call; if it fails, return NA instead of throwing an error
  tryCatch(
    rcrossref::cr_cn(dois = doi, format = fmt, style = style),
    error = function(e) NA_character_
  )
}


```

# --- Resolve APA/BibTeX via CrossRef first, with manual fallback + source flag --
```{r}
override_with_crossref <- function(doi_vec, meta_df) {
  # Load (or initialize) the on-disk cache of APA / BibTeX strings
  cache <- load_citation_cache()

  # Normalise all DOIs to lowercase for consistent matching & caching
  doi_vec <- tolower(doi_vec)

  # Prepare output vectors for final citations and where they came from
  apa_final     <- vector("character", length(doi_vec))
  bibtex_final  <- vector("character", length(doi_vec))
  apa_source    <- vector("character", length(doi_vec))
  bibtex_source <- vector("character", length(doi_vec))

  # Loop over each DOI and resolve APA + BibTeX
  for (i in seq_along(doi_vec)) {
    d <- doi_vec[i]

    # Look up this DOI in the metadata table (title, authors, journal, etc.)
    meta_row <- meta_df[match(d, meta_df$doi), ]

    # --- APA ------------------------------------------------------------------
    # Try to get APA from cache first; if missing, ask CrossRef (cr_cn with style = "apa")
    apa_cr <- cache$apa[[d]] %||% get_crossref_citation(d, "apa")
    cache$apa[[d]] <- apa_cr

   if (is.null(apa_cr) || is.na(apa_cr) || !nzchar(apa_cr)) {
      # CrossRef APA is missing or empty → build our own APA-style reference

      if (nrow(meta_row) == 1 && !is.na(meta_row$type) && meta_row$type == "journal-article") {
        # For journal articles: use full APA markdown format (old helper)
        apa_fallback <- apa_reference_md(
          doi         = meta_row$doi,
          title       = meta_row$title,
          author_json = meta_row$author_json,
          journal     = meta_row$journal,
          year        = meta_row$year,
          volume      = meta_row$volume,
          issue       = meta_row$issue,
          pages       = meta_row$pages
        )[1]
      } else if (nrow(meta_row) == 1 && !is.na(meta_row$title)) {
        # For non-article types: simpler "Title. https://doi.org/..." style
        apa_fallback <- glue("{meta_row$title}. https://doi.org/{meta_row$doi}")
      } else {
        # If we don't even have a title, just fall back to the DOI URL
        apa_fallback <- glue("https://doi.org/{d}")
      }

      apa_final[i]  <- apa_fallback
      apa_source[i] <- "manual"
    } else {
      # CrossRef APA worked → use it
      apa_final[i]  <- apa_cr
      apa_source[i] <- "crossref"
    }

    # --- BibTeX ---------------------------------------------------------------
    # Same idea: use cached BibTeX, otherwise query CrossRef for BibTeX
    bib_cr <- cache$bibtex[[d]] %||% get_crossref_citation(d, "bibtex")
    cache$bibtex[[d]] <- bib_cr

    if (is.null(bib_cr) || is.na(bib_cr) || !nzchar(bib_cr)) {
      # CrossRef BibTeX missing → build a manual BibTeX record

      if (nrow(meta_row) == 1) {
        # If we have metadata, build a full @article using our helper
        bibtex_fallback <- bibtex_reference(
          doi         = meta_row$doi,
          title       = meta_row$title,
          author_json = meta_row$author_json,
          journal     = meta_row$journal,
          year        = meta_row$year,
          volume      = meta_row$volume,
          issue       = meta_row$issue,
          pages       = meta_row$pages
        )[1]
      } else {
        # Minimal @misc fallback if we have almost no metadata
        bibtex_fallback <- glue("@misc{{{d}, doi = {{{d}}}}}")
      }

      bibtex_final[i]   <- bibtex_fallback
      bibtex_source[i]  <- "manual"
    } else {
      # CrossRef BibTeX worked → use it
      bibtex_final[i]   <- bib_cr
      bibtex_source[i]  <- "crossref"
    }
  }

  # Save updated APA/BibTeX cache to disk for future runs
  save_citation_cache(cache)

  # Return a tibble with final references and where they came from
  tibble(
    doi           = doi_vec,
    apa_ref       = apa_final,
    bibtex_ref    = bibtex_final,
    apa_source    = apa_source,
    bibtex_source = bibtex_source
  )
}

```





# --- Build APA/BibTeX references for originals and replications --------------
```{r}
# Resolve APA + BibTeX for all ORIGINAL DOIs (using CrossRef-first logic)
refs_o <- override_with_crossref(
  doi_vec = meta_o$doi,
  meta_df = meta_o
)

# Resolve APA + BibTeX for all REPLICATION DOIs
refs_r <- override_with_crossref(
  doi_vec = meta_r$doi,
  meta_df = meta_r
)

```



# --- Combine metadata + references and write final FLoRA CSV ------------------

```{r}
flora %>%
  rowwise() %>%
  mutate(
    # normalise DOIs and create short hashes for compact IDs
    doi_o     = tolower(doi_o),
    doi_r     = tolower(doi_r),
    doi_o_hash = str_sub(digest::digest(doi_o, "md5", serialize = FALSE), 1, 3),
    doi_r_hash = str_sub(digest::digest(doi_r, "md5", serialize = FALSE), 1, 3)
  ) %>%
  ungroup() %>%
  # attach original article metadata (title_o, journal_o, etc.)
  left_join(doi_data_o_parsed, by = c("doi_o" = "doi")) %>%
  # attach replication article metadata
  left_join(doi_data_r_parsed, by = c("doi_r" = "doi")) %>%
  # attach original APA/BibTeX references + source flags
  left_join(refs_o, by = c("doi_o" = "doi")) %>%
  rename(
    apa_ref_o       = apa_ref,
    bibtex_ref_o    = bibtex_ref,
    apa_source_o    = apa_source,
    bibtex_source_o = bibtex_source
  ) %>%
  # attach replication APA/BibTeX references + source flags
  left_join(refs_r, by = c("doi_r" = "doi")) %>%
  rename(
    apa_ref_r       = apa_ref,
    bibtex_ref_r    = bibtex_ref,
    apa_source_r    = apa_source,
    bibtex_source_r = bibtex_source
  ) %>%
  # choose the final set of columns to export
  select(
    doi_o, doi_o_hash, title_o, author_o, journal_o, year_o, volume_o, issue_o, pages_o,
    apa_ref_o, bibtex_ref_o,
    doi_r, doi_r_hash,rep_type, title_r, author_r, journal_r, year_r, volume_r, issue_r, pages_r,
    apa_ref_r, bibtex_ref_r,
    ref_o, ref_r, url_r, outcome
  ) %>%
  # keep only rows where we successfully have both original and replication titles
  filter(!is.na(title_o), !is.na(title_r)) %>%
  # write everything to the final CSV file used by tools/bots
  write_csv("flora.csv")

```


