---
title: "FLoRA"
format: html
---

## FLoRA for hackathon/tools

For the Zotero plugin and preprint bot, we need a list of originals and replications, with outcomes and full bibliographic information.

# --- Load replication data and keep only rows with valid DOIs ----------------
```{r setup}
# --- Load required packages for data + string handling ------------------------
library(dplyr)
library(readr)
library(rcrossref)
library(purrr)
library(jsonlite)
library(glue)
library(stringr)
library(rlang)

# --- Load replication data and keep only rows with valid DOIs ----------------

reported_success <- read_csv(
  "https://docs.google.com/spreadsheets/d/e/2PACX-1vSlPVImL7kjzMfpfX08JtAikvEiBg8uQ3HyEJs-YvuQMUa2RMvip-4v8-DilWl2PdwOD_UTD8gJ5wKA/pub?gid=984458430&single=true&output=csv",
  skip = 1,              # ignore first header row in the sheet
  show_col_types = FALSE # suppress column type messages
) %>%
  # keep rows where the original DOI looks like a proper DOI
  filter(str_detect(doi_o, regex("^10\\.[0-9]{4,9}/", ignore_case = TRUE))) %>%
  # and where the replication DOI also looks like a proper DOI
  filter(str_detect(doi_r, regex("^10\\.[0-9]{4,9}/", ignore_case = TRUE)))

```



# --- Select relevant fields and deduplicate O–R pairs ------------------------
```{r}

flora <- reported_success %>%
  # keep only the columns we care about for the FLoRA table
  select(doi_o, doi_r, ref_o, ref_r, url_r, outcome) %>%
  # group rows that refer to the same original–replication combination
  group_by(doi_o, doi_r, ref_o, ref_r, url_r) %>%
  summarise(
    # if there’s only one outcome in the group, keep it;
    # if there are multiple different outcomes, label as "mixed"
    outcome = if (length(unique(outcome)) == 1) unique(outcome) else "mixed",
    .groups = "drop"
  ) %>%
  # ensure only one row per (doi_o, doi_r) pair in the final data
  distinct(doi_o, doi_r, .keep_all = TRUE)

```




# --- Helpers for fetching and formatting CrossRef metadata --------------------
```{r}
get_doi_data <- function(doi) {
  # Clean and de-duplicate DOIs before querying CrossRef
  doi <- str_replace_all(doi, "https?://doi.org/", "")
  doi <- str_subset(doi, "^10\\.\\d+/\\S+$")
  doi <- unique(doi)

  message(glue("Fetching data for {length(doi)} DOIs from CrossRef..."))

  # If nothing left after cleaning, return an empty tibble
  if (length(doi) == 0) return(tibble())

  # Query CrossRef in batches of up to 100 DOIs to avoid URL length errors
  for (i in seq(1, length(doi), by = 100)) {
    doisub <- doi[i:min(i + 99, length(doi))]
    names(doisub) <- rep("doi", length(doisub))
    res <- cr_works(filter = doisub, limit = 100)
    message(glue("  Retrieved {nrow(res$data)} records (of {length(doisub)} requested)"))
    if (i == 1) out <- res$data else out <- bind_rows(out, res$data)
  }

  message(glue("Done. Retrieved {nrow(out)} records in total."))

  return(out)
}


# --- Convert CrossRef author lists into a JSON-encoded tibble -----------------

jsonify_authors <- function(author_list) {

  if (all(c("family", "given") %in% names(author_list))) {
    # Standard case: separate given + family fields
    out <- tibble(
      given    = author_list[["given"]]    %||% NA_character_,
      family   = author_list[["family"]]   %||% NA_character_,
      sequence = author_list[["sequence"]] %||% NA_character_,
      ORCID    = author_list[["ORCID"]]    %||% NA_character_
    )

  } else if ("name" %in% names(author_list)) {
    # Some records use a single "name" field instead of "family"
    out <- tibble(
      given    = author_list[["given"]] %||% NA_character_,
      family   = author_list[["name"]]  %||% NA_character_,
      sequence = author_list[["sequence"]] %||% NA_character_,
      ORCID    = author_list[["ORCID"]]    %||% NA_character_
    )

  } else {
    # Fallback for unknown formats
    message("Unknown author format - field names: ",
            paste(names(author_list), collapse = ", "),
            ". Returning NULL entry.")
    out <- tibble(
      given    = author_list[["given"]] %||% NA_character_,
      family   = author_list[["name"]]  %||% NA_character_,
      sequence = author_list[["sequence"]] %||% NA_character_,
      ORCID    = author_list[["ORCID"]]    %||% NA_character_
    )
  }

  # Encode the author tibble as compact JSON string
  out |>
    toJSON(pretty = FALSE, na = "null")
}


# --- Extract a publication year from CrossRef date fields ---------------------

safe_year <- function(df) {
  # Candidate date columns in CrossRef records
  pub_cols <- c("published.print", "published.online", "published")
  existing <- pub_cols[pub_cols %in% names(df)]

  # If there are no date fields, return NA
  if (length(existing) == 0) {
    return(NA_integer_)
  }

  # Take the first non-missing date and keep the first 4 chars as the year
  year_str <- dplyr::coalesce(!!!df[existing])
  as.integer(stringr::str_sub(as.character(year_str), 1, 4))
}


# --- Build a markdown-formatted APA-style reference (manual fallback) ---------

apa_reference_md <- function(doi, title, author_json, journal, year, volume, issue, pages) {
  n <- length(doi)

  map_chr(seq_len(n), function(i) {
    # Parse JSON authors back into a data frame
    authors <- tryCatch(fromJSON(author_json[i]), error = function(e) NULL)

    # Format a single author as "Family, G. G."
    fmt_author <- function(given, family) {
      initials <- if (!is.na(given) && nzchar(given)) {
        parts <- strsplit(given, "\\s+")[[1]]
        paste0(substr(parts, 1, 1), collapse = ". ")
      } else ""
      paste0(family, ", ", initials, ".")
    }

    # Build the full author string depending on number of authors
    if (is.null(authors) || nrow(authors) == 0) {
      author_str <- "(n.a.)"
    } else {
      author_strs <- pmap_chr(authors, \(given, family, sequence, ORCID)
        fmt_author(given, family))
      n_auth <- length(author_strs)
      author_str <-
        if (n_auth == 1) author_strs
        else if (n_auth == 2) paste(author_strs, collapse = " & ")
        else paste0(paste(author_strs[-n_auth], collapse = ", "), ", & ", author_strs[n_auth])
    }

    # Final APA-like reference with journal info and clickable DOI
    glue(
      "{author_str} ({year[i]}). {title[i]}. *{journal[i]}*, *{volume[i]}*({issue[i]}), {pages[i]}. [https://doi.org/{doi[i]}](https://doi.org/{doi[i]})"
    )
  })
}


# --- Build a BibTeX @article entry from metadata (manual fallback) ------------

bibtex_reference <- function(doi, title, author_json, journal, year, volume, issue, pages) {
  n <- length(doi)

  map_chr(seq_len(n), function(i) {
    # Parse JSON authors back into a data frame
    authors <- tryCatch(fromJSON(author_json[i]), error = function(e) NULL)

    # Format a single author as "Family, Given"
    fmt_author_bib <- function(given, family) {
      if (!is.na(given) && nzchar(given)) {
        paste0(family, ", ", given)
      } else {
        family
      }
    }

    # Build the BibTeX author list as "Name and Name and Name"
    if (is.null(authors) || nrow(authors) == 0) {
      author_str <- "Unknown"
    } else {
      author_strs <- pmap_chr(authors, \(given, family, sequence, ORCID)
        fmt_author_bib(given, family))
      author_str <- paste(author_strs, collapse = " and ")
    }

    # Create a simple key like "smith2020"
    key <- str_remove_all(paste0(str_to_lower(word(author_strs[1], 1)), year[i]), "[^a-z0-9]")

    # Full BibTeX @article entry
    glue(
"@article{{{key},
  author = {{{author_str}}},
  title = {{{title[i]}}},
  journal = {{{journal[i]}}},
  year = {{{year[i]}}},
  volume = {{{volume[i]}}},
  number = {{{issue[i]}}},
  pages = {{{pages[i]}}},
  doi = {{{doi[i]}}}
}}"
    )
  })
}
```


# --- Fetch and parse CrossRef metadata for originals and replications ---------
```{r}
# Get raw CrossRef records for ORIGINAL DOIs
doi_data_o <- get_doi_data(flora$doi_o)

# Keep and rename only the fields we care about for originals
doi_data_o_parsed <- doi_data_o %>% 
  transmute(
    doi       = tolower(doi),                  # normalize DOI to lowercase
    title_o   = title,                         # article title
    author_o  = map_chr(author, jsonify_authors),  # authors as JSON string
    journal_o = container.title,               # journal / container title
    year_o    = safe_year(pick(everything())), # extract publication year
    volume_o  = volume,
    issue_o   = issue,
    pages_o   = page,
    type_o    = type                           # CrossRef type (e.g., "journal-article")
  )

# Get raw CrossRef records for REPLICATION DOIs
doi_data_r <- get_doi_data(flora$doi_r)

# Same parsing for replications, with "_r" suffix
doi_data_r_parsed <- doi_data_r %>% 
  transmute(
    doi       = tolower(doi),
    title_r   = title,
    author_r  = map_chr(author, jsonify_authors),
    journal_r = container.title,
    year_r    = safe_year(pick(everything())),
    volume_r  = volume,
    issue_r   = issue,
    pages_r   = page,
    type_r    = type
  )
```

# --- Normalise metadata for originals and replications -----------------------
```{r}

# Build a unified metadata table for ORIGINAL articles
# (common column names so later functions can treat O and R the same)
meta_o <- doi_data_o_parsed %>%
  transmute(
    doi,
    title       = title_o,
    author_json = author_o,
    journal     = journal_o,
    year        = year_o,
    volume      = volume_o,
    issue       = issue_o,
    pages       = pages_o,
    type        = type_o
  )

# Build a unified metadata table for REPLICATION articles
meta_r <- doi_data_r_parsed %>%
  transmute(
    doi,
    title       = title_r,
    author_json = author_r,
    journal     = journal_r,
    year        = year_r,
    volume      = volume_r,
    issue       = issue_r,
    pages       = pages_r,
    type        = type_r
  )

```

# --- CrossRef citation cache + helper for APA / BibTeX lookups ----------------
```{r}
# File on disk where we cache citation strings from CrossRef
CIT_CACHE_FILE <- "crossref_citation_cache.rds"

# Load cached APA/BibTeX results if the cache file exists;
# otherwise start with an empty list structure
load_citation_cache <- function() {
  if (file.exists(CIT_CACHE_FILE)) {
    readRDS(CIT_CACHE_FILE)
  } else {
    list(apa = list(), bibtex = list())
  }
}

# Save the updated cache back to disk as an RDS file
save_citation_cache <- function(cache) {
  saveRDS(cache, CIT_CACHE_FILE)
}

# Fetch a citation for a DOI from CrossRef, either in BibTeX or APA style
get_crossref_citation <- function(doi, type = c("bibtex", "apa")) {
  type <- match.arg(type)

  if (type == "bibtex") {
    fmt   <- "bibtex"  # CrossRef content-negotiation format for BibTeX
    style <- NULL      # no citation style needed for BibTeX
  } else if (type == "apa") {
    fmt   <- "text"    # plain text output
    style <- "apa"     # APA citation style
  }

  # Try the CrossRef call; if it fails, return NA instead of throwing an error
  tryCatch(
    rcrossref::cr_cn(dois = doi, format = fmt, style = style),
    error = function(e) NA_character_
  )
}


```

# --- Resolve APA/BibTeX via CrossRef first, with manual fallback + source flag --
```{r}
override_with_crossref <- function(doi_vec, meta_df) {
  # Load (or initialize) the on-disk cache of APA / BibTeX strings
  cache <- load_citation_cache()

  # Normalise all DOIs to lowercase for consistent matching & caching
  doi_vec <- tolower(doi_vec)

  # Prepare output vectors for final citations and where they came from
  apa_final     <- vector("character", length(doi_vec))
  bibtex_final  <- vector("character", length(doi_vec))
  apa_source    <- vector("character", length(doi_vec))
  bibtex_source <- vector("character", length(doi_vec))

  # Loop over each DOI and resolve APA + BibTeX
  for (i in seq_along(doi_vec)) {
    d <- doi_vec[i]

    # Look up this DOI in the metadata table (title, authors, journal, etc.)
    meta_row <- meta_df[match(d, meta_df$doi), ]

    # --- APA ------------------------------------------------------------------
    # Try to get APA from cache first; if missing, ask CrossRef (cr_cn with style = "apa")
    apa_cr <- cache$apa[[d]] %||% get_crossref_citation(d, "apa")
    cache$apa[[d]] <- apa_cr

    if (is.na(apa_cr) || !nzchar(apa_cr)) {
      # CrossRef APA is missing or empty → build our own APA-style reference

      if (nrow(meta_row) == 1 && !is.na(meta_row$type) && meta_row$type == "journal-article") {
        # For journal articles: use full APA markdown format (old helper)
        apa_fallback <- apa_reference_md(
          doi         = meta_row$doi,
          title       = meta_row$title,
          author_json = meta_row$author_json,
          journal     = meta_row$journal,
          year        = meta_row$year,
          volume      = meta_row$volume,
          issue       = meta_row$issue,
          pages       = meta_row$pages
        )[1]
      } else if (nrow(meta_row) == 1 && !is.na(meta_row$title)) {
        # For non-article types: simpler "Title. https://doi.org/..." style
        apa_fallback <- glue("{meta_row$title}. https://doi.org/{meta_row$doi}")
      } else {
        # If we don't even have a title, just fall back to the DOI URL
        apa_fallback <- glue("https://doi.org/{d}")
      }

      apa_final[i]  <- apa_fallback
      apa_source[i] <- "manual"
    } else {
      # CrossRef APA worked → use it
      apa_final[i]  <- apa_cr
      apa_source[i] <- "crossref"
    }

    # --- BibTeX ---------------------------------------------------------------
    # Same idea: use cached BibTeX, otherwise query CrossRef for BibTeX
    bib_cr <- cache$bibtex[[d]] %||% get_crossref_citation(d, "bibtex")
    cache$bibtex[[d]] <- bib_cr

    if (is.na(bib_cr) || !nzchar(bib_cr)) {
      # CrossRef BibTeX missing → build a manual BibTeX record

      if (nrow(meta_row) == 1) {
        # If we have metadata, build a full @article using our helper
        bibtex_fallback <- bibtex_reference(
          doi         = meta_row$doi,
          title       = meta_row$title,
          author_json = meta_row$author_json,
          journal     = meta_row$journal,
          year        = meta_row$year,
          volume      = meta_row$volume,
          issue       = meta_row$issue,
          pages       = meta_row$pages
        )[1]
      } else {
        # Minimal @misc fallback if we have almost no metadata
        bibtex_fallback <- glue("@misc{{{d}, doi = {{{d}}}}}")
      }

      bibtex_final[i]   <- bibtex_fallback
      bibtex_source[i]  <- "manual"
    } else {
      # CrossRef BibTeX worked → use it
      bibtex_final[i]   <- bib_cr
      bibtex_source[i]  <- "crossref"
    }
  }

  # Save updated APA/BibTeX cache to disk for future runs
  save_citation_cache(cache)

  # Return a tibble with final references and where they came from
  tibble(
    doi           = doi_vec,
    apa_ref       = apa_final,
    bibtex_ref    = bibtex_final,
    apa_source    = apa_source,
    bibtex_source = bibtex_source
  )
}

```





# --- Build APA/BibTeX references for originals and replications --------------
```{r}
# Resolve APA + BibTeX for all ORIGINAL DOIs (using CrossRef-first logic)
refs_o <- override_with_crossref(
  doi_vec = meta_o$doi,
  meta_df = meta_o
)

# Resolve APA + BibTeX for all REPLICATION DOIs
refs_r <- override_with_crossref(
  doi_vec = meta_r$doi,
  meta_df = meta_r
)

```



# --- Combine metadata + references and write final FLoRA CSV ------------------

```{r}
flora %>%
  rowwise() %>%
  mutate(
    # normalise DOIs and create short hashes for compact IDs
    doi_o     = tolower(doi_o),
    doi_r     = tolower(doi_r),
    doi_o_hash = str_sub(digest::digest(doi_o, "md5", serialize = FALSE), 1, 3),
    doi_r_hash = str_sub(digest::digest(doi_r, "md5", serialize = FALSE), 1, 3)
  ) %>%
  ungroup() %>%
  # attach original article metadata (title_o, journal_o, etc.)
  left_join(doi_data_o_parsed, by = c("doi_o" = "doi")) %>%
  # attach replication article metadata
  left_join(doi_data_r_parsed, by = c("doi_r" = "doi")) %>%
  # attach original APA/BibTeX references + source flags
  left_join(refs_o, by = c("doi_o" = "doi")) %>%
  rename(
    apa_ref_o       = apa_ref,
    bibtex_ref_o    = bibtex_ref,
    apa_source_o    = apa_source,
    bibtex_source_o = bibtex_source
  ) %>%
  # attach replication APA/BibTeX references + source flags
  left_join(refs_r, by = c("doi_r" = "doi")) %>%
  rename(
    apa_ref_r       = apa_ref,
    bibtex_ref_r    = bibtex_ref,
    apa_source_r    = apa_source,
    bibtex_source_r = bibtex_source
  ) %>%
  # choose the final set of columns to export
  select(
    doi_o, doi_o_hash, title_o, author_o, journal_o, year_o, volume_o, issue_o, pages_o,
    apa_ref_o, bibtex_ref_o,
    doi_r, doi_r_hash, title_r, author_r, journal_r, year_r, volume_r, issue_r, pages_r,
    apa_ref_r, bibtex_ref_r,
    ref_o, ref_r, url_r, outcome
  ) %>%
  # keep only rows where we successfully have both original and replication titles
  filter(!is.na(title_o), !is.na(title_r)) %>%
  # write everything to the final CSV file used by tools/bots
  write_csv("flora.csv")

```


