---
title: "FLoRA"
format: html
---

## Prepration of FLoRA for hackathon/tools

For the Zotero plugin and preprint bot, we need a list of originals and replications, with outcomes and full bibliographic information.

# --- Load replication data and keep only rows with valid DOIs ----------------
```{r setup}
# --- Load required packages ---------------------------------------------------
# dplyr/readr: data wrangling + csv IO
# rcrossref: Crossref API client
# purrr: mapping helpers (map_chr, compact)
# jsonlite: JSON encode/decode for authors
# glue/stringr: string handling and readable messages
# rlang: provides %||% (null-coalescing) via rlang
# httr: DataCite API calls (fallback)
# digest: short DOI hashes for stable IDs

library(dplyr)
library(readr)
library(rcrossref)
library(purrr)
library(jsonlite)
library(glue)
library(stringr)
library(rlang)
library(httr)
library(digest)
library(tidyr)

# --- Load replication data ---------------------------------------------------
# Pull the public Google Sheet as CSV.
# skip = 1 removes the first "header row" in the sheet.
reported_success <- read_csv(
  "https://docs.google.com/spreadsheets/d/e/2PACX-1vSlPVImL7kjzMfpfX08JtAikvEiBg8uQ3HyEJs-YvuQMUa2RMvip-4v8-DilWl2PdwOD_UTD8gJ5wKA/pub?gid=984458430&single=true&output=csv",
  skip = 1, show_col_types = FALSE
) %>%
  # Keep only rows where original DOI and replication DOI look like a DOI prefix.
  filter(str_detect(doi_o, "^10\\.[0-9]{4,9}/")) %>%
  filter(str_detect(doi_r, "^10\\.[0-9]{4,9}/"))
```

# --- Select relevant fields and deduplicate O–R pairs ------------------------
```{r}

flora <- reported_success %>%
  # keep only the columns we care about for the FLoRA table
  select(doi_o, doi_r, ref_o, ref_r, url_r, outcome) %>%
  mutate(
    rep_type = "Replication"   # Sheet A always has DOI replications because you filtered for it
  ) %>%
  # group rows that refer to the same original–replication combination
  group_by(doi_o, doi_r, ref_o, ref_r, url_r,rep_type) %>%
  summarise(
    # if there’s only one outcome in the group, keep it;
    # if there are multiple different outcomes, label as "mixed"
    outcome = if (length(unique(outcome)) == 1) unique(outcome) else "mixed",
    .groups = "drop"
  ) %>%
  # ensure only one row per (doi_o, doi_r) pair in the final data
  distinct(doi_o, doi_r, .keep_all = TRUE)

```

# --- Convert CrossRef / DataCite author lists into a JSON-encoded tibble ------
```{r}

jsonify_authors <- function(author_list) {
  # Handle completely missing or empty author data
  if (is.null(author_list) || length(author_list) == 0) {
    return(jsonlite::toJSON(tibble(
      given = NA_character_,
      family = NA_character_,
      sequence = NA_character_,
      ORCID = NA_character_
    ), auto_unbox = TRUE, pretty = FALSE, na = "null"))
  }

  # Crossref often returns a data.frame; DataCite may return a list
  if (is.data.frame(author_list)) {
    out <- tibble(
      given    = author_list[["given"]]    %||% NA_character_,
      family   = author_list[["family"]]   %||% author_list[["name"]] %||% NA_character_,
      sequence = author_list[["sequence"]] %||% NA_character_,
      ORCID    = author_list[["ORCID"]]    %||% NA_character_
    )
    return(jsonlite::toJSON(out, auto_unbox = TRUE, pretty = FALSE, na = "null"))
  }

  # Single-author list (like list(name="World Bank"))
  if (is.list(author_list) && (("given" %in% names(author_list)) || ("name" %in% names(author_list)))) {
    out <- tibble(
      given    = author_list[["given"]]    %||% NA_character_,
      family   = author_list[["family"]]   %||% author_list[["name"]] %||% NA_character_,
      sequence = author_list[["sequence"]] %||% NA_character_,
      ORCID    = author_list[["ORCID"]]    %||% NA_character_
    )
    return(jsonlite::toJSON(out, auto_unbox = TRUE, pretty = FALSE, na = "null"))
  }

  # If we get something completely unexpected, return a neutral row instead of crashing
  message("Unknown author format - field names: ",
          paste(names(author_list), collapse = ", "),
          ". Returning empty entry.")
  out <- tibble(
    given = NA_character_,
    family = NA_character_,
    sequence = NA_character_,
    ORCID = NA_character_
  )
  jsonlite::toJSON(out, auto_unbox = TRUE, pretty = FALSE, na = "null")
}
```

# --- Helpers for RDS-based DOI cache -----------------------------------------
```{r}
load_doi_cache <- function(cache_file) {
  if (file.exists(cache_file)) {
    # Load the cached tibble/data.frame from RDS
    readRDS(cache_file)
  } else {
    # Empty tibble with the expected columns
    tibble::tibble(
      doi = character(), title = character(), author = character(),
      container.title = character(), published = character(),
      volume = character(), issue = character(), page = character(),
      type = character(), source = character()
    )
  }
}

save_doi_cache <- function(cache_df, cache_file) {
  saveRDS(cache_df, cache_file)
}

```


# --- DOI metadata fetcher (Crossref-first + DataCite fallback + cache) -------
```{r}
get_doi_data_unified_cached <- function(
  dois,
  cache_file = "flora_doi_cache.rds",
  save_every = 100,
  force_refresh = FALSE
) {
  # --- 1) Clean DOIs ---------------------------------------------------------
  clean_doi <- function(x) {
    x <- gsub("https?://(dx\\.)?doi\\.org/", "", x, ignore.case = TRUE)
    x <- URLdecode(x)
    x <- gsub("/%?2[fF]", "/", x, perl = TRUE)  # fix 2F artifacts
    x <- gsub("//+", "/", x)                    # collapse double slashes
    trimws(tolower(x))
  }

  dois <- unique(clean_doi(dois))
  dois <- stringr::str_subset(dois, "^10\\.\\d+/\\S+$")
  if (length(dois) == 0) return(tibble::tibble())

  # --- 2) Load cache from RDS ------------------------------------------------
  cache <- load_doi_cache(cache_file)
  cache[] <- lapply(cache, as.character)  # keep everything as character
  cached_dois <- unique(cache$doi)

  # Refetch if cached 'published' is missing/blank
  needs_refresh <- cache %>%
    dplyr::filter(
      doi %in% dois,
      is.na(published) | published %in% c("", "NA", "null")
    ) %>%
    dplyr::pull(doi) %>%
    unique()

  new_dois <- if (force_refresh) {
    dois
  } else {
    unique(c(setdiff(dois, cached_dois), needs_refresh))
  }

  message(glue::glue(
    "Total DOIs: {length(dois)} | Cached: {length(cached_dois)} | To fetch: {length(new_dois)}"
  ))

  # --- 3) Year extractor -----------------------------------------------------
  get_year_safe <- function(d) {
    extract_year <- function(x) {
      if (is.null(x)) return(NA_integer_)
      if (is.list(x) && length(x) == 1 && is.list(x[[1]]) && !is.null(names(x[[1]]))) {
        x <- x[[1]]
      }
      if (is.list(x) && "date-parts" %in% names(x)) {
        dp <- x[["date-parts"]]
        if (is.list(dp) && length(dp) >= 1 && length(dp[[1]]) >= 1) {
          return(as.integer(dp[[1]][1]))
        }
      }
      suppressWarnings(as.integer(stringr::str_sub(as.character(x), 1, 4)))
    }

    candidates <- c(
      "issued", "published.print", "published.online", "published",
      "created", "deposited", "indexed"
    )

    for (nm in candidates) {
      if (nm %in% names(d)) {
        val <- tryCatch(d[[nm]][[1]], error = function(e) d[[nm]][1])
        yr  <- tryCatch(extract_year(val), error = function(e) NA_integer_)
        if (!is.na(yr)) return(yr)
      }
    }
    NA_integer_
  }

  # --- 4) Crossref fetch -----------------------------------------------------
  try_crossref <- function(doi) {
    res <- tryCatch(rcrossref::cr_works(dois = doi), error = function(e) NULL)
    if (is.null(res) || nrow(res$data) == 0) return(NULL)

    d  <- res$data
    yr <- get_year_safe(d)

    out <- tibble::tibble(
      doi = tolower(doi),
      title = as.character(d$title[1]),
      author = jsonlite::toJSON(d$author[[1]] %||% list(), auto_unbox = TRUE),
      container.title = as.character(d$container.title[1] %||% NA_character_),
      published = as.character(yr),
      volume = as.character(d$volume[1] %||% NA_character_),
      issue = as.character(d$issue[1] %||% NA_character_),
      page = as.character(d$page[1] %||% NA_character_),
      type = as.character(d$type[1] %||% NA_character_),
      source = "crossref"
    )
    out[] <- lapply(out, as.character)
    out
  }

  # --- 5) DataCite fetch (fallback) ------------------------------------------
  try_datacite <- function(doi) {
    tryCatch({
      url <- glue::glue("https://api.datacite.org/dois/{URLencode(doi)}")
      res <- httr::GET(url)
      if (res$status_code != 200) return(NULL)

      j <- jsonlite::fromJSON(httr::content(res, "text", encoding = "UTF-8"))
      attr <- j$data$attributes

      out <- tibble::tibble(
        doi = tolower(doi),
        title = as.character(attr$title[1] %||% NA_character_),
        author = jsonlite::toJSON(attr$creators %||% list(), auto_unbox = TRUE),
        container.title = as.character(attr$publisher %||% NA_character_),
        published = as.character(attr$publicationYear %||% NA_character_),
        volume = NA_character_,
        issue = NA_character_,
        page = NA_character_,
        type = as.character(attr$type[1] %||% NA_character_),
        source = "datacite"
      )
      out[] <- lapply(out, as.character)
      out
    }, error = function(e) NULL)
  }

  # --- 6) Fetch single DOI ---------------------------------------------------
  fetch_one_doi <- function(doi) {
    if (!force_refresh && doi %in% cached_dois && !doi %in% needs_refresh) {
      return(NULL)
    }
    out <- try_crossref(doi)
    if (is.null(out) || nrow(out) == 0) out <- try_datacite(doi)
    if (is.null(out) || nrow(out) == 0) {
      message(glue::glue("⚠️ No metadata found for {doi}"))
      return(NULL)
    }
    out[] <- lapply(out, as.character)
    out
  }

  # --- 7) Fetch loop + periodic saves ----------------------------------------
  new_records <- vector("list", length(new_dois))

  if (length(new_dois) > 0) {
    for (i in seq_along(new_dois)) {
      d <- new_dois[i]
      new_records[[i]] <- fetch_one_doi(d)

      if (i %% save_every == 0) {
        temp <- dplyr::bind_rows(cache, dplyr::bind_rows(purrr::compact(new_records)))
        save_doi_cache(temp, cache_file)
        message(glue::glue("💾 Saved intermediate cache after {i} fetched DOIs"))
      }
    }
  }

  valid_new <- purrr::compact(new_records)
  new_tbl <- if (length(valid_new) > 0) dplyr::bind_rows(valid_new) else tibble::tibble()

  # --- 8) Combine + dedupe + save cache --------------------------------------
  all_data <- dplyr::bind_rows(cache, new_tbl) %>%
    dplyr::mutate(dplyr::across(dplyr::everything(), as.character)) %>%
    dplyr::distinct(doi, .keep_all = TRUE)

  save_doi_cache(all_data, cache_file)
  message(glue::glue("✅ Total {nrow(all_data)} records (cached + new)"))

  all_data
}



```

# --- Fetch + parse DOI metadata for originals and replications (from cache) --
```{r}
# --- Fetch metadata for originals and replications (CrossRef + DataCite + cache) ----
doi_data_o <- get_doi_data_unified_cached(flora$doi_o, cache_file = "flora_doi_cache.rds")
doi_data_r <- get_doi_data_unified_cached(flora$doi_r, cache_file = "flora_doi_cache.rds")


# --- Parse and rename the fields for consistency ------------------------------------
doi_data_o_parsed <- doi_data_o %>%
  mutate(
    # If author looks like JSON text, keep as-is; otherwise jsonify
    author = ifelse(
      str_detect(author, "^\\s*\\[\\s*\\{"),
      author,
      map_chr(author, jsonify_authors)
    )
  ) %>%
  transmute(
    doi       = tolower(doi),
    title_o   = title,
    author_o  = author,     # already JSON-safe now
    journal_o = container.title,
    year_o = readr::parse_integer(published, na = c("", "NA", "null")),
    volume_o  = volume,
    issue_o   = issue,
    pages_o   = page,
    type_o    = type,
    source_o  = source
  )

doi_data_r_parsed <- doi_data_r %>%
  mutate(
    author = ifelse(
      str_detect(author, "^\\s*\\[\\s*\\{"),
      author,
      map_chr(author, jsonify_authors)
    )
  ) %>%
  transmute(
    doi       = tolower(doi),
    title_r   = title,
    author_r  = author,
    journal_r = container.title,
    year_r = readr::parse_integer(published, na = c("", "NA", "null")),
    volume_r  = volume,
    issue_r   = issue,
    pages_r   = page,
    type_r    = type,
    source_r  = source
  )


```


# --- Normalise metadata for originals and replications -----------------------
```{r}

# Build a unified metadata table for ORIGINAL articles
# (common column names so later functions can treat O and R the same)
meta_o <- doi_data_o_parsed %>%
  transmute(
    doi,
    title       = title_o,
    author_json = author_o,
    journal     = journal_o,
    year        =  year_o,
    volume      = volume_o,
    issue       = issue_o,
    pages       = pages_o,
    type        = type_o
  )

# Build a unified metadata table for REPLICATION articles
meta_r <- doi_data_r_parsed %>%
  transmute(
    doi,
    title       = title_r,
    author_json = author_r,
    journal     = journal_r,
    year        =year_r,
    volume      = volume_r,
    issue       = issue_r,
    pages       = pages_r,
    type        = type_r
  )

```

# --- CrossRef citation cache + helper for APA / BibTeX lookups ----------------
```{r}
# File on disk where we cache citation strings from CrossRef
CIT_CACHE_FILE <- "crossref_citation_cache.rds"

# Load cached APA/BibTeX results if the cache file exists;
# otherwise start with an empty list structure
load_citation_cache <- function() {
  if (file.exists(CIT_CACHE_FILE)) {
    readRDS(CIT_CACHE_FILE)
  } else {
    list(apa = list(), bibtex = list())
  }
}

# Save the updated cache back to disk as an RDS file
save_citation_cache <- function(cache) {
  saveRDS(cache, CIT_CACHE_FILE)
}

# Fetch a citation for a DOI from CrossRef, either in BibTeX or APA style
get_crossref_citation <- function(doi, type = c("bibtex", "apa")) {
  type <- match.arg(type)

  if (type == "bibtex") {
    fmt   <- "bibtex"  # CrossRef content-negotiation format for BibTeX
    style <- NULL      # no citation style needed for BibTeX
  } else if (type == "apa") {
    fmt   <- "text"    # plain text output
    style <- "apa"     # APA citation style
  }

  # Try the CrossRef call; if it fails, return NA instead of throwing an error
  tryCatch(
    rcrossref::cr_cn(dois = doi, format = fmt, style = style),
    error = function(e) NA_character_
  )
}


```

# --- Resolve APA/BibTeX via CrossRef first, with manual fallback + source flag --
```{r}
override_with_crossref <- function(doi_vec, meta_df) {
  # Load (or initialize) the on-disk cache of APA / BibTeX strings
  cache <- load_citation_cache()

  # Normalise all DOIs to lowercase for consistent matching & caching
  doi_vec <- tolower(doi_vec)

  # Prepare output vectors for final citations and where they came from
  apa_final     <- vector("character", length(doi_vec))
  bibtex_final  <- vector("character", length(doi_vec))
  apa_source    <- vector("character", length(doi_vec))
  bibtex_source <- vector("character", length(doi_vec))

  # Loop over each DOI and resolve APA + BibTeX
  for (i in seq_along(doi_vec)) {
    d <- doi_vec[i]

    # Look up this DOI in the metadata table (title, authors, journal, etc.)
    meta_row <- meta_df[match(d, meta_df$doi), ]

    # --- APA ------------------------------------------------------------------
    # Try to get APA from cache first; if missing, ask CrossRef (cr_cn with style = "apa")
    apa_cr <- cache$apa[[d]] %||% get_crossref_citation(d, "apa")
    cache$apa[[d]] <- apa_cr

   if (is.null(apa_cr) || is.na(apa_cr) || !nzchar(apa_cr)) {
      # CrossRef APA is missing or empty → build our own APA-style reference

      if (nrow(meta_row) == 1 && !is.na(meta_row$type) && meta_row$type == "journal-article") {
        # For journal articles: use full APA markdown format (old helper)
        apa_fallback <- apa_reference_md(
          doi         = meta_row$doi,
          title       = meta_row$title,
          author_json = meta_row$author_json,
          journal     = meta_row$journal,
          year        = meta_row$year,
          volume      = meta_row$volume,
          issue       = meta_row$issue,
          pages       = meta_row$pages
        )[1]
      } else if (nrow(meta_row) == 1 && !is.na(meta_row$title)) {
        # For non-article types: simpler "Title. https://doi.org/..." style
        apa_fallback <- glue("{meta_row$title}. https://doi.org/{meta_row$doi}")
      } else {
        # If we don't even have a title, just fall back to the DOI URL
        apa_fallback <- glue("https://doi.org/{d}")
      }

      apa_final[i]  <- apa_fallback
      apa_source[i] <- "manual"
    } else {
      # CrossRef APA worked → use it
      apa_final[i]  <- apa_cr
      apa_source[i] <- "crossref"
    }

    # --- BibTeX ---------------------------------------------------------------
    # Same idea: use cached BibTeX, otherwise query CrossRef for BibTeX
    bib_cr <- cache$bibtex[[d]] %||% get_crossref_citation(d, "bibtex")
    cache$bibtex[[d]] <- bib_cr

    if (is.null(bib_cr) || is.na(bib_cr) || !nzchar(bib_cr)) {
      # CrossRef BibTeX missing → build a manual BibTeX record

      if (nrow(meta_row) == 1) {
        # If we have metadata, build a full @article using our helper
        bibtex_fallback <- bibtex_reference(
          doi         = meta_row$doi,
          title       = meta_row$title,
          author_json = meta_row$author_json,
          journal     = meta_row$journal,
          year        = meta_row$year,
          volume      = meta_row$volume,
          issue       = meta_row$issue,
          pages       = meta_row$pages
        )[1]
      } else {
        # Minimal @misc fallback if we have almost no metadata
        bibtex_fallback <- glue("@misc{{{d}, doi = {{{d}}}}}")
      }

      bibtex_final[i]   <- bibtex_fallback
      bibtex_source[i]  <- "manual"
    } else {
      # CrossRef BibTeX worked → use it
      bibtex_final[i]   <- bib_cr
      bibtex_source[i]  <- "crossref"
    }
  }

  # Save updated APA/BibTeX cache to disk for future runs
  save_citation_cache(cache)

  # Return a tibble with final references and where they came from
  tibble(
    doi           = doi_vec,
    apa_ref       = apa_final,
    bibtex_ref    = bibtex_final,
    apa_source    = apa_source,
    bibtex_source = bibtex_source
  )
}

```

# --- Build APA/BibTeX references for originals and replications --------------
```{r}
# Resolve APA + BibTeX for all ORIGINAL DOIs (using CrossRef-first logic)
refs_o <- override_with_crossref(
  doi_vec = meta_o$doi,
  meta_df = meta_o
)

# Resolve APA + BibTeX for all REPLICATION DOIs
refs_r <- override_with_crossref(
  doi_vec = meta_r$doi,
  meta_df = meta_r
)

```





# --- Combine metadata + references and write final FLoRA CSV ------------------
```{r}
flora %>%
  rowwise() %>%
  mutate(
    # normalise DOIs and create short hashes for compact IDs
    doi_o     = tolower(doi_o),
    doi_r     = tolower(doi_r),
    doi_o_hash = str_sub(digest::digest(doi_o, "md5", serialize = FALSE), 1, 3),
    doi_r_hash = str_sub(digest::digest(doi_r, "md5", serialize = FALSE), 1, 3)
  ) %>%
  ungroup() %>%
  # attach original article metadata (title_o, journal_o, etc.)
  left_join(doi_data_o_parsed, by = c("doi_o" = "doi")) %>%
  # attach replication article metadata
  left_join(doi_data_r_parsed, by = c("doi_r" = "doi")) %>%
  # attach original APA/BibTeX references + source flags
  left_join(refs_o, by = c("doi_o" = "doi")) %>%
  rename(
    apa_ref_o       = apa_ref,
    bibtex_ref_o    = bibtex_ref,
    apa_source_o    = apa_source,
    bibtex_source_o = bibtex_source
  ) %>%
  # attach replication APA/BibTeX references + source flags
  left_join(refs_r, by = c("doi_r" = "doi")) %>%
  rename(
    apa_ref_r       = apa_ref,
    bibtex_ref_r    = bibtex_ref,
    apa_source_r    = apa_source,
    bibtex_source_r = bibtex_source
  ) %>%
  # choose the final set of columns to export
  select(
    doi_o, doi_o_hash, title_o, author_o, journal_o, year_o, volume_o, issue_o, pages_o,
    apa_ref_o, bibtex_ref_o,
    doi_r, doi_r_hash, rep_type, title_r, author_r, journal_r, year_r, volume_r, issue_r, pages_r,
    apa_ref_r, bibtex_ref_r,
    ref_o, ref_r, url_r, outcome
  ) %>%
  # keep only rows where we successfully have both original and replication titles
  filter(!is.na(title_o), !is.na(title_r)) %>%
  # write everything to the final CSV file used by tools/bots
  identity() -> flora_a_out

```


# Sheet B Part 1 : (Helpers: convert JSON author metadata into APA/BibTeX strings + build manualAPA/BibTeX references (used when Crossref citations are missing) )
```{r}
# ---- Safely parse JSON without crashing -------------------------------------
safe_from_json <- function(x) {
  # Try to parse JSON text `x` into an R object (list/data.frame).
  # If parsing fails, return NULL instead of throwing an error.
  tryCatch(jsonlite::fromJSON(x), error = function(e) NULL)
}

# ---- Turn a given-name string into initials (APA style) ----------------------
given_to_initials <- function(given) {
  # If the given name is missing/empty, return empty string.
  if (is.null(given) || is.na(given) || !nzchar(trimws(given))) return("")

  # Split given name by whitespace into parts: "John Ronald" -> c("John","Ronald")
  parts <- unlist(strsplit(trimws(given), "\\s+"))

  # Take first letter of each part and add "." then collapse with spaces: "J. R."
  paste0(substr(parts, 1, 1), ".", collapse = " ")
}

# ---- Convert author JSON -> APA author list string --------------------------
# Expected author_json format (Crossref-style): JSON array of objects with `given` and `family`
# Output example: "Lacko, D., & Prošek, T."
authors_to_apa <- function(author_json) {
  # If missing/empty, return NA
  if (is.null(author_json) || is.na(author_json) || !nzchar(trimws(author_json))) return(NA_character_)

  # Parse JSON text to R object; if it isn't JSON, treat it as already-human-readable text
  a <- safe_from_json(author_json)
  if (is.null(a)) return(trimws(author_json))

  # If JSON parsed as a list (not data.frame), try converting to data.frame
  if (is.list(a) && !is.data.frame(a)) {
    a <- tryCatch(as.data.frame(a, stringsAsFactors = FALSE), error = function(e) NULL)
  }

  # If still unusable (NULL or no rows), return NA
  if (is.null(a) || nrow(a) == 0) return(NA_character_)

  # Ensure required columns exist (some sources might omit them)
  if (!("family" %in% names(a))) a$family <- NA_character_
  if (!("given"  %in% names(a))) a$given  <- NA_character_

  # Number of authors in the parsed list
  n <- nrow(a)

  # APA rule: if >20 authors, show first 19, ellipsis, last author.
  idx <- if (n <= 20) seq_len(n) else c(1:19, n)

  # Build each author string as "Family, I." (initials from given names)
  parts <- purrr::map_chr(idx, function(i) {
    # Extract family and given names safely, coercing to character and trimming whitespace
    fam <- trimws(as.character(a$family[i] %||% ""))
    giv <- trimws(as.character(a$given[i]  %||% ""))

    # Convert given name to initials: "David Paul" -> "D. P."
    ini <- trimws(given_to_initials(giv))

    # Prefer "Family, Initials"; else fall back to whichever part exists
    if (nzchar(fam) && nzchar(ini)) glue::glue("{fam}, {ini}")
    else if (nzchar(fam)) fam
    else if (nzchar(giv)) giv
    else NA_character_
  })

  # Drop empty/NA author entries
  parts <- parts[!is.na(parts) & nzchar(parts)]
  if (length(parts) == 0) return(NA_character_)

  # If there were >20 authors, insert "..." before the last author name
  if (n > 20) parts <- c(parts[1:19], "...", parts[length(parts)])

  # If only one author, return it directly
  if (length(parts) == 1) return(parts)

  # Join multiple authors with commas and an "&" before the last one (APA)
  paste0(
    paste(parts[1:(length(parts) - 1)], collapse = ", "),
    ", & ",
    parts[length(parts)]
  )
}

# ---- Convert author JSON -> BibTeX author field string -----------------------
# Output example: "Lacko, David and Prošek, Tomas"
authors_to_bibtex <- function(author_json) {
  # If missing/empty, return NA
  if (is.null(author_json) || is.na(author_json) || !nzchar(trimws(author_json))) return(NA_character_)

  # Parse JSON; if not JSON, return raw text
  a <- safe_from_json(author_json)
  if (is.null(a)) return(trimws(author_json))

  # Convert list to data.frame if needed
  if (is.list(a) && !is.data.frame(a)) {
    a <- tryCatch(as.data.frame(a, stringsAsFactors = FALSE), error = function(e) NULL)
  }

  # If unusable, return NA
  if (is.null(a) || nrow(a) == 0) return(NA_character_)

  # Ensure columns exist
  if (!("family" %in% names(a))) a$family <- NA_character_
  if (!("given"  %in% names(a))) a$given  <- NA_character_

  # Build each author as "Family, Given" (BibTeX accepts this well)
  parts <- purrr::map_chr(seq_len(nrow(a)), function(i) {
    fam <- trimws(as.character(a$family[i] %||% ""))
    giv <- trimws(as.character(a$given[i]  %||% ""))

    if (nzchar(fam) && nzchar(giv)) glue::glue("{fam}, {giv}")
    else if (nzchar(fam)) fam
    else if (nzchar(giv)) giv
    else NA_character_
  })

  # Drop empty entries; if none, return NA
  parts <- parts[!is.na(parts) & nzchar(parts)]
  if (length(parts) == 0) return(NA_character_)

  # BibTeX expects authors separated by " and "
  paste(parts, collapse = " and ")
}

# ---- Minimal "escaping"/cleanup for BibTeX fields ----------------------------
bibtex_escape <- function(x) {
  # If missing, return NA
  if (is.null(x) || is.na(x)) return(NA_character_)

  # Ensure character
  x <- as.character(x)

  # Replace newlines with spaces
  x <- gsub("[\r\n]+", " ", x)

  # Collapse repeated whitespace
  x <- gsub("\\s+", " ", x)

  # Trim leading/trailing whitespace
  trimws(x)
}

# ---- Create a stable BibTeX key --------------------------------------------
make_bibtex_key <- function(doi = NA_character_, title = NA_character_, year = NA_character_) {
  # Use DOI as the stable base if available; otherwise fall back to "title|year"
  base <- if (!is.na(doi) && nzchar(doi)) doi else paste0(title %||% "", "|", year %||% "")

  # Hash the base so keys are short but stable across runs
  h <- digest::digest(base, algo = "md5", serialize = FALSE)

  # Prefix + first 8 chars of md5 hash
  paste0("flora_", substr(h, 1, 8))
}

# ---- Build a manual APA-style reference (Markdown-friendly) -----------------
# Output example:
#   "Lacko, D., & Prošek, T. (2025). Title. *Journal*, *132*(4), 12-34. https://doi.org/..."
apa_reference_md <- function(doi, title, author_json, journal, year, volume, issue, pages) {
  # Convert authors JSON -> APA author string
  auth <- authors_to_apa(author_json)

  # Parse year to integer (if possible), suppress warnings if non-numeric
  yr   <- suppressWarnings(as.integer(year))

  # Normalize fields: replace NULL with NA (so later checks are consistent)
  title_txt   <- title %||% NA_character_
  journal_txt <- journal %||% NA_character_
  vol_txt     <- volume %||% NA_character_
  iss_txt     <- issue %||% NA_character_
  pages_txt   <- pages %||% NA_character_

  # Build DOI URL if DOI is present
  doi_url <- if (!is.null(doi) && !is.na(doi) && nzchar(doi)) glue::glue("https://doi.org/{doi}") else NA_character_

  # Start building reference parts into a vector
  out <- c()

  # Add authors
  if (!is.na(auth) && nzchar(auth)) {
    out <- c(out, auth)
  }

  # Add year in parentheses
  if (!is.na(yr)) {
    out <- c(out, glue::glue("({yr})."))
  }

  # Add title
  if (!is.na(title_txt) && nzchar(title_txt)) {
    out <- c(out, glue::glue("{title_txt}."))
  }

  # Build the journal/volume/issue/pages segment
  jpart <- ""

  # Journal in italics (Markdown)
  if (!is.na(journal_txt) && nzchar(journal_txt)) jpart <- glue::glue("*{journal_txt}*")

  # Volume in italics, appended after journal (APA style)
  if (!is.na(vol_txt) && nzchar(vol_txt)) {
    jpart <- if (nzchar(jpart)) glue::glue("{jpart}, *{vol_txt}*") else glue::glue("*{vol_txt}*")
  }

  # Issue in parentheses
  if (!is.na(iss_txt) && nzchar(iss_txt)) jpart <- glue::glue("{jpart}({iss_txt})")

  # Pages after comma
  if (!is.na(pages_txt) && nzchar(pages_txt)) jpart <- glue::glue("{jpart}, {pages_txt}")

  # Add journal block and DOI URL
  if (nzchar(jpart)) out <- c(out, glue::glue("{jpart}."))
  if (!is.na(doi_url) && nzchar(doi_url)) out <- c(out, doi_url)

  # Join non-empty pieces with spaces
  out <- paste(out[!is.na(out) & nzchar(out)], collapse = " ")

  # If everything else is empty but DOI exists, at least return DOI URL
  if (!nzchar(out) && !is.na(doi_url)) out <- doi_url

  out
}

# ---- Build a manual BibTeX entry (article if possible, else misc) ------------
bibtex_reference <- function(doi, title, author_json, journal, year, volume, issue, pages, key = NULL) {
  # Parse year to integer, suppress warnings on bad inputs
  yr <- suppressWarnings(as.integer(year))

  # Use provided key, or generate one based on DOI/title/year
  key <- key %||% make_bibtex_key(doi = doi, title = title, year = as.character(yr))

  # Convert authors JSON -> BibTeX author string ("Family, Given and ...")
  a <- authors_to_bibtex(author_json)

  # Clean fields for BibTeX output (mostly whitespace normalization)
  title_e   <- bibtex_escape(title)
  journal_e <- bibtex_escape(journal)
  pages_e   <- bibtex_escape(pages)
  volume_e  <- bibtex_escape(volume)
  issue_e   <- bibtex_escape(issue)
  doi_e     <- bibtex_escape(doi)

  # Decide whether we have enough info to call it an @article
  # (needs title, journal, and year)
  is_article <- !is.na(title_e) && nzchar(title_e) &&
    !is.na(journal_e) && nzchar(journal_e) &&
    !is.na(yr)

  # ---- Build @article if we have enough fields ------------------------------
  if (is_article) {
    # Build BibTeX fields; each field is included only if its value exists
    fields <- c(
      if (!is.na(a) && nzchar(a)) glue::glue("  author = {{{a}}}"),
      if (!is.na(title_e) && nzchar(title_e)) glue::glue("  title = {{{title_e}}}"),
      if (!is.na(journal_e) && nzchar(journal_e)) glue::glue("  journal = {{{journal_e}}}"),
      if (!is.na(yr)) glue::glue("  year = {{{yr}}}"),
      if (!is.na(volume_e) && nzchar(volume_e)) glue::glue("  volume = {{{volume_e}}}"),
      if (!is.na(issue_e) && nzchar(issue_e)) glue::glue("  number = {{{issue_e}}}"),
      if (!is.na(pages_e) && nzchar(pages_e)) glue::glue("  pages = {{{pages_e}}}"),
      if (!is.na(doi_e) && nzchar(doi_e)) glue::glue("  doi = {{{doi_e}}}"),
      if (!is.na(doi_e) && nzchar(doi_e)) glue::glue("  url = https://doi.org/{doi_e}")
    )

    # Drop any NULL/empty field lines
    fields <- fields[!is.na(fields) & nzchar(fields)]

    # Assemble the final BibTeX string for @article
    return(paste0(
      "@article{", key, ",\n",
      paste(fields, collapse = ",\n"),
      "\n}\n"
    ))
  }

  # ---- Otherwise build @misc fallback ---------------------------------------
  fields <- c(
    if (!is.na(a) && nzchar(a)) glue::glue("  author = {{{a}}}"),
    if (!is.na(title_e) && nzchar(title_e)) glue::glue("  title = {{{title_e}}}"),
    if (!is.na(yr)) glue::glue("  year = {{{yr}}}"),
    if (!is.na(doi_e) && nzchar(doi_e)) glue::glue("  doi = {{{doi_e}}}"),
    if (!is.na(doi_e) && nzchar(doi_e)) glue::glue("  url = https://doi.org/{doi_e}")
  )

  # Drop empty field lines
  fields <- fields[!is.na(fields) & nzchar(fields)]

  # Assemble the final BibTeX string for @misc
  paste0(
    "@misc{", key, ",\n",
    paste(fields, collapse = ",\n"),
    "\n}\n"
  )
}

```

# SHEET B PIPELINE PART 2 :append reproduction-list (Sheet B) to Sheet A output
```{r}
# workflow:
# 1) Assumes Sheet A pipeline already ran and produced an in-memory data frame called `flora_a_out`.
# 2) Loads Sheet B (reproduction list) from Google Sheets.
# 3) Normalizes DOIs, deduplicates, sets rep_type = "Reproduction".
# 4) Fetches metadata via Crossref/DataCite .
# 5) If Crossref metadata missing, fills from ref_o/ref_r by parsing citations.
# 6) Builds APA + BibTeX using your crossref-first override_with_crossref(),with a manual fallback (per-row) when DOI is missing or Crossref citation fails.
# 7) Aligns column types to avoid bind_rows() type conflicts.
# 8) Binds A + B and writes ONE file: flora.csv

# ---- 1) Helpers -------------------------------------------------------------

# Normalize DOI strings (strip doi.org prefix, decode, lowercase, validate)
normalize_doi <- function(x) {
  if (is.null(x)) return(NA_character_)
  x <- as.character(x)
  x <- gsub("https?://(dx\\.)?doi\\.org/", "", x, ignore.case = TRUE)
  x <- URLdecode(x)
  x <- gsub("/%?2[fF]", "/", x, perl = TRUE)
  x <- gsub("//+", "/", x)
  x <- trimws(tolower(x))
  ifelse(stringr::str_detect(x, "^10\\.[0-9]{4,9}/\\S+$"), x, NA_character_)
}

# Protect periods inside abbreviations so "first sentence" title extraction works
protect_abbrev_periods <- function(x) {
  if (is.na(x) || !nzchar(x)) return(x)
  x <- as.character(x)
  x <- gsub("\\bet\\s+al\\.", "et al§", x, ignore.case = TRUE)
  x <- gsub("\\be\\.g\\.", "e§g§", x, ignore.case = TRUE)
  x <- gsub("\\bi\\.e\\.", "i§e§", x, ignore.case = TRUE)
  x <- gsub("\\bU\\.S\\.", "U§S§", x)
  x
}

# Undo abbreviation protection
unprotect_abbrev_periods <- function(x) {
  if (is.na(x) || !nzchar(x)) return(x)
  x <- as.character(x)
  gsub("§", ".", x)
}

# Parse a reference string into best-effort components:
# author_raw, year, title, journal, volume, issue, pages
parse_ref_general <- function(ref) {
  ref <- as.character(ref)

  # Empty ref -> return NA fields
  if (is.na(ref) || !nzchar(trimws(ref))) {
    return(tibble::tibble(
      author_raw = NA_character_,
      year = NA_integer_,
      title = NA_character_,
      journal = NA_character_,
      volume = NA_character_,
      issue = NA_character_,
      pages = NA_character_
    ))
  }

  # --- APA-ish pattern: "... (2024). Title. Journal, 106(3), 638-654."
  if (stringr::str_detect(ref, "\\((19|20)\\d{2}\\)\\.")) {

    # Extract year "(2024)" then turn into integer 2024
    year_chr <- stringr::str_match(ref, "\\((19|20)\\d{2}\\)")[1]
    year_int <- suppressWarnings(as.integer(stringr::str_remove_all(year_chr, "[\\(\\)]")))

    # Split into two parts at the end of "(YYYY). "
    # parts[1] = authors (maybe includes year), parts[2] = rest (title + journal etc.)
    parts <- stringr::str_split_fixed(ref, "\\)\\.\\s+", 2)

    authors_raw <- trimws(parts[1])

    # Remove trailing "(YYYY)" if it stuck to authors_raw
    authors_raw <- stringr::str_remove(authors_raw, "\\s*\\((19|20)\\d{2}\\)\\s*$")

    rest <- parts[2]

    # Protect abbreviation periods so we don't cut title too early
    rest_p <- protect_abbrev_periods(rest)

    # Title = first sentence in rest
    title_p <- stringr::str_trim(stringr::str_extract(rest_p, "^[^.]+"))
    title   <- unprotect_abbrev_periods(title_p)

    # After-title portion = remainder after "Title. "
    after_title_p <- stringr::str_trim(stringr::str_remove(rest_p, "^[^.]+\\.\\s*"))
    after_title   <- unprotect_abbrev_periods(after_title_p)

    # Journal = text before first comma
    journal <- stringr::str_trim(stringr::str_extract(after_title, "^[^,]+"))

    # Volume = digits before "(" in ", 106(3)"
    volume  <- stringr::str_match(after_title, ",\\s*([0-9]+)\\s*\\(")[,2]

    # Issue = inside "(3)"
    issue   <- stringr::str_match(after_title, "\\(([0-9A-Za-z]+)\\)")[,2]

    # Pages = "638-654" or "638–654"
    pages   <- stringr::str_match(after_title, "([0-9]+\\s*[-–]\\s*[0-9]+)")[,2]

    return(tibble::tibble(
      author_raw = authors_raw,
      year = ifelse(is.finite(year_int), year_int, NA_integer_),
      title = title %||% NA_character_,
      journal = journal %||% NA_character_,
      volume = volume %||% NA_character_,
      issue = issue %||% NA_character_,
      pages = pages %||% NA_character_
    ))
  }

  # --- Listing style:
  # "Eliza Forsythe, Title, The Economic Journal, Volume 132, Issue 645, July 2022, Pages 1765–1789."
  author_raw  <- stringr::str_trim(stringr::str_extract(ref, "^[^,]+"))
  title_raw   <- stringr::str_trim(stringr::str_match(ref, "^[^,]+,\\s*([^,]+),")[,2])
  journal_raw <- stringr::str_trim(stringr::str_match(ref, "^[^,]+,\\s*[^,]+,\\s*([^,]+),")[,2])

  volume <- stringr::str_trim(stringr::str_match(ref, "Volume\\s+([^,]+)")[,2])
  issue  <- stringr::str_trim(stringr::str_match(ref, "Issue\\s+([^,]+)")[,2])

  year_int <- suppressWarnings(as.integer(stringr::str_match(ref, "(19|20)\\d{2}")[,1]))
  if (!is.finite(year_int)) year_int <- NA_integer_

  pages <- stringr::str_trim(stringr::str_match(ref, "Pages\\s+([0-9]+\\s*[-–]\\s*[0-9]+)")[,2])

  tibble::tibble(
    author_raw = author_raw %||% NA_character_,
    year = year_int,
    title = title_raw %||% NA_character_,
    journal = journal_raw %||% NA_character_,
    volume = volume %||% NA_character_,
    issue = issue %||% NA_character_,
    pages = pages %||% NA_character_
  )
}

# Convert author_raw (APA-ish) into JSON list of authors
# Expected input often like: "Lacko, D., & Prošek, T."
# Strategy:
# - Normalize "&" into commas
# - Split on commas
# - Interpret as pairs: family, given, family, given, ...
author_raw_to_json <- function(author_raw) {
  author_raw <- as.character(author_raw)

  if (is.na(author_raw) || !nzchar(trimws(author_raw))) {
    return(jsonlite::toJSON(tibble::tibble(
      given = NA_character_, family = NA_character_, name = NA_character_,
      sequence = NA_character_, ORCID = NA_character_, etal = NA
    ), auto_unbox = TRUE, pretty = FALSE, na = "null"))
  }

  x <- trimws(author_raw)

  # Fix common mojibake (broken encoding)
  if (grepl("Ã", x, fixed = TRUE)) {
    x <- iconv(x, from = "latin1", to = "UTF-8")
  }

  # Remove leftover "(YYYY...)" if present
  x <- sub("\\s*\\((19|20)\\d{2}.*$", "", x)

  # Detect "et al." and strip it but keep flag
  has_etal <- grepl("\\bet\\s+al\\b", x, ignore.case = TRUE)
  x <- gsub("\\s*,?\\s*et\\s+al\\.?\\s*$", "", x, ignore.case = TRUE)

  # Turn "&" into comma separator
  x <- gsub("\\s*&\\s*", ", ", x)

  toks <- trimws(unlist(strsplit(x, "\\s*,\\s*")))
  toks <- toks[nzchar(toks)]

  # If we cannot pair family,given -> treat as corporate/unknown name
  if (length(toks) < 2) {
    return(jsonlite::toJSON(tibble::tibble(
      given = NA_character_, family = NA_character_, name = x,
      sequence = NA_character_, ORCID = NA_character_, etal = has_etal
    ), auto_unbox = TRUE, pretty = FALSE, na = "null"))
  }

  authors <- list()
  i <- 1
  while (i <= length(toks) - 1) {
    fam <- toks[i]
    giv <- toks[i + 1]

    authors[[length(authors) + 1]] <- tibble::tibble(
      given = ifelse(nzchar(giv), giv, NA_character_),
      family = ifelse(nzchar(fam), fam, NA_character_),
      name = NA_character_,
      sequence = NA_character_,
      ORCID = NA_character_,
      etal = has_etal
    )
    i <- i + 2
  }

  if (length(authors) == 0) {
    return(jsonlite::toJSON(tibble::tibble(
      given = NA_character_, family = NA_character_, name = x,
      sequence = NA_character_, ORCID = NA_character_, etal = has_etal
    ), auto_unbox = TRUE, pretty = FALSE, na = "null"))
  }

  jsonlite::toJSON(dplyr::bind_rows(authors), auto_unbox = TRUE, pretty = FALSE, na = "null")
}

# Fill metadata fields from ref_o or ref_r when missing
fill_from_ref_if_missing <- function(df, side = c("o", "r")) {
  side <- match.arg(side)

  ref_col     <- paste0("ref_", side)
  title_col   <- paste0("title_", side)
  author_col  <- paste0("author_", side)
  journal_col <- paste0("journal_", side)
  year_col    <- paste0("year_", side)
  volume_col  <- paste0("volume_", side)
  issue_col   <- paste0("issue_", side)
  pages_col   <- paste0("pages_", side)

  # Parse each ref string
  parsed <- purrr::map(df[[ref_col]], parse_ref_general)

  # Pull parsed fields into vectors
  p_author_raw <- purrr::map_chr(parsed, ~ (.x$author_raw %||% NA_character_)[1])
  p_year       <- purrr::map_int(parsed, ~ as.integer((.x$year %||% NA_integer_)[1]))
  p_title      <- purrr::map_chr(parsed, ~ (.x$title %||% NA_character_)[1])
  p_journal    <- purrr::map_chr(parsed, ~ (.x$journal %||% NA_character_)[1])
  p_volume     <- purrr::map_chr(parsed, ~ (.x$volume %||% NA_character_)[1])
  p_issue      <- purrr::map_chr(parsed, ~ (.x$issue %||% NA_character_)[1])
  p_pages      <- purrr::map_chr(parsed, ~ (.x$pages %||% NA_character_)[1])

  # Convert parsed author_raw -> JSON
  p_author_json <- purrr::map_chr(p_author_raw, author_raw_to_json)

  # Fill missing values only
  df %>%
    dplyr::mutate(
      "{title_col}"   := dplyr::coalesce(.data[[title_col]],   p_title),
      "{author_col}"  := dplyr::coalesce(.data[[author_col]],  p_author_json),
      "{journal_col}" := dplyr::coalesce(.data[[journal_col]], p_journal),
      "{year_col}"    := dplyr::coalesce(.data[[year_col]],    p_year),
      "{volume_col}"  := dplyr::coalesce(.data[[volume_col]],  p_volume),
      "{issue_col}"   := dplyr::coalesce(.data[[issue_col]],   p_issue),
      "{pages_col}"   := dplyr::coalesce(.data[[pages_col]],   p_pages)
    )
}

# Parse DOI metadata table from get_doi_data_unified_cached()
# into standardized columns title_o/author_o/... or title_r/author_r/...
parse_doi_tbl <- function(doi_tbl, side = c("o", "r")) {
  side <- match.arg(side)

  # If empty input -> return empty tibble with expected schema
  if (is.null(doi_tbl) || nrow(doi_tbl) == 0) {
    return(tibble::tibble(
      doi = character(),
      !!paste0("title_", side)   := character(),
      !!paste0("author_", side)  := character(),
      !!paste0("journal_", side) := character(),
      !!paste0("year_", side)    := integer(),
      !!paste0("volume_", side)  := character(),
      !!paste0("issue_", side)   := character(),
      !!paste0("pages_", side)   := character(),
      !!paste0("type_", side)    := character()
    ))
  }

  # Ensure required columns exist
  needed <- c("doi", "title", "author", "container.title", "published", "volume", "issue", "page", "type")
  for (nm in needed) if (!(nm %in% names(doi_tbl))) doi_tbl[[nm]] <- NA_character_

  doi_tbl %>%
    dplyr::mutate(
      # Ensure `author` is JSON text; if not, convert to JSON with jsonify_authors()
      author = ifelse(
        stringr::str_detect(author, "^\\s*\\[\\s*\\{"),
        author,
        purrr::map_chr(author, jsonify_authors)   # <- from Sheet A chunk
      )
    ) %>%
    dplyr::transmute(
      doi = tolower(doi),
      !!paste0("title_", side)   := as.character(title),
      !!paste0("author_", side)  := as.character(author),
      !!paste0("journal_", side) := as.character(`container.title`),
      !!paste0("year_", side)    := readr::parse_integer(as.character(published), na = c("", "NA", "null")),
      !!paste0("volume_", side)  := as.character(volume),
      !!paste0("issue_", side)   := as.character(issue),
      !!paste0("pages_", side)   := as.character(page),
      !!paste0("type_", side)    := as.character(type)
    )
}

# ---- 2) Load Sheet B (reproduction list) ------------------------------------

sheet_b <- readr::read_csv(
  "https://docs.google.com/spreadsheets/d/e/2PACX-1vT0VnLyrf9GCYXtN6l1DgaoLlg6H5r-08Op9eJzSripS1QBSHL031Arc27yUDe0YY7cB4TOnYMm2Vh1/pub?gid=984458430&single=true&output=csv",
  show_col_types = FALSE
)

# Clean column names: lowercase + underscores
clean_names_simple <- function(nms) {
  nms %>%
    tolower() %>%
    trimws() %>%
    gsub("[^a-z0-9]+", "_", .) %>%
    gsub("^_+|_+$", "", .)
}
names(sheet_b) <- clean_names_simple(names(sheet_b))
message("Sheet B cleaned columns: ", paste(names(sheet_b), collapse = " | "))

# ---- 3) Build base Sheet B table + dedupe -----------------------------------

# Sheet B has these columns (as you stated):
# ref_o, doi_o, ref_r, doi_r, url_r, abstract_r, outcome, outcome_quote

flora_b <- sheet_b %>%
  dplyr::transmute(
    doi_o = normalize_doi(.data[["doi_o"]]),
    doi_r = normalize_doi(.data[["doi_r"]]),
    ref_o = .data[["ref_o"]],
    ref_r = .data[["ref_r"]],
    url_r = .data[["url_r"]],
    abstract_r = .data[["abstract_r"]],
    outcome = .data[["outcome"]],
    outcome_quote = .data[["outcome_quote"]]
  ) %>%
  dplyr::filter(!is.na(doi_o)) %>%                 # keep only rows with original DOI
  dplyr::mutate(rep_type = "Reproduction") %>%     # identity column for source B
  dplyr::group_by(doi_o, doi_r, ref_o, ref_r, url_r, abstract_r, rep_type) %>%
  dplyr::summarise(
    outcome = if (length(unique(outcome)) == 1) unique(outcome) else "mixed",
    outcome_quote = paste(unique(na.omit(outcome_quote)), collapse = " | "),
    .groups = "drop"
  )

# ---- 4) Fetch DOI metadata (Crossref/DataCite/cache) ------------------------

# Fetch metadata for original DOIs from Sheet B
doi_data_o_b <- get_doi_data_unified_cached(flora_b$doi_o, cache_file = "flora_doi_cache.rds")

# Fetch metadata for replication DOIs only where doi_r exists
rep_dois_b <- flora_b %>%
  dplyr::filter(!is.na(doi_r)) %>%
  dplyr::pull(doi_r) %>%
  unique()

doi_data_r_b <- get_doi_data_unified_cached(rep_dois_b, cache_file = "flora_doi_cache.rds")

# Parse DOI metadata tables into standardized columns (title_o, author_o, ...)
doi_data_o_b_parsed <- parse_doi_tbl(doi_data_o_b, "o")
doi_data_r_b_parsed <- parse_doi_tbl(doi_data_r_b, "r")

# ---- 5) Join metadata + compute hashes + fill missing from ref strings ------

flora_b_enriched <- flora_b %>%
  dplyr::mutate(
    doi_o = tolower(doi_o),
    doi_r = tolower(doi_r),

    # Always hash doi_o
    doi_o_hash = stringr::str_sub(digest::digest(doi_o, "md5", serialize = FALSE), 1, 3),

    # Hash doi_r only if present; otherwise NA
    doi_r_hash = ifelse(
      is.na(doi_r), NA_character_,
      stringr::str_sub(digest::digest(doi_r, "md5", serialize = FALSE), 1, 3)
    )
  ) %>%
  # Join DOI-derived metadata
  dplyr::left_join(doi_data_o_b_parsed, by = c("doi_o" = "doi")) %>%
  dplyr::left_join(doi_data_r_b_parsed, by = c("doi_r" = "doi")) %>%
  # Fill missing fields using ref_o/ref_r parsing
  fill_from_ref_if_missing("o") %>%
  fill_from_ref_if_missing("r")

# ---- 6) Build citations (Crossref first; manual fallback second) ------------

# Build meta table for originals, using enriched fields
meta_o_b <- flora_b_enriched %>%
  dplyr::distinct(doi_o, title_o, author_o, journal_o, year_o, volume_o, issue_o, pages_o, type_o) %>%
  dplyr::transmute(
    doi         = doi_o,
    title       = title_o,
    author_json = author_o,
    journal     = journal_o,
    year        = year_o,
    volume      = volume_o,
    issue       = issue_o,
    pages       = pages_o,
    type        = type_o
  )

# Crossref-first citations for originals
refs_o_b <- override_with_crossref(doi_vec = meta_o_b$doi, meta_df = meta_o_b)

# Build meta table for replications only where doi_r exists
meta_r_b <- flora_b_enriched %>%
  dplyr::filter(!is.na(doi_r)) %>%
  dplyr::distinct(doi_r, title_r, author_r, journal_r, year_r, volume_r, issue_r, pages_r, type_r) %>%
  dplyr::transmute(
    doi         = doi_r,
    title       = title_r,
    author_json = author_r,
    journal     = journal_r,
    year        = year_r,
    volume      = volume_r,
    issue       = issue_r,
    pages       = pages_r,
    type        = type_r
  )

# Crossref-first citations for replications (if any DOI-r exist)
refs_r_b <- if (nrow(meta_r_b) > 0) {
  override_with_crossref(doi_vec = meta_r_b$doi, meta_df = meta_r_b)
} else {
  tibble::tibble(
    doi = character(),
    apa_ref = character(),
    bibtex_ref = character(),
    apa_source = character(),
    bibtex_source = character()
  )
}

# ---- 7) Assemble Sheet B output rows (match Sheet A schema) -----------------
# IMPORTANT: rowwise() ensures manual fallback runs per-row, not on whole vectors.

flora_b_out <- flora_b_enriched %>%
  # Join original citations
  dplyr::left_join(refs_o_b, by = c("doi_o" = "doi")) %>%
  dplyr::rename(
    apa_ref_o = apa_ref,
    bibtex_ref_o = bibtex_ref
  ) %>%
  # Join replication citations (for DOI-r rows)
  dplyr::left_join(refs_r_b, by = c("doi_r" = "doi")) %>%
  dplyr::rename(
    apa_ref_r_doi = apa_ref,
    bibtex_ref_r_doi = bibtex_ref
  ) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    # Replication APA:
    # - If DOI-r exists and crossref-first produced something, use it
    # - Else build manually from parsed/ref-filled fields
    apa_ref_r = if (!is.na(doi_r) && !is.na(apa_ref_r_doi) && nzchar(apa_ref_r_doi)) {
      apa_ref_r_doi
    } else {
      apa_reference_md(
        doi         = if (!is.na(doi_r) && nzchar(doi_r)) doi_r else NA_character_,
        title       = title_r,
        author_json = author_r,
        journal     = journal_r,
        year        = year_r,
        volume      = volume_r,
        issue       = issue_r,
        pages       = pages_r
      )
    },

    # Replication BibTeX:
    # - If DOI-r exists and crossref-first produced something, use it
    # - Else build manually
    bibtex_ref_r = if (!is.na(doi_r) && !is.na(bibtex_ref_r_doi) && nzchar(bibtex_ref_r_doi)) {
      bibtex_ref_r_doi
    } else {
      bibtex_reference(
        doi         = if (!is.na(doi_r) && nzchar(doi_r)) doi_r else NA_character_,
        title       = title_r,
        author_json = author_r,
        journal     = journal_r,
        year        = year_r,
        volume      = volume_r,
        issue       = issue_r,
        pages       = pages_r
      )
    }
  ) %>%
  dplyr::ungroup() %>%
  # Final column order (aligned with your Sheet A export + Sheet B extras)
  dplyr::select(
    doi_o, doi_o_hash, title_o, author_o, journal_o, year_o, volume_o, issue_o, pages_o,
    apa_ref_o, bibtex_ref_o,

    doi_r, doi_r_hash, rep_type, title_r, author_r, journal_r, year_r, volume_r, issue_r, pages_r,
    apa_ref_r, bibtex_ref_r,

    ref_o, ref_r, url_r, abstract_r, outcome, outcome_quote
  )

# ---- 8) Bind A + B, fix type conflicts  -------------------

# Columns that commonly cause bind_rows() conflicts (numeric vs character)
# Force them to character in BOTH datasets before binding.
char_cols <- c(
  "doi_o", "doi_r", "doi_o_hash", "doi_r_hash",
  "volume_o", "issue_o", "pages_o",
  "volume_r", "issue_r", "pages_r",
  "year_o", "year_r"
)

make_char_if_exists <- function(df, cols) {
  cols <- intersect(cols, names(df))
  df %>% dplyr::mutate(dplyr::across(dplyr::all_of(cols), as.character))
}

if (!("rep_type" %in% names(flora_a_out))) {
  flora_a_out <- flora_a_out %>% dplyr::mutate(rep_type = "doi")
}

# IMPORTANT: create `flora` (final dataframe) but do NOT write the CSV yet
flora <- dplyr::bind_rows(
  make_char_if_exists(flora_a_out, char_cols),
  make_char_if_exists(flora_b_out, char_cols)
)


```
# MANUAL OVERRIDES : Reads manual_references.xlsx and replaces APA/BibTeX in `flora`
```{r}
if (!requireNamespace("readxl", quietly = TRUE)) {
  stop("Package 'readxl' is required. Please run: install.packages('readxl')")
}
library(readxl)

# Robust: folder where the .qmd/.Rmd lives when rendering; otherwise fall back to getwd()
here_dir <- tryCatch(knitr::current_input(dir = TRUE), error = function(e) NULL)
if (is.null(here_dir) || length(here_dir) == 0 || is.na(here_dir) || !nzchar(as.character(here_dir))) {
  here_dir <- getwd()
}
here_dir <- as.character(here_dir)[1]

manual_path <- file.path(here_dir, "manual_references.xlsx")



manual_refs <- readxl::read_excel(manual_path) %>%
  dplyr::rename(
    manual_key    = key,
    manual_apa    = reference_apa,
    manual_bibtex = reference_bibtex
  ) %>%
  dplyr::mutate(
    manual_key = as.character(manual_key),
    manual_key = trimws(manual_key),

    # Normalize DOI-looking keys (strip doi.org etc, lowercase).
    manual_key_norm = dplyr::if_else(
      stringr::str_detect(manual_key, "^\\s*(https?://(dx\\.)?doi\\.org/)?10\\.[0-9]{4,9}/"),
      {
        k <- gsub("^\\s*https?://(dx\\.)?doi\\.org/", "", manual_key, ignore.case = TRUE)
        k <- URLdecode(k)
        k <- trimws(tolower(k))
        k
      },
      trimws(manual_key)  # URL keys: keep as-is (trimmed)
    ),

    # Treat blanks as NA
    manual_apa    = dplyr::na_if(trimws(as.character(manual_apa)), ""),
    manual_bibtex = dplyr::na_if(trimws(as.character(manual_bibtex)), "")
  ) %>%
  dplyr::select(manual_key_norm, manual_apa, manual_bibtex) %>%
  dplyr::distinct(manual_key_norm, .keep_all = TRUE)

# ---- Prepare match keys in flora --------------------------------------------
flora_overridden <- flora %>%
  dplyr::mutate(
    doi_o = as.character(doi_o),
    doi_r = as.character(doi_r),

    # url_o may or may not exist (we handle both safely)
    url_o = if ("url_o" %in% names(.)) as.character(url_o) else NA_character_,
    url_r = if ("url_r" %in% names(.)) as.character(url_r) else NA_character_,

    doi_o_norm = ifelse(is.na(doi_o) | !nzchar(trimws(doi_o)), NA_character_, trimws(tolower(doi_o))),
    doi_r_norm = ifelse(is.na(doi_r) | !nzchar(trimws(doi_r)), NA_character_, trimws(tolower(doi_r))),
    url_o_norm = ifelse(is.na(url_o) | !nzchar(trimws(url_o)), NA_character_, trimws(url_o)),
    url_r_norm = ifelse(is.na(url_r) | !nzchar(trimws(url_r)), NA_character_, trimws(url_r))
  )

# ---- Override ORIGINAL refs by matching manual_key_norm == doi_o ------------
flora_overridden <- flora_overridden %>%
  dplyr::left_join(
    manual_refs,
    by = c("doi_o_norm" = "manual_key_norm")
  ) %>%
  dplyr::mutate(
    apa_ref_o    = dplyr::coalesce(manual_apa, apa_ref_o),
    bibtex_ref_o = dplyr::coalesce(manual_bibtex, bibtex_ref_o)
  ) %>%
  dplyr::select(-manual_apa, -manual_bibtex)

# ---- Override ORIGINAL refs by matching manual_key_norm == url_o (ADDED) ----
flora_overridden <- flora_overridden %>%
  dplyr::left_join(
    manual_refs,
    by = c("url_o_norm" = "manual_key_norm")
  ) %>%
  dplyr::mutate(
    apa_ref_o    = dplyr::coalesce(manual_apa, apa_ref_o),
    bibtex_ref_o = dplyr::coalesce(manual_bibtex, bibtex_ref_o)
  ) %>%
  dplyr::select(-manual_apa, -manual_bibtex)

# ---- Override REPLICATION refs by matching manual_key_norm == doi_r ----------
flora_overridden <- flora_overridden %>%
  dplyr::left_join(
    manual_refs,
    by = c("doi_r_norm" = "manual_key_norm")
  ) %>%
  dplyr::mutate(
    apa_ref_r    = dplyr::coalesce(manual_apa, apa_ref_r),
    bibtex_ref_r = dplyr::coalesce(manual_bibtex, bibtex_ref_r)
  ) %>%
  dplyr::select(-manual_apa, -manual_bibtex)

# ---- Override REPLICATION refs by matching manual_key_norm == url_r ----------
flora_overridden <- flora_overridden %>%
  dplyr::left_join(
    manual_refs,
    by = c("url_r_norm" = "manual_key_norm")
  ) %>%
  dplyr::mutate(
    apa_ref_r    = dplyr::coalesce(manual_apa, apa_ref_r),
    bibtex_ref_r = dplyr::coalesce(manual_bibtex, bibtex_ref_r)
  ) %>%
  dplyr::select(-manual_apa, -manual_bibtex)

# ---- Clean helper columns + write final CSV ---------------------------------
flora_final <- flora_overridden %>%
  dplyr::select(-doi_o_norm, -doi_r_norm, -url_o_norm, -url_r_norm)

sum(flora$apa_ref_o != flora_final$apa_ref_o, na.rm = TRUE)
sum(flora$apa_ref_r != flora_final$apa_ref_r, na.rm = TRUE)

readr::write_excel_csv(flora_final, "flora.csv")
```
