---
title: "FLoRA: FORRT Library of Replications and Reproductions"
format: html
---

# FLoRA - FORRT Library of Replications and Reproductions

This document generates the **FLoRA** (FORRT Library of Replications and Reproductions) dataset, which consolidates replication and reproduction data from the FORRT Replication Database.

The dataset includes:
- Bibliometric details from CrossRef and OpenAlex
- Outcome information with quotes and sources
- Metapaper coding
- Publication status classification
- Citation caching for efficiency

## Setup

```{r setup}
library(dplyr)
library(readr)
library(rcrossref)
library(purrr)
library(jsonlite)
library(glue)
library(stringr)
library(rlang)
library(httr)

# Load replication data and keep only rows with valid DOIs
reported_success <- read_csv(
  "https://docs.google.com/spreadsheets/d/e/2PACX-1vSlPVImL7kjzMfpfX08JtAikvEiBg8uQ3HyEJs-YvuQMUa2RMvip-4v8-DilWl2PdwOD_UTD8gJ5wKA/pub?gid=984458430&single=true&output=csv",
  skip = 1,
  show_col_types = FALSE
) %>%
  # ✅ validate DOI prefix: starts with 10., has 4–9 digits, and then a slash
  filter(str_detect(doi_o, regex("^10\\.[0-9]{4,9}/", ignore_case = TRUE))) %>%
  filter(str_detect(doi_r, regex("^10\\.[0-9]{4,9}/", ignore_case = TRUE)))

```

## Select relevant fields and deduplicate

We keep only validated entries and deduplicate at the original-replication pair level.

```{r deduplication}
flora <- reported_success %>% 
  filter(validation == "validated - chosen")  %>% 
  select(doi_o, doi_r, ref_o, ref_r, url_r, outcome, outcome_quote, out_quote_source) %>%
  group_by(doi_o, doi_r, ref_o, ref_r, url_r) %>%
  summarise(
    outcome = if (length(unique(outcome)) == 1) unique(outcome) else "mixed",
    outcome_quote = if (length(unique(outcome_quote)) == 1) unique(outcome_quote) else paste(outcome_quote, collapse = "; "),
    out_quote_source = paste(out_quote_source, collapse = "; "),
    .groups = "drop"
  ) %>% 
  distinct(doi_o, doi_r, .keep_all = TRUE)
```

## Helpers for CrossRef metadata

```{r crossref_helpers}
'%||%' <- function(a, b) if (!is.null(a) && length(a) > 0) a else b

get_doi_data <- function(doi) {
  # Clean and de-duplicate DOIs before querying CrossRef
  doi <- str_replace_all(doi, "https?://doi.org/", "")
  doi <- str_subset(doi, "^10\\.\\d+/\\S+$")
  doi <- unique(doi)
  
  message(glue("Fetching data for {length(doi)} DOIs from CrossRef..."))
  
  if (length(doi) == 0) return(tibble())
  
  # Query CrossRef in batches of up to 100 DOIs to avoid URL length errors
  for (i in seq(1, length(doi), by = 100)) {
    doisub <- doi[i:min(i+99, length(doi))]
    names(doisub) <- rep("doi", length(doisub))
    res <- cr_works(filter = doisub, limit = 100)
    message(glue("  Retrieved {nrow(res$data)} records (of {length(doisub)} requested)"))
    if(i == 1) out <- res$data else out <- bind_rows(out, res$data)
  }
  message(glue("Done. Retrieved {nrow(out)} records in total."))
  
  return(out)
}


jsonify_authors <- function(author_list) {
  
  if (all(c("family", "given") %in% names(author_list))) {
    out <- tibble(
      given = author_list[["given"]] %||% NA_character_,
      family = author_list[["family"]] %||% NA_character_,
      sequence = author_list[["sequence"]] %||% NA_character_,
      ORCID = author_list[["ORCID"]] %||% NA_character_
    )
  } else if ("name" %in% names(author_list)) {
    out <- tibble(
      given = author_list[["given"]] %||% NA_character_,
      family = author_list[["name"]] %||% NA_character_,
      sequence = author_list[["sequence"]] %||% NA_character_,
      ORCID = author_list[["ORCID"]] %||% NA_character_
    )
  } else {
    message("Unknown author format - field names: ", paste(names(author_list), collapse = ", "), ". Returning NULL entry.")
    out <- tibble(
      given = author_list[["given"]] %||% NA_character_,
      family = author_list[["name"]] %||% NA_character_,
      sequence = author_list[["sequence"]] %||% NA_character_,
      ORCID = author_list[["ORCID"]] %||% NA_character_
    )
  }
  
  out |>
    toJSON(pretty = FALSE, na = "null")
}


safe_year <- function(df) {
  pub_cols <- c("published.print", "published.online", "published")
  existing <- pub_cols[pub_cols %in% names(df)]

  if (length(existing) == 0) {
    return(NA_integer_)
  }

  year_str <- dplyr::coalesce(!!!df[existing])
  as.integer(stringr::str_sub(as.character(year_str), 1, 4))
}

apa_reference_md <- function(doi, title, author_json, journal, year, volume, issue, pages) {
  n <- length(doi)

  map_chr(seq_len(n), function(i) {
    # --- parse authors safely ---
    authors <- tryCatch(fromJSON(author_json[i]), error = function(e) NULL)

    fmt_author <- function(given, family) {
      initials <- if (!is.na(given) && nzchar(given)) {
        parts <- strsplit(given, "\\s+")[[1]]
        paste0(substr(parts, 1, 1), collapse = ". ")
      } else ""
      paste0(family, ", ", initials, ".")
    }

    if (is.null(authors) || nrow(authors) == 0) {
      author_str <- "(n.a.)"
    } else {
      author_strs <- pmap_chr(authors, \(given, family, sequence, ORCID)
        fmt_author(given, family))
      n_auth <- length(author_strs)
      author_str <-
        if (n_auth == 1) author_strs
        else if (n_auth == 2) paste(author_strs, collapse = " & ")
        else paste0(paste(author_strs[-n_auth], collapse = ", "), ", & ", author_strs[n_auth])
    }

    # --- APA-style markdown (title not bold) ---
    glue(
      "{author_str} ({year[i]}). {title[i]}. *{journal[i]}*, *{volume[i]}*({issue[i]}), {pages[i]}. [https://doi.org/{doi[i]}](https://doi.org/{doi[i]})"
    )
  })
}

bibtex_reference <- function(doi, title, author_json, journal, year, volume, issue, pages) {
  n <- length(doi)

  map_chr(seq_len(n), function(i) {
    # --- parse authors safely ---
    authors <- tryCatch(fromJSON(author_json[i]), error = function(e) NULL)

    fmt_author_bib <- function(given, family) {
      if (!is.na(given) && nzchar(given)) {
        paste0(family, ", ", given)
      } else {
        family
      }
    }

    if (is.null(authors) || nrow(authors) == 0) {
      author_str <- "Unknown"
    } else {
      author_strs <- pmap_chr(authors, \(given, family, sequence, ORCID)
        fmt_author_bib(given, family))
      author_str <- paste(author_strs, collapse = " and ")
    }

    key <- str_remove_all(paste0(str_to_lower(word(author_strs[1], 1)), year[i]), "[^a-z0-9]")

    glue(
"@article{{{key},
  author = {{{author_str}}},
  title = {{{title[i]}}},
  journal = {{{journal[i]}}},
  year = {{{year[i]}}},
  volume = {{{volume[i]}}},
  number = {{{issue[i]}}},
  pages = {{{pages[i]}}},
  doi = {{{doi[i]}}}
}}"
    )
  })
}
```

## Citation cache for CrossRef lookups

```{r citation_cache}
# File on disk where we cache citation strings from CrossRef
CIT_CACHE_FILE <- "crossref_citation_cache.rds"

load_citation_cache <- function() {
  if (file.exists(CIT_CACHE_FILE)) {
    readRDS(CIT_CACHE_FILE)
  } else {
    list(apa = list(), bibtex = list())
  }
}

save_citation_cache <- function(cache) {
  saveRDS(cache, CIT_CACHE_FILE)
}

get_crossref_citation <- function(doi, type = c("bibtex", "apa")) {
  type <- match.arg(type)

  if (type == "bibtex") {
    fmt   <- "bibtex"
    style <- NULL
  } else if (type == "apa") {
    fmt   <- "text"
    style <- "apa"
  }

  tryCatch(
    rcrossref::cr_cn(dois = doi, format = fmt, style = style),
    error = function(e) NA_character_
  )
}
```

## Fetch and parse CrossRef metadata

```{r fetch_crossref}
doi_data_o <- get_doi_data(flora$doi_o)

doi_data_o_parsed <- doi_data_o %>% 
  transmute(
    doi = tolower(doi),
    title_o = title,
    author_o = map_chr(author, jsonify_authors),
    journal_o = container.title,
    year_o = safe_year(pick(everything())),
    volume_o = volume,
    issue_o = issue,
    pages_o = page,
    type_o = type
  )

doi_data_r <- get_doi_data(flora$doi_r)

doi_data_r_parsed <- doi_data_r %>% 
  transmute(
    doi = tolower(doi),
    title_r = title,
    author_r = map_chr(author, jsonify_authors),
    journal_r = container.title,
    year_r = safe_year(pick(everything())),
    volume_r = volume,
    issue_r = issue,
    pages_r = page,
    type_r = type
  )
```

## Build APA/BibTeX references with CrossRef caching

```{r build_references}
# Build unified metadata tables for use in citation generation
meta_o <- doi_data_o_parsed %>%
  transmute(
    doi,
    title = title_o,
    author_json = author_o,
    journal = journal_o,
    year = year_o,
    volume = volume_o,
    issue = issue_o,
    pages = pages_o,
    type = type_o
  )

meta_r <- doi_data_r_parsed %>%
  transmute(
    doi,
    title = title_r,
    author_json = author_r,
    journal = journal_r,
    year = year_r,
    volume = volume_r,
    issue = issue_r,
    pages = pages_r,
    type = type_r
  )

# Function to resolve APA/BibTeX via CrossRef first, with manual fallback
override_with_crossref <- function(doi_vec, meta_df) {
  cache <- load_citation_cache()
  doi_vec <- tolower(doi_vec)

  apa_final <- vector("character", length(doi_vec))
  bibtex_final <- vector("character", length(doi_vec))

  for (i in seq_along(doi_vec)) {
    d <- doi_vec[i]
    meta_row <- meta_df[match(d, meta_df$doi), ]

    # --- APA ---
    apa_cr <- cache$apa[[d]] %||% get_crossref_citation(d, "apa")
    cache$apa[[d]] <- apa_cr

    if (is.na(apa_cr) || !nzchar(apa_cr)) {
      if (nrow(meta_row) == 1 && !is.na(meta_row$type) && meta_row$type == "journal-article") {
        apa_fallback <- apa_reference_md(
          doi = meta_row$doi,
          title = meta_row$title,
          author_json = meta_row$author_json,
          journal = meta_row$journal,
          year = meta_row$year,
          volume = meta_row$volume,
          issue = meta_row$issue,
          pages = meta_row$pages
        )[1]
      } else if (nrow(meta_row) == 1 && !is.na(meta_row$title)) {
        apa_fallback <- glue("{meta_row$title}. https://doi.org/{meta_row$doi}")
      } else {
        apa_fallback <- glue("https://doi.org/{d}")
      }
      apa_final[i] <- apa_fallback
    } else {
      apa_final[i] <- apa_cr
    }

    # --- BibTeX ---
    bib_cr <- cache$bibtex[[d]] %||% get_crossref_citation(d, "bibtex")
    cache$bibtex[[d]] <- bib_cr

    if (is.na(bib_cr) || !nzchar(bib_cr)) {
      if (nrow(meta_row) == 1) {
        bibtex_fallback <- bibtex_reference(
          doi = meta_row$doi,
          title = meta_row$title,
          author_json = meta_row$author_json,
          journal = meta_row$journal,
          year = meta_row$year,
          volume = meta_row$volume,
          issue = meta_row$issue,
          pages = meta_row$pages
        )[1]
      } else {
        bibtex_fallback <- glue("@misc{{{d}, doi = {{{d}}}}}")
      }
      bibtex_final[i] <- bibtex_fallback
    } else {
      bibtex_final[i] <- bib_cr
    }
  }

  save_citation_cache(cache)

  tibble(
    doi = doi_vec,
    apa_ref = apa_final,
    bibtex_ref = bibtex_final
  )
}

# Generate references for originals and replications
refs_o <- override_with_crossref(doi_vec = meta_o$doi, meta_df = meta_o)
refs_r <- override_with_crossref(doi_vec = meta_r$doi, meta_df = meta_r)
```

## OpenAlex metadata (keywords, field, language)

```{r openalex}
get_openalex_meta <- function(doi, mailto = NULL, fallback_to_concepts = TRUE, n_concepts = 10) {
  base_url <- "https://api.openalex.org/works/"
  encoded_doi <- URLencode(doi, reserved = TRUE)
  url <- paste0(base_url, "doi:", encoded_doi)
  
  q <- list(select = "id,doi,display_name,language,keywords,concepts,primary_topic")
  if (!is.null(mailto)) q$mailto <- mailto
  
  resp <- GET(url, query = q, user_agent("R script (OpenAlex lookup)"))
  if (http_error(resp)) {
    stop(sprintf("HTTP %s: %s", status_code(resp), content(resp, as = "text", encoding = "UTF-8")))
  }
  x <- fromJSON(content(resp, as = "text", encoding = "UTF-8"), simplifyVector = TRUE)
  
  # language
  lang <- x$language %||% NA_character_
  
  # keywords
  kws <- NULL
  if (!is.null(x$keywords)) {
    if (is.character(x$keywords)) {
      kws <- x$keywords
    } else if (is.data.frame(x$keywords)) {
      nm <- names(x$keywords)
      pick <- intersect(c("display_name", "keyword", "name", "value"), nm)
      if (length(pick)) kws <- unique(na.omit(as.character(x$keywords[[pick[1]]])))
    } else if (is.list(x$keywords)) {
      kws <- unique(na.omit(as.character(unlist(lapply(
        x$keywords,
        function(k) k[["display_name"]] %||% k[["keyword"]] %||% k[["name"]] %||% k
      )))))
    }
  }
  
  # Fallback to top-N concepts for keywords if needed
  if ((is.null(kws) || length(kws) == 0) && fallback_to_concepts && !is.null(x$concepts)) {
    cons <- x$concepts
    if (is.data.frame(cons) && "display_name" %in% names(cons)) {
      if ("score" %in% names(cons)) cons <- cons[order(cons$score, decreasing = TRUE), , drop = FALSE]
      kws <- head(unique(as.character(cons$display_name)), n_concepts)
    } else if (is.list(cons)) {
      dn <- unlist(lapply(cons, function(cn) cn[["display_name"]]))
      sc <- unlist(lapply(cons, function(cn) cn[["score"]]))
      ord <- order(ifelse(is.na(sc), 0, sc), decreasing = TRUE)
      kws <- head(unique(as.character(dn[ord])), n_concepts)
    }
  }
  kws <- unique(trimws(kws))
  kws <- kws[nzchar(kws)]
  
  # field: prefer primary_topic$field$display_name, otherwise infer from concepts
  field <- NA_character_
  if (!is.null(x$primary_topic)) {
    pt <- x$primary_topic
    if (!is.null(pt$field) && !is.null(pt$field$display_name)) {
      field <- as.character(pt$field$display_name)
    } else if (!is.null(pt$subfield) && !is.null(pt$subfield$display_name)) {
      field <- as.character(pt$subfield$display_name)
    } else if (!is.null(pt$display_name)) {
      field <- as.character(pt$display_name)
    }
  }
  if ((is.na(field) || !nzchar(field)) && !is.null(x$concepts)) {
    cons <- x$concepts
    # normalize to vectors
    if (is.data.frame(cons)) {
      dn <- as.character(cons$display_name %||% rep(NA_character_, nrow(cons)))
      lv <- suppressWarnings(as.integer(cons$level %||% rep(NA_integer_, nrow(cons))))
      sc <- suppressWarnings(as.numeric(cons$score %||% rep(NA_real_, nrow(cons))))
    } else if (is.list(cons)) {
      dn <- as.character(unlist(lapply(cons, function(cn) cn[["display_name"]] %||% NA_character_)))
      lv <- suppressWarnings(as.integer(unlist(lapply(cons, function(cn) cn[["level"]] %||% NA_integer_))))
      sc <- suppressWarnings(as.numeric(unlist(lapply(cons, function(cn) cn[["score"]] %||% NA_real_))))
    } else {
      dn <- lv <- sc <- NULL
    }
    if (!is.null(dn) && length(dn)) {
      # prefer level 1 (field), then 0 (domain), else best-scoring
      pick_idx <- which(lv == 1)
      if (!length(pick_idx)) pick_idx <- which(lv == 0)
      if (!length(pick_idx)) pick_idx <- seq_along(dn)
      # choose highest scoring among candidates
      sc_sub <- ifelse(is.na(sc[pick_idx]), 0, sc[pick_idx])
      field <- dn[pick_idx[order(sc_sub, decreasing = TRUE)][1]]
    }
  }
  
  list(
    keywords = if (length(kws)) kws else character(0),
    field = field %||% NA_character_,
    language = lang %||% NA_character_
  )
}

# Prepare DOIs
dois <- unique(flora$doi_o)
dois <- dois[!is.na(dois) & grepl("^10", dois)]

# Download from OpenAlex
kwdat <- data.frame(
  doi_o = dois,
  keywords_o = NA_character_,
  field_o = NA_character_,
  language_o = NA_character_,
  error = NA_character_,
  stringsAsFactors = FALSE
)

for (idx in seq_along(dois)) {
  res <- tryCatch(
    {
      out <- get_openalex_meta(dois[idx], mailto = "lukas.roeseler@uni-muenster.de")
      list(out = out, err = NA_character_)
    },
    error = function(e) list(out = NULL, err = conditionMessage(e))
  )
  
  if (!is.null(res$out)) {
    kwdat$keywords_o[idx] <- if (length(res$out$keywords) > 0) paste(res$out$keywords, collapse = ", ") else NA_character_
    kwdat$field_o[idx] <- res$out$field %||% NA_character_
    kwdat$language_o[idx] <- res$out$language %||% NA_character_
  }
  kwdat$error[idx] <- res$err
}

kwdat <- kwdat[, 1:4]  # drop error column
```

## Metapaper coding

```{r metapaper_coding}
# code any of these entries as meta papers
metapapers <- c(
  "10.1038--s41562-024-02062-9",
  "10.3389/fcomm.2022.1048896", # 10 cases of sensory research
  "10.1073/pnas.2103313118", # scarcity effects
  "10.1177/0956797619831612", # Soto
  "10.1007/s13164-018-0400-9", # X-PHI
  "10.1038/s41562-018-0399-z", # Camerer 2018
  "10.1098/rsos.231240", # Boyce
  "10.1027/1864-9335/a000178", # ML1
  "10.1177/2515245918810225", # ML2
  "10.1016/j.jesp.2015.10.012", # ML3
  "10.1126/science.aac4716", # RP:P
  "10.1038/s41562-024-02062-9", # Holzmeister et al. 2024
  "10.1177/08902070221094216", # SVO development
  "10.1126/science.aaf0918" # Camerer 2016
)
```

## Publication status

```{r publication_status}
preprint_servers <- c(
  "SSRN Electronic Journal",
  "Open Research Europe"
)
```

## Combine and finalize FLoRA dataset

```{r combine_finalize}
flora <- flora %>% 
  rowwise() %>% 
  mutate(
    doi_o = tolower(doi_o),
    doi_r = tolower(doi_r),
    doi_o_hash = str_sub(digest::digest(doi_o, "md5", serialize = FALSE), 1, 3),
    doi_r_hash = str_sub(digest::digest(doi_r, "md5", serialize = FALSE), 1, 3)
  ) %>%
  ungroup() %>% 
  left_join(doi_data_o_parsed, by = c("doi_o" = "doi")) %>%
  left_join(doi_data_r_parsed, by = c("doi_r" = "doi")) %>%
  left_join(refs_o, by = c("doi_o" = "doi")) %>%
  rename(apa_ref_o = apa_ref, bibtex_ref_o = bibtex_ref) %>%
  left_join(refs_r, by = c("doi_r" = "doi")) %>%
  rename(apa_ref_r = apa_ref, bibtex_ref_r = bibtex_ref) %>%
  left_join(kwdat, by = "doi_o") %>%
  # Add metapaper coding
  mutate(metapaper_r = as.integer(doi_r %in% metapapers)) %>%
  # Add publication status
  mutate(
    publicationstatus_r = case_when(
      is.na(journal_r) ~ "Working paper",
      journal_r %in% preprint_servers ~ "Not peer-reviewed (e.g., preprint)",
      TRUE ~ "Peer-reviewed (published in a journal)"
    )
  ) %>%
  # Select final columns
  select(
    doi_o, doi_o_hash, title_o, author_o, journal_o, year_o, volume_o, issue_o, pages_o,
    apa_ref_o, bibtex_ref_o,
    doi_r, doi_r_hash, title_r, author_r, journal_r, year_r, volume_r, issue_r, pages_r,
    apa_ref_r, bibtex_ref_r,
    ref_o, ref_r, url_r, outcome, outcome_quote, out_quote_source,
    keywords_o, field_o, language_o,
    metapaper_r, publicationstatus_r
  ) %>%
  # Filter out rows without both titles
  filter(!is.na(title_o), !is.na(title_r))

# Apply exclusions: keep only those metastudies with url_r non-missing
flora <- flora %>%
  filter(!(metapaper_r == 1 & is.na(url_r)))
```

## Save FLoRA dataset

```{r save}
write_csv(flora, "FLoRA.csv")
openxlsx::write.xlsx(flora, file = "FLoRA.xlsx")
```
