---
title: "FRiDa Creator"
format: html
---

# LibRe²

The successor to the meta-analyses-focused FORRT Replication Database (FReD) is a library of replications **and reproductions** that does not include methodological features such as sample and effect sizes.

Notes: This is heavily based on the FLoRA creator. Currently, it includes code to download crossref keywords. It is, however, also possible to attach the already downloaded keywords instead of getting them all from scratch.

## TO DOs

-   add further variables

```{r setup}
library(dplyr)
library(readr)
library(rcrossref)
library(purrr)
library(jsonlite)
library(glue)
library(stringr)
library(rlang)
library(httr)

reported_success <- read_csv(
  "https://docs.google.com/spreadsheets/d/e/2PACX-1vSlPVImL7kjzMfpfX08JtAikvEiBg8uQ3HyEJs-YvuQMUa2RMvip-4v8-DilWl2PdwOD_UTD8gJ5wKA/pub?gid=984458430&single=true&output=csv",
  skip = 1,
  show_col_types = FALSE
) %>%
  # ✅ validate DOI prefix: starts with 10., has 4–9 digits, and then a slash
  filter(str_detect(doi_o, regex("^10\\.[0-9]{4,9}/", ignore_case = TRUE))) %>%
  filter(str_detect(doi_r, regex("^10\\.[0-9]{4,9}/", ignore_case = TRUE)))

```

## Select relevant fields and deduplicate (we do not report study level)

```{r deduplication}
flora <- reported_success %>% filter(validation == "validated - chosen")  %>% select(doi_o, doi_r, ref_o, ref_r, url_r, outcome, outcome_quote, out_quote_source) %>%
  group_by(doi_o, doi_r, ref_o, ref_r, url_r) %>%
  summarise(
  outcome = if (length(unique(outcome)) == 1) unique(outcome) else "mixed",
  outcome_quote = if (length(unique(outcome_quote)) == 1) unique(outcome_quote) else paste(outcome_quote, collapse = "; "),
  out_quote_source = paste(out_quote_source, collapse = "; "),
  .groups = "drop"
  ) %>% 
  distinct(doi_o,doi_r,.keep_all = TRUE)
```

## Augment with bibliometric details

```{r bibliometric augmentation}
get_doi_data <- function(doi) {
  #100 at a time to avoid URL too long CR error
  doi <- str_replace_all(doi, "https?://doi.org/", "")
  doi <- str_subset(doi, "^10\\.\\d+/\\S+$")
  doi <- unique(doi)
  
  message(glue("Fetching data for {length(doi)} DOIs from CrossRef..."))
  
  if (length(doi) == 0) return(tibble())
  
  for (i in seq(1, length(doi), by = 100)) {
    doisub <- doi[i:min(i+99, length(doi))]
    names(doisub) <- rep("doi", length(doisub))
    res <- cr_works(filter = doisub, limit = 100)
    message(glue("  Retrieved {nrow(res$data)} records (of {length(doisub)} requested)"))
    if(i == 1) out <- res$data else out <- bind_rows(out, res$data)
  }
  message(glue("Done. Retrieved {nrow(out)} records in total."))
  
  return(out)
}


jsonify_authors <- function(author_list) {
  
  if (all(c("family", "given") %in% names(author_list))) {
    out <-   tibble(
    given = author_list[["given"]] %||% NA_character_,
    family = author_list[["family"]] %||% NA_character_,
    sequence = author_list[["sequence"]] %||% NA_character_,
    ORCID = author_list[["ORCID"]] %||% NA_character_
  )
  } else if ("name" %in% names(author_list)) {
      out <-   tibble(
    given = author_list[["given"]] %||% NA_character_,
    family = author_list[["name"]] %||% NA_character_,
    sequence = author_list[["sequence"]] %||% NA_character_,
    ORCID = author_list[["ORCID"]] %||% NA_character_
  )

} else {
  message("Unknown author format - field names: ", paste(names(author_list), collapse = ", "), ". Returning NULL entry.")
        out <-   tibble(
    given = author_list[["given"]] %||% NA_character_,
    family = author_list[["name"]] %||% NA_character_,
    sequence = author_list[["sequence"]] %||% NA_character_,
    ORCID = author_list[["ORCID"]] %||% NA_character_
  )
}
  
 out |>
    toJSON(pretty = FALSE, na = "null")
}


safe_year <- function(df) {
  pub_cols <- c("published.print", "published.online", "published")
  existing <- pub_cols[pub_cols %in% names(df)]

  if (length(existing) == 0) {
    return(NA_integer_)
  }

  year_str <- dplyr::coalesce(!!!df[existing])
  as.integer(stringr::str_sub(as.character(year_str), 1, 4))
}

apa_reference_md <- function(doi, title, author_json, journal, year, volume, issue, pages) {
  n <- length(doi)

  map_chr(seq_len(n), function(i) {
    # --- parse authors safely ---
    authors <- tryCatch(fromJSON(author_json[i]), error = function(e) NULL)

    fmt_author <- function(given, family) {
      initials <- if (!is.na(given) && nzchar(given)) {
        parts <- strsplit(given, "\\s+")[[1]]
        paste0(substr(parts, 1, 1), collapse = ". ")
      } else ""
      paste0(family, ", ", initials, ".")
    }

    if (is.null(authors) || nrow(authors) == 0) {
      author_str <- "(n.a.)"
    } else {
      author_strs <- pmap_chr(authors, \(given, family, sequence, ORCID)
        fmt_author(given, family))
      n_auth <- length(author_strs)
      author_str <-
        if (n_auth == 1) author_strs
        else if (n_auth == 2) paste(author_strs, collapse = " & ")
        else paste0(paste(author_strs[-n_auth], collapse = ", "), ", & ", author_strs[n_auth])
    }

    # --- APA-style markdown (title not bold) ---
    glue(
      "{author_str} ({year[i]}). {title[i]}. *{journal[i]}*, *{volume[i]}*({issue[i]}), {pages[i]}. [https://doi.org/{doi[i]}](https://doi.org/{doi[i]})"
    )
  })
}

bibtex_reference <- function(doi, title, author_json, journal, year, volume, issue, pages) {
  n <- length(doi)

  map_chr(seq_len(n), function(i) {
    # --- parse authors safely ---
    authors <- tryCatch(fromJSON(author_json[i]), error = function(e) NULL)

    fmt_author_bib <- function(given, family) {
      if (!is.na(given) && nzchar(given)) {
        paste0(family, ", ", given)
      } else {
        family
      }
    }

    if (is.null(authors) || nrow(authors) == 0) {
      author_str <- "Unknown"
    } else {
      author_strs <- pmap_chr(authors, \(given, family, sequence, ORCID)
        fmt_author_bib(given, family))
      author_str <- paste(author_strs, collapse = " and ")
    }

    key <- str_remove_all(paste0(str_to_lower(word(author_strs[1], 1)), year[i]), "[^a-z0-9]")

    glue(
"@article{{{key},
  author = {{{author_str}}},
  title = {{{title[i]}}},
  journal = {{{journal[i]}}},
  year = {{{year[i]}}},
  volume = {{{volume[i]}}},
  number = {{{issue[i]}}},
  pages = {{{pages[i]}}},
  doi = {{{doi[i]}}}
}}"
    )
  })
}

doi_data_o <- get_doi_data(flora$doi_o)

doi_data_o_parsed <- doi_data_o %>% 
  transmute(
    doi = tolower(doi),
    title_o = title,
    author_o = map_chr(author, jsonify_authors),
    journal_o = container.title,
 year_o = safe_year(pick(everything())),
    volume_o = volume,
    issue_o = issue,
    pages_o = page,
 apa_ref_o = apa_reference_md(doi, title, author_o, journal_o, year_o, volume_o, issue_o, pages_o),
 bibtex_ref_o = bibtex_reference(doi, title, author_o, journal_o, year_o, volume_o, issue_o, pages_o)
  )

doi_data_r <- get_doi_data(flora$doi_r)

doi_data_r_parsed <- doi_data_r %>% 
  transmute(
    doi = tolower(doi),
    title_r = title,
    author_r = map_chr(author, jsonify_authors),
    journal_r = container.title,
    year_r = safe_year(pick(everything())),
    volume_r = volume,
    issue_r = issue,
    pages_r = page,
    apa_ref_r = apa_reference_md(doi, title, author_r, journal_r, year_r, volume_r, issue_r, pages_r),
    bibtex_ref_r = bibtex_reference(doi, title, author_r, journal_r, year_r, volume_r, issue_r, pages_r)
  )

flora <- flora %>% 
  rowwise() %>% 
  mutate(doi_o = tolower(doi_o),
         doi_r = tolower(doi_r),
         doi_o_hash = str_sub(digest::digest(doi_o, "md5", serialize = FALSE), 1, 3),
         doi_r_hash = str_sub(digest::digest(doi_r, "md5", serialize = FALSE), 1, 3)
         ) %>%
  ungroup() %>% 
  left_join(doi_data_o_parsed, by = c("doi_o" = "doi")) %>%
  left_join(doi_data_r_parsed, by = c("doi_r" = "doi")) %>%
  select(doi_o, doi_o_hash, title_o, author_o, journal_o, year_o, volume_o, issue_o, pages_o, apa_ref_o, bibtex_ref_o,
         doi_r, doi_r_hash, title_r, author_r, journal_r, year_r, volume_r, issue_r, pages_r, apa_ref_r, bibtex_ref_r,
         ref_o, ref_r, url_r, outcome, outcome_quote, out_quote_source) %>%
  filter(!is.na(title_o), !is.na(title_r))
```

## Keywords

```{r keywords}

get_openalex_keywords <- function(doi, mailto = NULL, fallback_to_concepts = TRUE, n_concepts = 10) {
  base_url <- "https://api.openalex.org/works/"
  # Encode only the DOI part so slashes become %2F but keep "doi:" literal
  encoded_doi <- URLencode(doi, reserved = TRUE)
  url <- paste0(base_url, "doi:", encoded_doi)

  q <- list(select = "id,doi,display_name,keywords,concepts")
  if (!is.null(mailto)) q$mailto <- mailto

  resp <- GET(url, query = q, user_agent("R script (OpenAlex lookup)"))
  if (http_error(resp)) {
    stop(sprintf("HTTP %s: %s", status_code(resp), content(resp, as = "text", encoding = "UTF-8")))
  }
  x <- fromJSON(content(resp, as = "text", encoding = "UTF-8"), simplifyVector = TRUE)

  # Try explicit keywords first
  kws <- NULL
  if (!is.null(x$keywords)) {
    if (is.character(x$keywords)) {
      kws <- x$keywords
    } else if (is.data.frame(x$keywords)) {
      nm <- names(x$keywords)
      pick <- intersect(c("display_name", "keyword", "name", "value"), nm)
      if (length(pick)) kws <- unique(na.omit(as.character(x$keywords[[pick[1]]])))
    } else if (is.list(x$keywords)) {
      # list of lists
      kws <- unique(na.omit(as.character(unlist(lapply(
        x$keywords,
        function(k) k[["display_name"]] %||% k[["keyword"]] %||% k[["name"]] %||% k
      )))))
    }
  }

  # Fallback to top-N concepts if no explicit keywords
  if ((is.null(kws) || length(kws) == 0) && fallback_to_concepts && !is.null(x$concepts)) {
    cons <- x$concepts
    if (is.data.frame(cons) && "display_name" %in% names(cons)) {
      if ("score" %in% names(cons)) {
        cons <- cons[order(cons$score, decreasing = TRUE), , drop = FALSE]
      }
      kws <- head(unique(as.character(cons$display_name)), n_concepts)
    } else if (is.list(cons)) {
      # list of concepts
      dn <- unlist(lapply(cons, function(cn) cn[["display_name"]]))
      sc <- unlist(lapply(cons, function(cn) cn[["score"]]))
      ord <- order(ifelse(is.na(sc), 0, sc), decreasing = TRUE)
      kws <- head(unique(as.character(dn[ord])), n_concepts)
    }
  }

  # Clean up
  kws <- unique(trimws(kws))
  kws[nzchar(kws)]
}

`%||%` <- function(a, b) if (!is.null(a) && length(a) > 0) a else b


dois <- unique(flora$doi_o)
dois <- dois[!is.na(dois) & grepl("^10", dois)]
dois <- dois[!is.na(dois)]

# dois is vector of DOIs
kwdat <- data.frame(
  doi_o = dois,
  keywords_o = NA_character_, # variable name of keywords is determined here
  error = NA_character_,
  stringsAsFactors = FALSE
)

# # download from openalex
# for (idx in seq_along(dois)) {
#   res <- tryCatch(
#     {
#       kw <- get_openalex_keywords(dois[idx], mailto = "lukas.roeseler@uni-muenster.de")
#       list(kw = kw, err = NA_character_)
#     },
#     error = function(e) list(kw = character(0), err = conditionMessage(e))
#   )
# 
#   kwdat$keywords[idx] <- if (length(res$kw) > 0) paste(res$kw, collapse = ", ") else NA_character_
#   kwdat$error[idx]    <- res$err
# }

# open already downloaded keywords
kwdat <- read.csv("https://raw.githubusercontent.com/forrtproject/FReD-data/refs/heads/main/data%20augmentation/doio_keywords.csv")
kwdat <- kwdat[ ,-1]

flora <- merge(flora, kwdat[, 1:2], by = "doi_o")
```

## Manually coded values

```{r code metapapers}

# code any of these entries as meta papers
metapapers <- c(
"10.1038--s41562-024-02062-9",
"10.3389/fcomm.2022.1048896", # 10 cases of sensory research
"10.1073/pnas.2103313118", # scarcity effects
"10.1177/0956797619831612", # Soto
"10.1007/s13164-018-0400-9", # X-PHI
"10.1038/s41562-018-0399-z", # Camerer 2018
"10.1098/rsos.231240", # Boyce
"10.1027/1864-9335/a000178", # ML1
"10.1177/2515245918810225", # ML2
"10.1016/j.jesp.2015.10.012", # ML3
# "10.1177/2515245920958687", # ML5 NOT because it is focusing on a single study
"10.1126/science.aac4716", # RP:P
"10.1038/s41562-024-02062-9", # Holzmeister et al. 2024
"10.1177/08902070221094216", # SVO development
"10.1126/science.aaf0918" # Camerer 2016
)


# determine metapaper_r = 1 for metapapers
flora$metapaper_r <- as.integer(flora$doi_r %in% metapapers)


```

## Exclusions

```{r Exclusion of cases}

# TO DO: [check others]

# show meta studies
metarefs <- unique(flora[flora$metapaper_r == 1, "ref_r"])

# keep only those metastudies with url_r non missing: Boyce, Soto, XPHI
keep <- !(flora$metapaper_r == 1 & is.na(flora$url_r))
flora <- flora[keep, ]
```

## Peer Review Status

```{r review_r}

# NA = working paper
# "Thesis"
# all else: "Peer-reviewed (published in a journal)"
# preprint servers: "Not peer-reviewed (e.g., preprint)"
preprint_servers <- c(
"SSRN Electronic Journal"
, "Open Research Europe"
)


flora <- flora %>%
mutate(
publicationstatus_r = case_when(
is.na(journal_r) ~ "Working paper",
journal_r %in% .env$preprint_servers ~ "Not peer-reviewed (e.g., preprint)",
TRUE ~ "Peer-reviewed (published in a journal)"
)
)

```

## Variable names

```{r renaming}

# nothing is happening here, yet

varnames <- data.frame("old" = names(flora)
                            , "new" = NA)
# varnames$new <- dplyr::case_match(varnames$old, )
```

## Save dataset

```{r saving}
write_csv(flora, "libre2.csv")
openxlsx::write.xlsx(flora, file = "libre2.xlsx")
```
