---
title: "FRiDa Creator"
format: html
---

# LibRe²

The successor to the meta-analyses-focused FORRT Replication Database (FReD) is a library of replications **and reproductions** that does not include methodological features such as sample and effect sizes.

Notes: This is heavily based on the fred-mini creator. Currently, it includes code to download crossref keywords. It is, however, also possible to attach the already downloaded keywords instead of getting them all from scratch.

## TO DOs

-   add further variables

```{r setup}
library(dplyr)
library(readr)
library(rcrossref)
library(purrr)
library(jsonlite)
library(glue)
library(stringr)
library(rlang)
library(httr)

reported_success <- read_csv(
  "https://docs.google.com/spreadsheets/d/e/2PACX-1vSlPVImL7kjzMfpfX08JtAikvEiBg8uQ3HyEJs-YvuQMUa2RMvip-4v8-DilWl2PdwOD_UTD8gJ5wKA/pub?gid=984458430&single=true&output=csv",
  skip = 1,
  show_col_types = FALSE
) %>%
  # ✅ validate DOI prefix: starts with 10., has 4–9 digits, and then a slash
  filter(str_detect(doi_o, regex("^10\\.[0-9]{4,9}/", ignore_case = TRUE))) %>%
  filter(str_detect(doi_r, regex("^10\\.[0-9]{4,9}/", ignore_case = TRUE)))

```

## Select relevant fields and deduplicate (we do not report study level)

```{r deduplication}
fred_mini <- reported_success %>% filter(validation == "validated - chosen")  %>% select(doi_o, doi_r, ref_o, ref_r, url_r, outcome, outcome_quote, out_quote_source) %>%
  group_by(doi_o, doi_r, ref_o, ref_r, url_r) %>%
  summarise(
  outcome = if (length(unique(outcome)) == 1) unique(outcome) else "mixed",
  outcome_quote = if (length(unique(outcome_quote)) == 1) unique(outcome_quote) else paste(outcome_quote, collapse = "; "),
  out_quote_source = paste(out_quote_source, collapse = "; "),
  .groups = "drop"
  ) %>% 
  distinct(doi_o,doi_r,.keep_all = TRUE)
```

## Augment with bibliometric details

```{r bibliometric augmentation}
get_doi_data <- function(doi) {
  #100 at a time to avoid URL too long CR error
  doi <- str_replace_all(doi, "https?://doi.org/", "")
  doi <- str_subset(doi, "^10\\.\\d+/\\S+$")
  doi <- unique(doi)
  
  message(glue("Fetching data for {length(doi)} DOIs from CrossRef..."))
  
  if (length(doi) == 0) return(tibble())
  
  for (i in seq(1, length(doi), by = 100)) {
    doisub <- doi[i:min(i+99, length(doi))]
    names(doisub) <- rep("doi", length(doisub))
    res <- cr_works(filter = doisub, limit = 100)
    message(glue("  Retrieved {nrow(res$data)} records (of {length(doisub)} requested)"))
    if(i == 1) out <- res$data else out <- bind_rows(out, res$data)
  }
  message(glue("Done. Retrieved {nrow(out)} records in total."))
  
  return(out)
}


jsonify_authors <- function(author_list) {
  
  if (all(c("family", "given") %in% names(author_list))) {
    out <-   tibble(
    given = author_list[["given"]] %||% NA_character_,
    family = author_list[["family"]] %||% NA_character_,
    sequence = author_list[["sequence"]] %||% NA_character_,
    ORCID = author_list[["ORCID"]] %||% NA_character_
  )
  } else if ("name" %in% names(author_list)) {
      out <-   tibble(
    given = author_list[["given"]] %||% NA_character_,
    family = author_list[["name"]] %||% NA_character_,
    sequence = author_list[["sequence"]] %||% NA_character_,
    ORCID = author_list[["ORCID"]] %||% NA_character_
  )

} else {
  message("Unknown author format - field names: ", paste(names(author_list), collapse = ", "), ". Returning NULL entry.")
        out <-   tibble(
    given = author_list[["given"]] %||% NA_character_,
    family = author_list[["name"]] %||% NA_character_,
    sequence = author_list[["sequence"]] %||% NA_character_,
    ORCID = author_list[["ORCID"]] %||% NA_character_
  )
}
  
 out |>
    toJSON(pretty = FALSE, na = "null")
}


safe_year <- function(df) {
  pub_cols <- c("published.print", "published.online", "published")
  existing <- pub_cols[pub_cols %in% names(df)]

  if (length(existing) == 0) {
    return(NA_integer_)
  }

  year_str <- dplyr::coalesce(!!!df[existing])
  as.integer(stringr::str_sub(as.character(year_str), 1, 4))
}

apa_reference_md <- function(doi, title, author_json, journal, year, volume, issue, pages) {
  n <- length(doi)

  map_chr(seq_len(n), function(i) {
    # --- parse authors safely ---
    authors <- tryCatch(fromJSON(author_json[i]), error = function(e) NULL)

    fmt_author <- function(given, family) {
      initials <- if (!is.na(given) && nzchar(given)) {
        parts <- strsplit(given, "\\s+")[[1]]
        paste0(substr(parts, 1, 1), collapse = ". ")
      } else ""
      paste0(family, ", ", initials, ".")
    }

    if (is.null(authors) || nrow(authors) == 0) {
      author_str <- "(n.a.)"
    } else {
      author_strs <- pmap_chr(authors, \(given, family, sequence, ORCID)
        fmt_author(given, family))
      n_auth <- length(author_strs)
      author_str <-
        if (n_auth == 1) author_strs
        else if (n_auth == 2) paste(author_strs, collapse = " & ")
        else paste0(paste(author_strs[-n_auth], collapse = ", "), ", & ", author_strs[n_auth])
    }

    # --- APA-style markdown (title not bold) ---
    glue(
      "{author_str} ({year[i]}). {title[i]}. *{journal[i]}*, *{volume[i]}*({issue[i]}), {pages[i]}. [https://doi.org/{doi[i]}](https://doi.org/{doi[i]})"
    )
  })
}

bibtex_reference <- function(doi, title, author_json, journal, year, volume, issue, pages) {
  n <- length(doi)

  map_chr(seq_len(n), function(i) {
    # --- parse authors safely ---
    authors <- tryCatch(fromJSON(author_json[i]), error = function(e) NULL)

    fmt_author_bib <- function(given, family) {
      if (!is.na(given) && nzchar(given)) {
        paste0(family, ", ", given)
      } else {
        family
      }
    }

    if (is.null(authors) || nrow(authors) == 0) {
      author_str <- "Unknown"
    } else {
      author_strs <- pmap_chr(authors, \(given, family, sequence, ORCID)
        fmt_author_bib(given, family))
      author_str <- paste(author_strs, collapse = " and ")
    }

    key <- str_remove_all(paste0(str_to_lower(word(author_strs[1], 1)), year[i]), "[^a-z0-9]")

    glue(
"@article{{{key},
  author = {{{author_str}}},
  title = {{{title[i]}}},
  journal = {{{journal[i]}}},
  year = {{{year[i]}}},
  volume = {{{volume[i]}}},
  number = {{{issue[i]}}},
  pages = {{{pages[i]}}},
  doi = {{{doi[i]}}}
}}"
    )
  })
}

doi_data_o <- get_doi_data(fred_mini$doi_o)

doi_data_o_parsed <- doi_data_o %>% 
  transmute(
    doi = tolower(doi),
    title_o = title,
    author_o = map_chr(author, jsonify_authors),
    journal_o = container.title,
 year_o = safe_year(pick(everything())),
    volume_o = volume,
    issue_o = issue,
    pages_o = page,
 apa_ref_o = apa_reference_md(doi, title, author_o, journal_o, year_o, volume_o, issue_o, pages_o),
 bibtex_ref_o = bibtex_reference(doi, title, author_o, journal_o, year_o, volume_o, issue_o, pages_o)
  )

doi_data_r <- get_doi_data(fred_mini$doi_r)

doi_data_r_parsed <- doi_data_r %>% 
  transmute(
    doi = tolower(doi),
    title_r = title,
    author_r = map_chr(author, jsonify_authors),
    journal_r = container.title,
    year_r = safe_year(pick(everything())),
    volume_r = volume,
    issue_r = issue,
    pages_r = page,
    apa_ref_r = apa_reference_md(doi, title, author_r, journal_r, year_r, volume_r, issue_r, pages_r),
    bibtex_ref_r = bibtex_reference(doi, title, author_r, journal_r, year_r, volume_r, issue_r, pages_r)
  )

fred_mini <- fred_mini %>% 
  rowwise() %>% 
  mutate(doi_o = tolower(doi_o),
         doi_r = tolower(doi_r),
         doi_o_hash = str_sub(digest::digest(doi_o, "md5", serialize = FALSE), 1, 3),
         doi_r_hash = str_sub(digest::digest(doi_r, "md5", serialize = FALSE), 1, 3)
         ) %>%
  ungroup() %>% 
  left_join(doi_data_o_parsed, by = c("doi_o" = "doi")) %>%
  left_join(doi_data_r_parsed, by = c("doi_r" = "doi")) %>%
  select(doi_o, doi_o_hash, title_o, author_o, journal_o, year_o, volume_o, issue_o, pages_o, apa_ref_o, bibtex_ref_o,
         doi_r, doi_r_hash, title_r, author_r, journal_r, year_r, volume_r, issue_r, pages_r, apa_ref_r, bibtex_ref_r,
         ref_o, ref_r, url_r, outcome, outcome_quote, out_quote_source) %>%
  filter(!is.na(title_o), !is.na(title_r))
```

## Open Alex Data

-   keywords

-   field (e.g, psychology)

-   language

```{r get openalex data}
# keywords, language, field


'%||%' <- function(a, b) if (!is.null(a) && length(a) > 0) a else b

get_openalex_meta <- function(doi, mailto = NULL, fallback_to_concepts = TRUE, n_concepts = 10) {
base_url <- "https://api.openalex.org/works/"
encoded_doi <- URLencode(doi, reserved = TRUE)
url <- paste0(base_url, "doi:", encoded_doi)

q <- list(select = "id,doi,display_name,language,keywords,concepts,primary_topic")
if (!is.null(mailto)) q$mailto <- mailto

resp <- GET(url, query = q, user_agent("R script (OpenAlex lookup)"))
if (http_error(resp)) {
stop(sprintf("HTTP %s: %s", status_code(resp), content(resp, as = "text", encoding = "UTF-8")))
}
x <- fromJSON(content(resp, as = "text", encoding = "UTF-8"), simplifyVector = TRUE)

# language
lang <- x$language %||% NA_character_

# keywords (as in your original)
kws <- NULL
if (!is.null(x$keywords)) {
if (is.character(x$keywords)) {
kws <- x$keywords
} else if (is.data.frame(x$keywords)) {
nm <- names(x$keywords)
pick <- intersect(c("display_name", "keyword", "name", "value"), nm)
if (length(pick)) kws <- unique(na.omit(as.character(x$keywords[[pick[1]]])))
} else if (is.list(x$keywords)) {
kws <- unique(na.omit(as.character(unlist(lapply(
x$keywords,
function(k) k[["display_name"]] %||% k[["keyword"]] %||% k[["name"]] %||% k
)))))
}
}

# Fallback to top-N concepts for keywords if needed
if ((is.null(kws) || length(kws) == 0) && fallback_to_concepts && !is.null(x$concepts)) {
cons <- x$concepts
if (is.data.frame(cons) && "display_name" %in% names(cons)) {
if ("score" %in% names(cons)) cons <- cons[order(cons$score, decreasing = TRUE), , drop = FALSE]
kws <- head(unique(as.character(cons$display_name)), n_concepts)
} else if (is.list(cons)) {
dn <- unlist(lapply(cons, function(cn) cn[["display_name"]]))
sc <- unlist(lapply(cons, function(cn) cn[["score"]]))
ord <- order(ifelse(is.na(sc), 0, sc), decreasing = TRUE)
kws <- head(unique(as.character(dn[ord])), n_concepts)
}
}
kws <- unique(trimws(kws))
kws <- kws[nzchar(kws)]

# field: prefer primary_topic$field$display_name, otherwise infer from concepts
field <- NA_character_
if (!is.null(x$primary_topic)) {
pt <- x$primary_topic
if (!is.null(pt$field) && !is.null(pt$field$display_name)) {
field <- as.character(pt$field$display_name)
} else if (!is.null(pt$subfield) && !is.null(pt$subfield$display_name)) {
field <- as.character(pt$subfield$display_name)
} else if (!is.null(pt$display_name)) {
field <- as.character(pt$display_name)
}
}
if ((is.na(field) || !nzchar(field)) && !is.null(x$concepts)) {
cons <- x$concepts
# normalize to vectors
if (is.data.frame(cons)) {
dn <- as.character(cons$display_name %||% rep(NA_character_, nrow(cons)))
lv <- suppressWarnings(as.integer(cons$level %||% rep(NA_integer_, nrow(cons))))
sc <- suppressWarnings(as.numeric(cons$score %||% rep(NA_real_, nrow(cons))))
} else if (is.list(cons)) {
dn <- as.character(unlist(lapply(cons, function(cn) cn[["display_name"]] %||% NA_character_)))
lv <- suppressWarnings(as.integer(unlist(lapply(cons, function(cn) cn[["level"]] %||% NA_integer_))))
sc <- suppressWarnings(as.numeric(unlist(lapply(cons, function(cn) cn[["score"]] %||% NA_real_))))
} else {
dn <- lv <- sc <- NULL
}
if (!is.null(dn) && length(dn)) {
# prefer level 1 (field), then 0 (domain), else best-scoring
pick_idx <- which(lv == 1)
if (!length(pick_idx)) pick_idx <- which(lv == 0)
if (!length(pick_idx)) pick_idx <- seq_along(dn)
# choose highest scoring among candidates
sc_sub <- ifelse(is.na(sc[pick_idx]), 0, sc[pick_idx])
field <- dn[pick_idx[order(sc_sub, decreasing = TRUE)][1]]
}
}

list(
keywords = if (length(kws)) kws else character(0),
field = field %||% NA_character_,
language = lang %||% NA_character_
)
}

# Prepare DOIs
dois <- unique(fred_mini$doi_o)
dois <- dois[!is.na(dois) & grepl("^10", dois)]

# Download from OpenAlex
kwdat <- data.frame(
doi_o = dois,
keywords_o = NA_character_,
field_o = NA_character_,
language_o = NA_character_,
error = NA_character_,
stringsAsFactors = FALSE
)

for (idx in seq_along(dois)) {
res <- tryCatch(
{
out <- get_openalex_meta(dois[idx], mailto = "lukas.roeseler@uni-muenster.de")
list(out = out, err = NA_character_)
},
error = function(e) list(out = NULL, err = conditionMessage(e))
)

if (!is.null(res$out)) {
kwdat$keywords_o[idx] <- if (length(res$out$keywords) > 0) paste(res$out$keywords, collapse = ", ") else NA_character_
kwdat$field_o[idx] <- res$out$field %||% NA_character_
kwdat$language_o[idx] <- res$out$language %||% NA_character_
}
kwdat$error[idx] <- res$err
}

kwdat

fred_mini <- merge(fred_mini, kwdat[, 1:4], by = "doi_o")
```

## Manually coded values

```{r code metapapers}

# code any of these entries as meta papers
metapapers <- c(
"10.1038--s41562-024-02062-9",
"10.3389/fcomm.2022.1048896", # 10 cases of sensory research
"10.1073/pnas.2103313118", # scarcity effects
"10.1177/0956797619831612", # Soto
"10.1007/s13164-018-0400-9", # X-PHI
"10.1038/s41562-018-0399-z", # Camerer 2018
"10.1098/rsos.231240", # Boyce
"10.1027/1864-9335/a000178", # ML1
"10.1177/2515245918810225", # ML2
"10.1016/j.jesp.2015.10.012", # ML3
# "10.1177/2515245920958687", # ML5 NOT because it is focusing on a single study
"10.1126/science.aac4716", # RP:P
"10.1038/s41562-024-02062-9", # Holzmeister et al. 2024
"10.1177/08902070221094216", # SVO development
"10.1126/science.aaf0918" # Camerer 2016
)


# determine metapaper_r = 1 for metapapers
fred_mini$metapaper_r <- as.integer(fred_mini$doi_r %in% metapapers)


```

## Exclusions

```{r Exclusion of cases}

# TO DO: [check others]

# show meta studies
metarefs <- unique(fred_mini[fred_mini$metapaper_r == 1, "ref_r"])

# keep only those metastudies with url_r non missing: Boyce, Soto, XPHI
keep <- !(fred_mini$metapaper_r == 1 & is.na(fred_mini$url_r))
fred_mini <- fred_mini[keep, ]
```

## Peer Review Status

```{r review_r}

# NA = working paper
# "Thesis"
# all else: "Peer-reviewed (published in a journal)"
# preprint servers: "Not peer-reviewed (e.g., preprint)"
preprint_servers <- c(
"SSRN Electronic Journal"
, "Open Research Europe"
)


fred_mini <- fred_mini %>%
mutate(
publicationstatus_r = case_when(
is.na(journal_r) ~ "Working paper",
journal_r %in% .env$preprint_servers ~ "Not peer-reviewed (e.g., preprint)",
TRUE ~ "Peer-reviewed (published in a journal)"
)
)

```

## Variable names

```{r renaming}

# nothing is happening here, yet

varnames <- data.frame("old" = names(fred_mini)
                            , "new" = NA)
# varnames$new <- dplyr::case_match(varnames$old, )
```

## Save dataset

```{r saving}
write_csv(fred_mini, "libre2.csv")
openxlsx::write.xlsx(fred_mini, file = "libre2.xlsx")
```
