---
title: "Dataset release"
author: "Lukas Wallrich"
date: "2024-02-02"
output: html_document
---

## Instructions

This set of functions can be used to automatically release and track FReD and FLoRA dataset updates.

For this, the user needs to create an OSF Personal Access Token [here](https://osf.io/settings/tokens) and (ideally) save it to an environment variable.

```{r}
source("release_dataset.R")
Sys.setenv("OSF_TOKEN" = "YOUR OSF TOKEN")
```

## Releasing new versions

Only a single function call is needed to:
- Bump the dataset version and log changes
- Archive the previous dataset
- Download the current dataset and release it on OSF

To distinguish magnitude of changes, version numbers can be split into major, minor and patch releases. All that needs to be specified are release notes and the magnitude of the bump.

To note:
- Currently, this works on a dummy repository - for production, the IDs of project and version log need to be changed in release_dataset.R
- It cannot update nested folders - so to use this, the Data folder would need to be at the top level. (See [osfr issue](https://github.com/ropensci/osfr/issues/130))

```{r}
release_notes <- glue::glue("- First automated release")

release_new_version(release_notes, type = "FReD", version = "patch") #Or major/minor

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
