---
title: "Duplicate Detection Report"
author: "FReD Data Team"
date: "`r Sys.Date()`"
format:
  html:
    embed-resources: true
    toc: true
    self-contained: true
editor: visual
---

------------------------------------------------------------------------

# Duplicate Detection Report

This report identifies potential duplicate rows in the FRED dataset based on matching `n_r` (replication sample size) and `es_value_r` (replication effect size) values. It shows whether these potential duplicates also overlap on other identifiers.

------------------------------------------------------------------------

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}
library(FReD)
library(lubridate)
library(ggplot2)
library(dplyr)
library(stringr)
library(tidyverse)
library(glue)
library(DT)
source("https://raw.githubusercontent.com/forrtproject/FReD-data/refs/heads/main/clean_cos_fred.R")
```

## Data Loading and Cleaning

This section replicates the data processing from cos_report.qmd to ensure consistency.

```{r load-and-clean, include = FALSE, echo = FALSE, warning = FALSE, message = FALSE}

# Load raw data from Google Sheets
ds_fully_raw <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vR_62H8liC0P3-WQ1xx4UalDnvqzPfObWBFwQAgVH6JhF3dPBFXd7AgFK_OljxFWQkNtUYdnUbuFTpe/pub?output=csv", skip = 1)

# Clean data
cleaned <- clean_fred_data(ds_fully_raw)
ds_raw <- cleaned$cleaned_data

# Helper function to clean filenames
replace_na_filename <- function(filename) {
  filename %>%
    na_if(".pdf") %>%
    na_if("na.pdf") %>%
    na_if("NA - USE URL")
}

ds_raw <- ds_raw %>%
  mutate(file_o = replace_na_filename(file_o),
         file_r = replace_na_filename(file_r))

# Apply main filters
ds_raw <- ds_raw %>%
  # filter empty rows
  filter(!is.na(ref_o) | !is.na(ref_r)) %>%
  # add ID to distinguish replications from same meta-paper
  mutate(id_r = coalesce(url_r, tolower(doi_r))) %>%
  # filter out excluded studies
  filter(is.na(exclusion)) %>%
  # filter out non-validated extra rows
  filter(!(approach == "extra rows" & status != "validated - chosen") |
         is.na(approach) | is.na(status))

# Filter for validated - chosen entries
ds <- ds_raw %>% filter(status == "validated - chosen")

# Update references with crossref data
refs <- try(readxl::read_excel("../crossref_retrieved_references.xlsx"), silent = TRUE)

if (!inherits(refs, "try-error")) {
  ds <- ds %>%
    mutate(doi_r = tolower(doi_r),
           doi_o = tolower(doi_o)) %>%
    left_join(
      refs %>% rename(doi_o = doi, ref_o_new = reference)
    ) %>% left_join(
      refs %>% rename(doi_r = doi, ref_r_new = reference)
    ) %>%
    mutate(ref_o = coalesce(ref_o_new, ref_o),
           ref_r = coalesce(ref_r_new, ref_r)) %>%
    select(-ref_o_new, -ref_r_new)
}

```

## Identifying Potential Duplicates

The following analysis identifies rows that share the same `n_r` (replication sample size) and `es_value_r` (replication effect size) values, as these combinations may indicate duplicate entries.

```{r identify-duplicates, echo = FALSE, warning = FALSE, message = FALSE}

# Find groups with duplicate n_r and es_value_r
duplicates <- ds %>%
  group_by(n_r, es_value_r) %>%
  filter(n() > 1) %>%
  ungroup() %>%
  arrange(n_r, es_value_r)

# Check if there are any duplicates
if (nrow(duplicates) == 0) {
  cat("✓ **No duplicate combinations of n_r and es_value_r found in the dataset.**\n\n")
} else {
  cat(glue("Found {n_distinct(duplicates %>% select(n_r, es_value_r))} unique combinations of (n_r, es_value_r) that appear multiple times.\n\n"))

  # Create a summary table showing overlap information
  overlap_summary <- duplicates %>%
    group_by(n_r, es_value_r) %>%
    summarise(
      count = n(),
      unique_entries = n_distinct(entry_id),
      unique_id_r = n_distinct(id_r),
      unique_ref_o = n_distinct(ref_o),
      unique_ref_r = n_distinct(ref_r),
      unique_fred_id = n_distinct(fred_id),
      unique_n_o = n_distinct(n_o),
      unique_es_value_o = n_distinct(es_value_o),
      .groups = "drop"
    ) %>%
    mutate(
      id_r_overlap = case_when(
        unique_id_r == 1 ~ "✓ All same",
        TRUE ~ "✗ Different"
      ),
      ref_o_overlap = case_when(
        unique_ref_o == 1 ~ "✓ All same",
        TRUE ~ "✗ Different"
      ),
      ref_r_overlap = case_when(
        unique_ref_r == 1 ~ "✓ All same",
        TRUE ~ "✗ Different"
      ),
      fred_id_overlap = case_when(
        unique_fred_id == 1 ~ "✓ All same",
        TRUE ~ "✗ Different"
      ),
      original_overlap = case_when(
        unique_n_o == 1 & unique_es_value_o == 1 ~ "✓ Both same",
        unique_n_o == 1 ~ "✓ n_o same",
        unique_es_value_o == 1 ~ "✓ es_value_o same",
        TRUE ~ "✗ Both differ"
      )
    ) %>%
    select(n_r, es_value_r, count, original_overlap, id_r_overlap, ref_o_overlap, ref_r_overlap, fred_id_overlap)

  # Prepare detailed rows for each group (for expansion)
  detailed_rows <- duplicates %>%
    select(n_r, es_value_r, entry_id, fred_id, ref_o, id_r, ref_r, study_r, doi_r) %>%
    arrange(n_r, es_value_r, entry_id)

  # Create child row HTML for each summary row
  child_rows_html <- list()
  for (i in 1:nrow(overlap_summary)) {
    n_val <- overlap_summary$n_r[i]
    es_val <- overlap_summary$es_value_r[i]

    group_details <- detailed_rows %>%
      filter(n_r == n_val, es_value_r == es_val) %>%
      select(entry_id, fred_id, ref_o, id_r, ref_r, study_r, doi_r)

    # Create HTML table for this group
    html_content <- "<div style='padding: 15px; background-color: #fafafa;'><table style='width: 100%; border-collapse: collapse; border: 1px solid #ddd;'>"
    html_content <- paste0(html_content,
      "<tr style='background-color: #e8f5e9;'><th style='border: 1px solid #ddd; padding: 8px; text-align: left;'>entry_id</th>",
      "<th style='border: 1px solid #ddd; padding: 8px; text-align: left;'>fred_id</th>",
      "<th style='border: 1px solid #ddd; padding: 8px; text-align: left;'>ref_o</th>",
      "<th style='border: 1px solid #ddd; padding: 8px; text-align: left;'>id_r</th>",
      "<th style='border: 1px solid #ddd; padding: 8px; text-align: left;'>ref_r</th>",
      "<th style='border: 1px solid #ddd; padding: 8px; text-align: left;'>study_r</th>",
      "<th style='border: 1px solid #ddd; padding: 8px; text-align: left;'>doi_r</th></tr>")

    for (j in 1:nrow(group_details)) {
      row <- group_details[j, ]
      html_content <- paste0(html_content,
        "<tr style='border-bottom: 1px solid #ddd;'>",
        "<td style='border: 1px solid #ddd; padding: 8px;'>", row$entry_id, "</td>",
        "<td style='border: 1px solid #ddd; padding: 8px;'><strong>", row$fred_id, "</strong></td>",
        "<td style='border: 1px solid #ddd; padding: 8px;'>", row$ref_o, "</td>",
        "<td style='border: 1px solid #ddd; padding: 8px;'><strong>", row$id_r, "</strong></td>",
        "<td style='border: 1px solid #ddd; padding: 8px;'>", row$ref_r, "</td>",
        "<td style='border: 1px solid #ddd; padding: 8px;'>", row$study_r, "</td>",
        "<td style='border: 1px solid #ddd; padding: 8px;'>", row$doi_r, "</td>",
        "</tr>")
    }
    html_content <- paste0(html_content, "</table></div>")
    child_rows_html[[i]] <- html_content
  }

  # Add expand column to overlap_summary (in first position)
  overlap_summary_display <- overlap_summary %>%
    mutate(expand = "▶ Click to expand") %>%
    select(expand, everything())

  # Color code the table with expandable rows
  datatable(
    overlap_summary_display,
    colnames = c(
      "Expand",
      "Sample Size (n_r)",
      "Effect Size (es_value_r)",
      "# Rows",
      "Original (n_o, es_value_o)",
      "id_r",
      "ref_o",
      "ref_r",
      "fred_id"
    ),
    escape = FALSE,
    options = list(
      pageLength = 50,
      dom = 'lftp',
      scrollX = TRUE,
      ordering = FALSE,
      columnDefs = list(
        list(
          targets = 0,
          render = JS("function(data, type, row) {
            return '<span style=\"cursor: pointer; color: #0066cc; font-weight: bold; user-select: none;\">' + data + '</span>';
          }")
        )
      )
    ),
    callback = JS(
      paste0("
      var colors = {'✓ All same': '#d4edda', '✓ Both same': '#c8e6c9', '✓ n_o same': '#e0f2f1', '✓ es_value_o same': '#e0f2f1', '✗ Different': '#f8d7da', '✗ Both differ': '#ffcdd2'};
      var childData = ", jsonlite::toJSON(child_rows_html), ";

      // Color the overlap columns (now 5-8 instead of 4-7)
      for (var i = 5; i <= 8; i++) {
        for (var row = 0; row < table.rows().count(); row++) {
          var cell = $('td:eq(' + i + ')', table.row(row).node());
          var text = cell.text();
          if (colors[text]) {
            cell.css('background-color', colors[text]);
            cell.css('font-weight', 'bold');
          }
        }
      }

      // Handle expand/collapse
      $('#' + table.table().node().id + ' tbody').on('click', 'tr', function() {
        var tr = $(this);
        var row = table.row(this);
        var rowIdx = table.rows({order:'current'}).indexes().indexOf(table.row(this).index());

        if (row.child.isShown()) {
          row.child.hide();
          tr.removeClass('shown');
          tr.find('span:first').text('▶ Click to expand');
        } else {
          row.child(childData[rowIdx]).show();
          tr.addClass('shown');
          tr.find('span:first').text('▼ Click to collapse');
        }
      });
      ")
    )
  )
}

```


## Summary Statistics

```{r summary, echo = FALSE, warning = FALSE, message = FALSE}

if (nrow(duplicates) > 0) {

  summary_stats <- tibble(
    Metric = c(
      "Total rows in dataset",
      "Rows involved in potential duplicates",
      "Unique (n_r, es_value_r) combinations with duplicates",
      "Percentage of rows in duplicates"
    ),
    Value = c(
      nrow(ds),
      nrow(duplicates),
      n_distinct(duplicates %>% select(n_r, es_value_r)),
      paste0(round(100 * nrow(duplicates) / nrow(ds), 2), "%")
    )
  )

  datatable(summary_stats, options = list(dom = 't', paging = FALSE))

} else {
  cat("No duplicates found. Summary: All {nrow(ds)} rows have unique combinations of (n_r, es_value_r).\n")
}

```

## Groups with Matching Original Study Values

The following groups have both `n_o` and `es_value_o` matching across all rows (indicating potential true duplicates from the same original study):

```{r original-matches, echo = FALSE, warning = FALSE, message = FALSE}

if (nrow(duplicates) > 0) {
  original_matches <- overlap_summary %>%
    filter(original_overlap == "✓ Both same") %>%
    select(n_r, es_value_r, count, original_overlap, id_r_overlap, ref_o_overlap, fred_id_overlap)

  if (nrow(original_matches) > 0) {
    cat(glue("**Found {nrow(original_matches)} group(s) with matching original study values:**\n\n"))

    # Create child rows for original matches
    original_child_rows <- list()
    for (i in 1:nrow(original_matches)) {
      n_val <- original_matches$n_r[i]
      es_val <- original_matches$es_value_r[i]

      group_details <- detailed_rows %>%
        filter(n_r == n_val, es_value_r == es_val) %>%
        select(entry_id, fred_id, ref_o, id_r, ref_r, study_r, doi_r)

      # Create HTML table for this group
      html_content <- "<div style='padding: 15px; background-color: #fafafa;'><table style='width: 100%; border-collapse: collapse; border: 1px solid #ddd;'>"
      html_content <- paste0(html_content,
        "<tr style='background-color: #e8f5e9;'><th style='border: 1px solid #ddd; padding: 8px; text-align: left;'>entry_id</th>",
        "<th style='border: 1px solid #ddd; padding: 8px; text-align: left;'>fred_id</th>",
        "<th style='border: 1px solid #ddd; padding: 8px; text-align: left;'>ref_o</th>",
        "<th style='border: 1px solid #ddd; padding: 8px; text-align: left;'>id_r</th>",
        "<th style='border: 1px solid #ddd; padding: 8px; text-align: left;'>ref_r</th>",
        "<th style='border: 1px solid #ddd; padding: 8px; text-align: left;'>study_r</th>",
        "<th style='border: 1px solid #ddd; padding: 8px; text-align: left;'>doi_r</th></tr>")

      for (j in 1:nrow(group_details)) {
        row <- group_details[j, ]
        html_content <- paste0(html_content,
          "<tr style='border-bottom: 1px solid #ddd;'>",
          "<td style='border: 1px solid #ddd; padding: 8px;'>", row$entry_id, "</td>",
          "<td style='border: 1px solid #ddd; padding: 8px;'><strong>", row$fred_id, "</strong></td>",
          "<td style='border: 1px solid #ddd; padding: 8px;'>", row$ref_o, "</td>",
          "<td style='border: 1px solid #ddd; padding: 8px;'><strong>", row$id_r, "</strong></td>",
          "<td style='border: 1px solid #ddd; padding: 8px;'>", row$ref_r, "</td>",
          "<td style='border: 1px solid #ddd; padding: 8px;'>", row$study_r, "</td>",
          "<td style='border: 1px solid #ddd; padding: 8px;'>", row$doi_r, "</td>",
          "</tr>")
      }
      html_content <- paste0(html_content, "</table></div>")
      original_child_rows[[i]] <- html_content
    }

    # Add expand column
    original_matches_display <- original_matches %>%
      mutate(expand = "▶ Click to expand") %>%
      select(expand, everything())

    datatable(
      original_matches_display,
      colnames = c(
        "Expand",
        "Sample Size (n_r)",
        "Effect Size (es_value_r)",
        "# Rows",
        "Original Values",
        "id_r",
        "ref_o",
        "fred_id"
      ),
      escape = FALSE,
      options = list(
        pageLength = 50,
        dom = 'lftp',
        scrollX = TRUE,
        ordering = FALSE,
        columnDefs = list(
          list(
            targets = 0,
            render = JS("function(data, type, row) {
              return '<span style=\"cursor: pointer; color: #0066cc; font-weight: bold; user-select: none;\">' + data + '</span>';
            }")
          )
        )
      ),
      callback = JS(
        paste0("
        var colors = {'✓ All same': '#d4edda', '✗ Different': '#f8d7da'};
        var childData = ", jsonlite::toJSON(original_child_rows), ";

        // Color the overlap columns
        for (var i = 5; i <= 7; i++) {
          for (var row = 0; row < table.rows().count(); row++) {
            var cell = $('td:eq(' + i + ')', table.row(row).node());
            var text = cell.text();
            if (colors[text]) {
              cell.css('background-color', colors[text]);
              cell.css('font-weight', 'bold');
            }
          }
        }

        // Handle expand/collapse
        $('#' + table.table().node().id + ' tbody').on('click', 'tr', function() {
          var tr = $(this);
          var row = table.row(this);
          var rowIdx = table.rows({order:'current'}).indexes().indexOf(table.row(this).index());

          if (row.child.isShown()) {
            row.child.hide();
            tr.removeClass('shown');
            tr.find('span:first').text('▶ Click to expand');
          } else {
            row.child(childData[rowIdx]).show();
            tr.addClass('shown');
            tr.find('span:first').text('▼ Click to collapse');
          }
        });
        ")
      )
    )
  } else {
    cat("**No groups with matching original study values found.** Rows within duplicate groups differ on either `n_o` or `es_value_o` (or both).\n")
  }
} else {
  cat("No duplicate groups to analyze.\n")
}

```

## Legend

### Status Colors:
- **✓ All same** (light green): All rows in this duplicate group share the same value for this field
- **✓ Both same** (medium green): Both `n_o` and `es_value_o` match across all rows
- **✓ n_o same** / **✓ es_value_o same** (light teal): Only one of the original values matches
- **✗ Different** (light red): Rows in this duplicate group differ on this field
- **✗ Both differ** (darker red): Both `n_o` and `es_value_o` differ across rows

### Fields Checked:
- **n_r**: Replication sample size
- **es_value_r**: Replication effect size value
- **n_o**: Original study sample size
- **es_value_o**: Original study effect size value
- **id_r**: Replication identifier (URL or DOI)
- **ref_o**: Original study reference
- **ref_r**: Replication study reference
- **fred_id**: FRED database identifier
